### 2018-11-07 paper/paper.Rmd
start paper outline

---
title: "Assessing the quality of Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
author: "Andreas Beger and Michael D. Ward"
date: "`r Sys.Date()`"
output: 
  pdf_document
abstract: "\\noindent HFC \\newline \\newline \\newline
\\textbf{Funding acknowledgement}: This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005. The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S. Government. The U.S. Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Hybrid Forecasting Competition

Background on HFC

## SAGE approach

Cover the basics: there is a platform that can show a chart and machine forecast; how are the machine forecasts generated.

## Research design for RCT-A

Go over the conditions, and the complications that happened: turkers only in two conditions, and the data and forecaster code changes over time.

Two key questions we want to examine:

1. How accurate were Turker forecasts compared to volunteer forecasts?
2. What was the impact of the machine forecasts on human accuracy?

# Design, data, and method

Key leverage point is that only some questions had a machine forecast.

Go over what the data for this are, how accuracy was assessed

# Results

## Basic findings

Machine forecasts were not more accurate than human forecasts.

## Turker performance compared to volunteer forecasters

## Machine forecast impact

# Discussion/conclusion

Based on the quick check I did during the site visit, probably something like this:

- the machine forecasts were not accurate enough, although sans data and code issues they may have been 
- Turkers don't forecast as well as volunteers
- seeing the machine forecasts had a slight detrimental impact on accuracy, although maybe this changed as the forecasts became more accurate later in the RCT
- seeing the charts improved performance, but I don't think it is clear why. Seeing the chart for instances where the data are hard to acquire (e.g. some data transformation are needed) did not boost accuracy more than in other series.

### 2018-11-09 slides/slides.Rmd
add template for xaringan slides

---
title: "draft slides"
author: "Andreas Beger, Predictive Heuristics"
date: "`r format(Sys.Date(), '%e %b %Y')`"
output: xaringan::moon_reader
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<!-- https://apreshill.github.io/data-vis-labs-2018/slides/06-slides_xaringan.html#13 -->

# First slide

---

# Second slides

### 2018-11-09 R/analysis.R
move scripts to R

@@ -2,6 +2,18 @@ library("tidyverse")
 
 dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
 
+
+
+
+user_clean %>%
+  group_by(condition) %>%
+  summarize(total_users = length(unique(user_id)),
+            turkers     = length(unique(user_id[turker])),
+            volunteers  = total_users - turkers,
+            forecasts = n())
+
+
+
 mbrier <- function(f, o, ordered = TRUE) {
   prediction = f
   outcome = o

### 2018-11-09 R/debug-data-read.R
add data maker and debug script

@@ -1,0 +1,33 @@
+#
+#   Why is read_csv not parsing some of the dump_user... lines?
+#
+#   9 November 2018
+#
+
+library("tidyverse")
+library("future.apply")
+
+plan(multiprocess, workers = 3L)
+
+dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
+out_dir <- file.path(dropbox_path, "output")
+
+user_str <- "dump_user_forecasts_09_07_2018_09_15_12.csv" %>%
+  file.path(dropbox_path, "input-data", .) %>%
+  read_lines() 
+
+str_header <- user_str[1]
+str_row    <- as.list(tail(user_str, -1))
+
+# this takes a while
+parsed_rows <- future_lapply(str_row, function(x) {
+  suppressWarnings(read_csv(paste0(c(str_header, x), collapse = "\n")))
+})
+
+write_rds(parsed_rows, file.path(out_dir, "debug_user_dump_read.rds"))
+
+
+user_raw <- "dump_user_forecasts_09_07_2018_09_15_12.csv" %>%
+  file.path(dropbox_path, "input-data", .) %>%
+  read_csv(., quote = "\"") 
+

### 2018-11-09 R/make-data.R

@@ -1,0 +1,157 @@
+#
+#   Process the raw staff dumps to make the analysis data
+#
+
+library("tidyverse")
+library("futile.logger")
+
+dropbox_path  <- "~/Dropbox/Work/asia-polmeth-2019"
+in_dir  <- file.path(dropbox_path, "input-data")
+out_dir <- file.path(dropbox_path, "output")
+user_dump     <- "dump_user_forecasts_09_07_2018_09_15_12.csv"
+question_dump <- "dump_questions_09_07_2018_09_30_55.csv" 
+
+
+mbrier <- function(f, o, ordered = TRUE) {
+  prediction = f
+  outcome = o
+  
+  stopifnot(length(prediction)==length(outcome))
+  stopifnot(all(outcome %in% c(0L, 1L, NA)))
+  stopifnot(all.equal(sum(prediction), sum(outcome), 1L) | any(is.na(prediction)))
+  
+  if (ordered) {
+    pairs <- NULL
+    for (i in 1:(length(prediction)-1)) {
+      p_i <- sapply(split(prediction, 1:length(prediction) > i), sum)
+      o_i <- sapply(split(outcome, 1:length(outcome) > i), max)
+      pairs <- c(pairs, sum((p_i - o_i)^2))
+    }
+    brier <- mean(pairs)
+  } else {
+    brier <- sum((prediction - outcome)^2)
+  }
+  
+  return(brier)
+}
+
+
+q_dump <- file.path(dropbox_path, "input-data", question_dump) %>%
+  read_csv(.,
+           col_types = cols(
+             .default = col_character(),
+             ifp_id = col_integer(),
+             discover_id = col_integer(),
+             start_date = col_datetime(format = ""),
+             end_date = col_datetime(format = ""),
+             resolved_date = col_datetime(format = ""),
+             voided_date = col_datetime(format = ""),
+             num_options = col_integer(),
+             is_ordinal = col_logical(),
+             is_resolved = col_logical(),
+             is_voided = col_logical(),
+             has_chart = col_logical()
+           )) %>%
+  mutate_at(vars(contains("date")), as.Date) %>%
+  mutate_at(vars(starts_with("has"), starts_with("is")), ~.x=="True")
+
+# Questions that require data aggregation
+agg_q <- q_dump %>%
+  group_by(ifp_id, title) %>%
+  summarize() %>%
[…]

### 2018-11-12 R/make-data.R
Update question dump to add more questions with resolution, close #2

@@ -9,7 +9,7 @@ dropbox_path  <- "~/Dropbox/Work/asia-polmeth-2019"
 in_dir  <- file.path(dropbox_path, "input-data")
 out_dir <- file.path(dropbox_path, "output")
 user_dump     <- "dump_user_forecasts_09_07_2018_09_15_12.csv"
-question_dump <- "dump_questions_09_07_2018_09_30_55.csv" 
+question_dump <- "dump_questions_11_12_2018_08_51_25.csv" 
 
 
 mbrier <- function(f, o, ordered = TRUE) {
@@ -35,13 +35,25 @@ mbrier <- function(f, o, ordered = TRUE) {
   return(brier)
 }
 
+# The latest question dump don't have IFP_id's anymore, need to back-engineer 
+# this based on title
+ifp_id_mapping <- file.path(in_dir, "dump_questions_09_07_2018_09_30_55.csv") %>%
+  read_csv(., col_types = cols(
+    .default = col_character(),
+    ifp_id = col_integer(),
+    discover_id = col_integer(),
+    start_date = col_datetime(format = ""),
+    end_date = col_datetime(format = ""),
+    resolved_date = col_datetime(format = ""),
+    voided_date = col_datetime(format = ""),
+    num_options = col_integer()
+  )) %>%
+  select(ifp_id, title)
 
-q_dump <- file.path(dropbox_path, "input-data", question_dump) %>%
+q_dump <- file.path(in_dir, question_dump) %>%
   read_csv(.,
            col_types = cols(
              .default = col_character(),
-             ifp_id = col_integer(),
-             discover_id = col_integer(),
              start_date = col_datetime(format = ""),
              end_date = col_datetime(format = ""),
              resolved_date = col_datetime(format = ""),
@@ -49,11 +61,20 @@ q_dump <- file.path(dropbox_path, "input-data", question_dump) %>%
              num_options = col_integer(),
              is_ordinal = col_logical(),
              is_resolved = col_logical(),
-             is_voided = col_logical(),
-             has_chart = col_logical()
+             is_voided = col_logical()
            )) %>%
   mutate_at(vars(contains("date")), as.Date) %>%
-  mutate_at(vars(starts_with("has"), starts_with("is")), ~.x=="True")
+  mutate_at(vars(starts_with("has"), starts_with("is")), ~.x=="True") %>%
+  # add in IFP ID
[…]

### 2018-11-13 paper/paper.Rmd
work on paper

@@
 ---
-title: "Assessing the quality of Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
+title: "Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
 author: "Andreas Beger and Michael D. Ward"
@@
 output: 
-  pdf_document
-abstract: "\\noindent HFC \\newline \\newline \\newline
-\\textbf{Funding acknowledgement}: This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
+  pdf_document:
+    latex_engine: xelatex
+abstract: "\\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project to develop hybrid geopolitical forecasting systems that combine human- and machine-generated forecasts.
+The first project trial period took place in 2018, during the course of which volunteer participants, Amazon Mechanical Turkers, and an automated time series forecasting module we developed provided forecasts for more than 150 questions covering a diverse set of topics.
+We investigate two questions: (1) how does the performance of turker forecasters compare to volunteers, and (2) what impact did access to the machine forecasts have on forecaster accuracy?
+\\newline \\newline
+"
+bibliography: "master.bib"
+header-includes:
+  - 
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+
+library("tidyverse")
+```
+
+```{r load-data}
+dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
+in_dir  <- file.path(dropbox_path, "input-data")
+out_dir <- file.path(dropbox_path, "output")
+
+fcasts <- read_rds(file.path(out_dir, "user_forecasts_w_features.rds")) %>%
+  mutate(date = as.Date(date),
+         x = as.numeric(date) - min(as.numeric(date)),
+         sees_chart = sees_chart==1,
+         sees_model = sees_model==1) 
+```
+
+The Hybrid Forecasting Competition (HFC)^[**Funding acknowledgement**: This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
 The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S.
@@
 The U.S.
-Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein."
----
+Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.] is an IARPA program that seeks to develop methods for hybrid geopolitical forecasting system that combine human and machine forecasts to answer a broad range of questions about economic, political, health, and other events and trends.
+The first trial period, or RCT, took place in 2018, during the course of which hundreds of volunteer and Amazon Mechnial Turk forecasters, as well as automated machine models, answered more than 150 questions covering a broad range of issues.
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+We worked on one of the competition teams, and specifically by contributing a time series forecasting module.
+Out of the large set of interesting questions one could examine with the results so far, given our specific focus on this project, we will try to examine in this paper two questions: (1) how does the accuracy of turker forecasters compare to volunteer forecasters, and (2) what was the impact of the machine forecasts on human forecaster accuracy?
 
@@
 
-Background on HFC
+The goal of the HFC is to find ways to optimally combine human and machine forecasts.
+For example, machine forecasts can be reliable and scalable, but are constrained by available data, idiosyncratic questions, and cold start problems when a corpus of historical data is not available.
+Human-generated forecasts on the other hand are more flexible, but also more costly to scale and subject to various cognitive biases.[e.g.
+@kahneman:2011].
+
+The competition is organized around a series of official trial periods ("RCTs") during which forecasters can actively forecast on official competition questions, and the results of which are used for subsequent project evaluations.
+Time between the trial periods is reserved for development and improvements, as well as any internal, non-official experiments performers might choose to run.
+
+During the trial periods, performer teams are competing against a reference forecast, with the ultimate goal being to find methods or tools that provide superior forecasts.
+
+Each entry, at the core, consists of a Good Judgement-style platform that can receive competition questions (IFPs in project parlance) and show them to users, who then provide probabilistic forecasts for a set of 2-5 answer options.
+
+The original goal was for the project to recruit a sufficient number of volunteer forecasters to provide each performing team with the numbers needed to forecast at the rate at which questions are generated during the trial periods.
+When it became apparent, several weeks into the first trial period, that the level of forecasting activity was not going to be sufficient for adequate evaluations, a contingency plan was activated whereby additional forecasters (turkers) were provided with Amazon Mechnanical Turk.
 
@@
 
-Cover the basics: there is a platform that can show a chart and machine forecast; how are the machine forecasts generated.
+One of the features of our project's platform is functionality that for questions that have a clearly associated data stream, e.g.
+questions about interest rates or gold prices, acquires the relevant data source and displays it to users in a chart.
+
+This same data platform also feeds a module, "basil-ts", that uses a simple time series model to derive probabilties for the question answer options.
+This was one of our main machine-forecast providers during the first trial period.
+
+To ensure scalability, both the data platform and forecasting module are required, by competition rules, to be largely automated.
+
+Most of the difficulty in implementing basil-ts was related to the automation requirements, not the actual modeling.
+E.g.
+it has to be able to identify the temporal resolution in both the question and the input time series, e.g.
+is it daily or monthly, how far ahead to forecast, whether there are any constraints on the data and thus forecast, e.g.
+is it a count series that cannot take on negative values.
 
@@
 
-Go over the conditions, and the complications that happened: turkers only in two conditions, and the data and forecaster code changes over time.
+The original research design split forecasters into three groups.
+The first, control, group only had access to a basic version of the platform that displays the question and answer options, but without any additional information.
+The second group, B, could also see a chart displaying historic data values relevant to the question at hand.
+The third group, C, in addition to the chart could also see the machine forecast.
+All forecasters, regardless of group, were forecasting on the same set of questions.
 
-Two key questions we want to examine:
+| Condition | Treatment                 | Population             | 
+|-----------|---------------------------|------------------------|
+|A          |None; control              | Volunteers and Turkers |
+|B          |Chart                      | Volunteeers            |
+|C          |Chart and machine forecast | Volunteers and Turkers |
 
-1. How accurate were Turker forecasts compared to volunteer forecasts?
-2. What was the impact of the machine forecasts on human accuracy?
+Table: Summary of original research design and eventual forecaster populations.
+
+Due to lower than expected activity levels, turker forecasters started to be provided several weeks intot he first trial period.
+Given activity levels at that time, a decision was made to assign turkers to either condition A or C, but not B.
+This mean that the group whose treatment was to only see charts (B) consisted only of volunteers, while the other two groups contained mixed populations of volunteers and turkers, thus adding a confounding factor for assessing the impact of charts and machine models.
+
+Additionally, chartable data was only available for about a third of the questions.
+The chart and model treatments thus are really part of a nested three layer set of factors: whether a question had chartable data, then whether a chart was displayed to a user, then whether a machine forecast was also displayed to a user.
+Whether a question had associated data depended very much on the type of question, ranging from questions like "What will be the gold price on [date]" where there is a clearly relevant time series that is easy to acquire and update, to "How many battle deaths will ACLED report in [country] in [date]?" where data are available but have to be transformed to a relevant series, to, on the other end, questions like "Who will win the World Cup?" where it is exceedingly difficult to acquire relevant time series data.
+Thus it is also possible that questions with chartable and without chartable data were systemtically different in their difficutly and average acccuracy levels.
+
+The effects of the two randomly assigned treatments, seeing a chart and seeing a model, are thus confounded by two other non-ranom factors: whether a forecaster was a volunteer or turker, and whether a question had chartable data or not.
+This effectively gives us an experiment design like in Table XXXX.
+
+| Condition | Is Turker? | Chartable data | Sees chart? | Sees models? |
+|-----------|------------|----------------|-------------|--------------|
+| A.1 |   |
+| A.2 | X |
+| A.3 |   | X |
+| A.4 | X | X |
+| B.1 |   |   |
+| B.2 |   | X | X |
+| C.1 |   |   |
+| C.2 | X |   |
+| C.3 |   | X | X | X |
+| C.4 | X | X | X | X |
+
+Table: Treatment by group after addition of turkers to conditions A and C.
+
+To complicate matters further, the quality of the data acquisition and machine forecasting modules changed over time during the RCT.
+Once a broader set of questions started flowing into the platform when the RCT started, a range of issues affecting both data quality and machine forecast quality arose.
+Examples include:
+
+- Incorrect aggregation of event data like ACLED and ICEWS that produced incorrect monthly counts of events or casualties.
+- Several questions were about monthly oil production rates for various countries.
+The resolution data are in PDF tables in monthly OPEC oil market reports.
+As this is hard to automatically scrape, an alternative source was used, but the values in this source do not always match the OPEC values.
+Nor, for that matter, do OPEC values always match in successive monthly reports.
+- Parsing failures that led the forecaster to produce non-sensical forecasts like negative values for count time series, or complete failures, e.g.
+when one question had a date like "April (04) 2018" instead of the expected "April 2018".
+- A limitation in the way Elastic Search aggregates non-standard fixed length time periods, e.g.
+periods of 40 days, required that for some questions data aggregation had to occur in the forecasting module, as opposed to data platform module, unlike planned.
+
+Most of these issues and bugs were fixed during the course of the RCT, and thus the quality of both the data and machine forecasts should have improved over time.
+
+Given that we (Predictive Heuristics) implemented the time series forecasting module, we are primarily interested in the impact that seeing machine forecasts had on the quality of human forecasts.
+We know from early results that the machine forecasts were generally not as accurate as the human forecasts.
 
@@
 
-Key leverage point is that only some questions had a machine forecast.
+## DGP
 
-Go over what the data for this are, how accuracy was assessed
+Question comes from T&E to our platform.
+
+Now on two tracks:
+
+1. Human
+  - sees question, and depending on condition the chart and machine forecast
+  - forecasts, possibly multiple times per day
+  - Mean daily Brier score
+
+2. Platform
+  - parses question, finds associated data
+  - sends question and data to forecaster
+  - forecaster parses question and data, TS forecast, convert to MN forecast
+  - mean daily Brier score and/or shown to user
+
+## Data
+
+The first trial period lasted from March 7th to September 7th 2018.
+The data used in the analysis below consist of `r nrow(fcasts) %>% format(big.mark = ",")` forecasts made by `r length(unique(fcasts$user_id)) %>% format(big.mark = ",")` distinct users on `r length(unique(fcasts$ifp_id))` questions during this period.
+The first turker forecasts were on May 2nd.
+
+![Average mean daily Brier score (MDB) by forecaster type over time. Both volunteers and turkers improved over time.](../output/figures/mdb-turker-over-time.png)
 
@@
 
+```{r tbl-accuracy-by-condition}
+read_csv("../output/tables/brier-by-conditions.csv") %>%
+  mutate(Condition = paste0(Condition, ".", c(1:4, 1:2, 1:4))) %>%
+  knitr::kable(digits = 2)
+```
+
+Table: Mean Brier scores by condition
+
 ## Turker performance compared to volunteer forecasters
@@
 ## Machine forecast impact
+
+![Average mean daily Brier score (MDB) by forecaster type over time.
+Both volunteers and turkers improved over time.\label{fig:volunteers-only}](../output/figures/volunteers-only-accuracy-by-chartability-over-time.png)
+
+Figure \ref{fig:volunteers-only} looks only at volunteer forecasters in conditions B and C, showing again mean daily Brier scores over time.
+The right chart shows that forecasters who saw both a chart and model had slighly worse performance throughout the RCT compared to condition B forecasters who only saw the chart.
+This matches the hypothesis that seeing the machine forecasts hurt performance.
+However, if we look at the chart on the left, which shows performance for questions for which no data was available, i.e.
+where both condition B and C forecasters saw neither a chart nor a model, condition C performance was also worse up until sometime in late July.
 
@@
 some data transformation are needed) did not boost accuracy more than in other series.
+
+# References

### 2018-11-13 R/analysis.R

@@ -1,8 +1,336 @@
 library("tidyverse")
+library("lme4")
+library("mgcv")
+library("hrbrthemes")
+library("MuMIn")
+library("betareg")
+library("rstan")
+library("rstanarm")
+library("forcats")
 
 dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
+in_dir  <- file.path(dropbox_path, "input-data")
+out_dir <- file.path(dropbox_path, "output")
 
 
+fcasts <- read_rds(file.path(out_dir, "user_forecasts_w_features.rds")) %>%
+  mutate(date = as.Date(date),
+         x = as.numeric(date) - min(as.numeric(date)),
+         sees_chart = sees_chart==1,
+         sees_model = sees_model==1) 
+
+#fcasts$sees_model <- factor(fcasts$sees_model, labels = c("No", "Yes"))
+#fcasts$sees_chart <- factor(fcasts$sees_chart, labels = c("No", "Yes"))
+
+fcasts %>%
+  group_by(condition) %>% 
+  summarize(mean_brier = mean(brier))
+
+fcasts %>%
+  group_by(turker, sees_chart, sees_model) %>%
+  summarize(mean_brier = mean(brier))
+
+cond_tbl <- fcasts %>%
+  group_by(condition, turker, sees_chart, sees_model, has_arima) %>%
+  summarize(mean_brier = mean(brier),
+            n_ifps = length(unique(ifp_id)),
+            n_forecasts = n()) %>%
+  ungroup() %>%
+  select(condition, turker, has_arima, sees_chart, sees_model, everything()) %>%
+  arrange(condition, turker, has_arima, sees_chart, mean_brier)
+
+cond_tbl %>%
+  arrange(condition, has_arima) %>%
+  mutate(condition = fct_recode(condition, A = "a", B = "b", C = "c"),
+         turker = fct_recode(as.character(turker), Yes = "TRUE", No = "FALSE"),
+         has_arima = fct_recode(as.character(has_arima), Yes = "TRUE", No = "FALSE")
+         ) %>%
+  rename(Condition = condition, Turker = turker, `Has data` = has_arima, 
+         `Saw chart` = sees_chart, `Saw model` = sees_model, 
+         `Mean Brier` = mean_brier, Questions = n_ifps, Forecasts = n_forecasts) %>%
+  write_csv("output/tables/brier-by-conditions.csv")
+
+
[…]

### 2018-11-13 R/debug-data-read.R

@@ -7,8 +7,6 @@
 library("tidyverse")
 library("future.apply")
 
-plan(multiprocess, workers = 3L)
-
 dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
 out_dir <- file.path(dropbox_path, "output")
 
@@ -16,18 +14,98 @@ user_str <- "dump_user_forecasts_09_07_2018_09_15_12.csv" %>%
   file.path(dropbox_path, "input-data", .) %>%
   read_lines() 
 
-str_header <- user_str[1]
-str_row    <- as.list(tail(user_str, -1))
-
-# this takes a while
-parsed_rows <- future_lapply(str_row, function(x) {
-  suppressWarnings(read_csv(paste0(c(str_header, x), collapse = "\n")))
-})
-
-write_rds(parsed_rows, file.path(out_dir, "debug_user_dump_read.rds"))
-
-
 user_raw <- "dump_user_forecasts_09_07_2018_09_15_12.csv" %>%
   file.path(dropbox_path, "input-data", .) %>%
   read_csv(., quote = "\"") 
 
+# Parsing each row as a CSV, separately; this takes a while
+if (FALSE) {
+  plan(multiprocess, workers = 3L)
+  
+  str_header <- user_str[1]
+  str_row    <- as.list(tail(user_str, -1))
+  
+  # this takes a while
+  parsed_rows <- future_lapply(str_row, function(x) {
+    suppressWarnings(read_csv(paste0(c(str_header, x), collapse = "\n"), 
+                              quote = "\"",
+                              col_types = cols(
+                                .default    = col_character(),
+                                date        = col_datetime(),
+                                user_id     = col_integer(),
+                                ifp_id      = col_integer(),
+                                discover_id = col_integer(),
+                                num_options = col_integer(),
+                                option_1 = col_double(),
+                                option_2 = col_double(),
+                                option_3 = col_double(),
+                                option_4 = col_double(),
+                                option_5 = col_double()
+                              )))
+  })
+  
+  write_rds(parsed_rows, file.path(out_dir, "debug_user_dump_read.rds"))
+} else {
+  parsed_rows <- read_rds(file.path(out_dir, "debug_user_dump_read.rds"))
+}
+
+row_N <- sapply(parsed_rows, nrow)
+table(row_N)
+
[…]

### 2018-11-13 R/make-data.R

@@ -4,6 +4,7 @@
 
 library("tidyverse")
 library("futile.logger")
+library("future.apply")
 
 dropbox_path  <- "~/Dropbox/Work/asia-polmeth-2019"
 in_dir  <- file.path(dropbox_path, "input-data")
@@ -74,7 +75,9 @@ no_q  <- filter(q_dump, !is.na(ifp_id) & is.na(question_id))
 no_id <- filter(q_dump, is.na(ifp_id))
 
 q_dump <- q_dump %>%
-  filter(!is.na(ifp_id))
+  filter(!is.na(ifp_id)) %>%
+  # filter practice questions
+  filter(!str_detect(title, "Practice Q"))
 
 # Questions that require data aggregation
 agg_q <- q_dump %>%
@@ -95,7 +98,7 @@ answers <- q_dump %>%
   select(ifp_id, has_historic_data, has_arima, correct_option)
   
 
-
+# The warnings are related to missing option_2, option_3, etc.
 user_raw <-  file.path(dropbox_path, "input-data", user_dump) %>%
   read_csv(.)
 

### 2018-11-30 paper/paper.Rmd
redo data and add in machine forecasts

@@
 "
-bibliography: "master.bib"
+bibliography: "../../whistle/master.bib"
 header-includes:
@@
 
-The competition is organized around a series of official trial periods ("RCTs") during which forecasters can actively forecast on official competition questions, and the results of which are used for subsequent project evaluations.
-Time between the trial periods is reserved for development and improvements, as well as any internal, non-official experiments performers might choose to run.
+The competition is organized around Good Judgement-style forecasting tournaments [@tetlock:2005, @tetlock:etal:2017], where forecasters answer and are scored on a diverse pool of questions.
+Examples from the first trial period, RCT-A, include:
 
-During the trial periods, performer teams are competing against a reference forecast, with the ultimate goal being to find methods or tools that provide superior forecasts.
+- What will be the long-term interest rate for South Africa (ZAF) in July 2018?
+- How many deaths perpetrated by Boko Haram will the Council on Foreign Relations report for June 2018?
+- What will be the daily closing spot price of Brent crude oil (USD per barrel) on 31 May 2018, according to the U.S. EIA?
 
-Each entry, at the core, consists of a Good Judgement-style platform that can receive competition questions (IFPs in project parlance) and show them to users, who then provide probabilistic forecasts for a set of 2-5 answer options.
+Each question includes 2 to 5 answer options to which forecasters must assign weights summing to 1. Performance for a single forecast is then based on an ordered multinomial Brier score.
 
-The original goal was for the project to recruit a sufficient number of volunteer forecasters to provide each performing team with the numbers needed to forecast at the rate at which questions are generated during the trial periods.
-When it became apparent, several weeks into the first trial period, that the level of forecasting activity was not going to be sufficient for adequate evaluations, a contingency plan was activated whereby additional forecasters (turkers) were provided with Amazon Mechnanical Turk.
+In addition to this basic "human forecaster" tournament, competitors were expected to implement features that would in some fashion or agument these human forecasts with machine-generated tools and forecasts.
+An explicity requirement for the latter is that they are automated systems, e.g.
+an ad-hoc hand-tuned and expert-implemented machine model to forecast on a question would not be allowed, rather it would have to be a system that can generate such a model.
+On the other hand, if a forecaster has the skills and inclination to use data and model as a forecasting tool, that is perfectly allowed.
 
@@
 
-One of the features of our project's platform is functionality that for questions that have a clearly associated data stream, e.g.
-questions about interest rates or gold prices, acquires the relevant data source and displays it to users in a chart.
+One of the distinguishing features of our team's approach was a system that could automatically associate some of the RCT-A questions with a clearly corresponding time series and display these to a forcaster.
+This is based on a data platform which automatically collects and updates data from a variety of sources, and which can associate questions, based on their title, to an appropriate transformation of a data set if it is in the platform and matches a known, pre-specified pattern or template.
+If data is found for a question, it can then be shown to users as a simple time series chart accompanying the relevant question.
 
-This same data platform also feeds a module, "basil-ts", that uses a simple time series model to derive probabilties for the question answer options.
-This was one of our main machine-forecast providers during the first trial period.
-
-To ensure scalability, both the data platform and forecasting module are required, by competition rules, to be largely automated.
-
-Most of the difficulty in implementing basil-ts was related to the automation requirements, not the actual modeling.
-E.g.
-it has to be able to identify the temporal resolution in both the question and the input time series, e.g.
-is it daily or monthly, how far ahead to forecast, whether there are any constraints on the data and thus forecast, e.g.
-is it a count series that cannot take on negative values.
+Additionally, we could use the time series to generate a machine forecast based on a univariate ARIMA model, and which then would either be shown to a user, and/or submitted seperately as a standalone forecast.
+The module generating these forecasts, "basil-ts", was based on the Auto ARIMA model in @hyndman:khandakar:2008, which consists of a ARIMA-family model with an automated algorithm for determining a reasonble specific model structure, e.g.
+whether and how many differencing orders to apply, AR and MA orders, and some other parameters.
+This is wrapped in additional functionality needed to meet the automation requirements, e.g.
+recognizing how far a forecast needs to extend, data pre-processing, converting time series to answer option forecasts, updating forecasts with additional information if available, etc.
 
@@
 
-The original research design split forecasters into three groups.
-The first, control, group only had access to a basic version of the platform that displays the question and answer options, but without any additional information.
-The second group, B, could also see a chart displaying historic data values relevant to the question at hand.
-The third group, C, in addition to the chart could also see the machine forecast.
+Since the scientific goal of HFC is to evaluate various hybrid forecasting techniques, the pripmary aspect of our research design for RCT-A was designed to assess what impact exposure to the time series charts and model forecasts would have on human forecaster accuracy.
+Incoming forecasters were assigned to one of three experimental conditions.
+The first group, A, served as control group and only had access to a basic version of the online platform showin the question information and tools to enter weights for the answer options.
+The second group, B, could also see the time series charts, and the third group, C, could see the chart and machine forecast.
 All forecasters, regardless of group, were forecasting on the same set of questions.
 
-| Condition | Treatment                 | Population             | 
-|-----------|---------------------------|------------------------|
-|A          |None; control              | Volunteers and Turkers |
-|B          |Chart                      | Volunteeers            |
-|C          |Chart and machine forecast | Volunteers and Turkers |
+There were elements in the research design to assess other design choices but they are not relevant to the set of questions we seek to examine here, thus we will not discuss them.
 
-Table: Summary of original research design and eventual forecaster populations.
+| Condition | Treatment                 | 
+|-----------|---------------------------|
+|A          |None; control              |
+|B          |Chart                      |
+|C          |Chart and machine forecast |
+
+Table: Summary of original research design.
+
+### Turker
 
@@
 This mean that the group whose treatment was to only see charts (B) consisted only of volunteers, while the other two groups contained mixed populations of volunteers and turkers, thus adding a confounding factor for assessing the impact of charts and machine models.
+
+### Time series data was not available for all questions
 
@@
 
-Table: Treatment by group after addition of turkers to conditions A and C.
+Table: Effective treatments by group after addition of turkers to conditions A and C.
+
+### Varying quality of data and machine forecasts over time
 
@@
 Most of these issues and bugs were fixed during the course of the RCT, and thus the quality of both the data and machine forecasts should have improved over time.
-
-Given that we (Predictive Heuristics) implemented the time series forecasting module, we are primarily interested in the impact that seeing machine forecasts had on the quality of human forecasts.
-We know from early results that the machine forecasts were generally not as accurate as the human forecasts.
 
@@
 
-The first trial period lasted from March 7th to September 7th 2018.
-The data used in the analysis below consist of `r nrow(fcasts) %>% format(big.mark = ",")` forecasts made by `r length(unique(fcasts$user_id)) %>% format(big.mark = ",")` distinct users on `r length(unique(fcasts$ifp_id))` questions during this period.
-The first turker forecasts were on May 2nd.
-
-![Average mean daily Brier score (MDB) by forecaster type over time. Both volunteers and turkers improved over time.](../output/figures/mdb-turker-over-time.png)
+RCT-A lasted from March 7th to September 7th 2018.
+However, Turkers did not enter until May 2nd, and for an unrelated reason we also discard data after August 2nd.
+This leaves a total of 48,931 forecasts---7,816 from volunteer forecasters, 39,140 from Turkers, and 1,975 from machine---for 156 IFPs, of which 101 did not have TS data, 6 had TS data but not a machine forecast, and 49 had both TS data and a machine forecast.
 
@@
 
+### Brier over time
+
+Make point here that Brier generaly decreased over time? Maybe questions easier, maybe forecasters learn over time? Just establish that all analysis from here is for May 2nd and on only.
+
+### Other
+
 Machine forecasts were not more accurate than human forecasts.
+
+|Forecaster |Question         | A: no chart| B: chart only| C: chart and model|
+|:----------|:----------------|-----------:|-------------:|------------------:|
+|Volunteer  |IFP without data |        0.15|          0.09|               0.12|
+|Volunteer  |IFP had data     |        0.25|          0.22|               0.28|
+|Turker     |IFP without data |        0.15|            NA|               0.14|
+|Turker     |IFP had data     |        0.28|            NA|               0.27|
+
+Table: MMDB by forecaster, data availability, and condition.
+
+The low MMDB for condition B forecasters appears to be primarily driven by good performance on questions that *did not* have charts (keeping in mind that 2/3 of questions would not have had a chart).
 

### 2018-11-30 R/analysis.R

@@ -13,39 +13,346 @@ in_dir  <- file.path(dropbox_path, "input-data")
 out_dir <- file.path(dropbox_path, "output")
 
 
-fcasts <- read_rds(file.path(out_dir, "user_forecasts_w_features.rds")) %>%
+user_fcasts <- read_rds(file.path(out_dir, "user_forecasts.rds")) %>%
   mutate(date = as.Date(date),
+         # x is just a date index for plotting, i think
          x = as.numeric(date) - min(as.numeric(date)),
          sees_chart = sees_chart==1,
          sees_model = sees_model==1) 
 
+machine_fcasts <- read_rds(file.path(out_dir, "machine_forecasts.rds"))
+
+data_sources <- read_csv(file.path(in_dir, "questions-with-answers.csv")) 
+
 #fcasts$sees_model <- factor(fcasts$sees_model, labels = c("No", "Yes"))
 #fcasts$sees_chart <- factor(fcasts$sees_chart, labels = c("No", "Yes"))
 
+df1 <- machine_fcasts %>%
+  select(ifp_id, date, has_historic_data, has_arima, needs_data_agg,
+         num_options, p_correct, brier) %>%
+  mutate(Forecaster = "Machine", condition = "n/a", sub_condition_1 = "n/a",
+         sees_chart = NA, sees_model = NA)
+df2 <- user_fcasts %>%
+  mutate(Forecaster = ifelse(turker, "Turker", "Volunteer")) %>%
+  select(ifp_id, date, has_historic_data, has_arima, needs_data_agg,
+         num_options, p_correct, brier, Forecaster, condition, sub_condition_1,
+         sees_chart, sees_model)
+all_fcasts <- bind_rows(df1, df2) %>%
+  left_join(data_sources %>% select(ifp_id, data_source), by = "ifp_id") %>%
+  replace_na(list(data_source = "other")) %>%
+  # add in summary info for chart/arima status
+  mutate(Data_status = case_when(
+    has_historic_data & has_arima ~ "Chart and model",
+    has_historic_data & !has_arima ~ "Chart only",
+    !has_historic_data ~ "None",
+    TRUE ~ NA_character_
+  ))
+
+# Human and machine forecasts during common period
+common_period <- all_fcasts %>%
+  filter(date >= "2018-05-02" & date <= "2018-08-02")
+
+
+# Summary statistics ------------------------------------------------------
+
[…]

### 2018-11-30 R/make-data.R

@@ -1,181 +1,0 @@
-#
-#   Process the raw staff dumps to make the analysis data
-#
-
-library("tidyverse")
-library("futile.logger")
-library("future.apply")
-
-dropbox_path  <- "~/Dropbox/Work/asia-polmeth-2019"
-in_dir  <- file.path(dropbox_path, "input-data")
-out_dir <- file.path(dropbox_path, "output")
-user_dump     <- "dump_user_forecasts_09_07_2018_09_15_12.csv"
-question_dump <- "dump_questions_11_12_2018_08_51_25.csv" 
-
-
-mbrier <- function(f, o, ordered = TRUE) {
-  prediction = f
-  outcome = o
-  
-  stopifnot(length(prediction)==length(outcome))
-  stopifnot(all(outcome %in% c(0L, 1L, NA)))
-  stopifnot(all.equal(sum(prediction), sum(outcome), 1L) | any(is.na(prediction)))
-  
-  if (ordered) {
-    pairs <- NULL
-    for (i in 1:(length(prediction)-1)) {
-      p_i <- sapply(split(prediction, 1:length(prediction) > i), sum)
-      o_i <- sapply(split(outcome, 1:length(outcome) > i), max)
-      pairs <- c(pairs, sum((p_i - o_i)^2))
-    }
-    brier <- mean(pairs)
-  } else {
-    brier <- sum((prediction - outcome)^2)
-  }
-  
-  return(brier)
-}
-
-# The latest question dump don't have IFP_id's anymore, need to back-engineer 
-# this based on title
-ifp_id_mapping <- file.path(in_dir, "dump_questions_09_07_2018_09_30_55.csv") %>%
-  read_csv(., col_types = cols(
-    .default = col_character(),
-    ifp_id = col_integer(),
-    discover_id = col_integer(),
-    start_date = col_datetime(format = ""),
-    end_date = col_datetime(format = ""),
-    resolved_date = col_datetime(format = ""),
-    voided_date = col_datetime(format = ""),
-    num_options = col_integer()
-  )) %>%
-  select(ifp_id, title)
-
-q_dump <- file.path(in_dir, question_dump) %>%
-  read_csv(.,
-           col_types = cols(
-             .default = col_character(),
-             start_date = col_datetime(format = ""),
-             end_date = col_datetime(format = ""),
-             resolved_date = col_datetime(format = ""),
-             voided_date = col_datetime(format = ""),
[…]

### 2018-11-30 R/make-human-forecast-data.R

@@ -1,0 +1,100 @@
+#
+#   Process the raw staff dumps to make the analysis data
+#
+
+library("tidyverse")
+library("futile.logger")
+library("future.apply")
+
+dropbox_path  <- "~/Dropbox/Work/asia-polmeth-2019"
+in_dir  <- file.path(dropbox_path, "input-data")
+out_dir <- file.path(dropbox_path, "output")
+user_dump     <- "dump_user_forecasts_09_07_2018_09_15_12.csv"
+answer_key <- "answer_key.rds"
+
+
+source("R/mbrier.R")
+
+answers <- read_rds(file.path(out_dir, answer_key))
+
+# The warnings are related to missing option_2, option_3, etc.
+user_raw <-  file.path(dropbox_path, "input-data", user_dump) %>%
+  read_csv(.)
+
+user_clean <- user_raw %>%
+  # turker is string true/false, make logical
+  mutate(turker = case_when(
+    turker=="False" ~ FALSE,
+    turker=="True"  ~ TRUE,
+    TRUE ~ NA
+  )) %>%
+  # convert options from percent to probability
+  mutate_at(vars(starts_with("option_")), ~ ./100) %>%
+  # binary questions show as having only 1 option, make it 2 for "yes"/"no"
+  # and fill in 1-p for 2nd option
+  mutate(
+    num_options = case_when(
+      num_options==1 ~ 2L,
+      TRUE ~ num_options
+    ),
+    option_2 = case_when(
+      is.na(option_2) ~ 1 - option_1,
+      TRUE ~ option_2
+    )) 
+
+# Check that options sum to 1
+p_range <- user_clean %>% 
+  gather(option, p, starts_with("option_")) %>%
+  group_by(date, user_id, ifp_id) %>% 
+  summarize(p = sum(p, na.rm = TRUE)) %>%
+  pull(p) %>%
+  range(.)
+if (p_range[1]<0.999999 | p_range[2]>1.000001) {
+  stop("p is wrong")
+}
+
+
+
+# Add in answers
+user_w_answer <- left_join(user_clean, answers, by = "ifp_id") %>%
+  # 801, some weird test IFP or something
+  filter(!is.na(correct_option)) %>%
+  arrange(ifp_id, user_id, date) %>%
+  rename(needs_data_agg = needs_agg) %>%
+  select(ifp_id, user_id, date, everything(), 
+         # don't need these
+         -discover_id, -rationale)
[…]

### 2018-11-30 R/make-machine-forecast-data.R

@@ -1,0 +1,153 @@
+
+dropbox_path  <- "~/Dropbox/Work/asia-polmeth-2019"
+in_dir  <- file.path(dropbox_path, "input-data")
+out_dir <- file.path(dropbox_path, "output")
+
+
+answers <- read_rds(file.path(out_dir, "answer_key.rds")) %>%
+  rename(needs_data_agg = needs_agg)
+
+# Load fcast dump file and add resolutions
+# Use machine forecasts in "aggreagted_..." file instead of "dump_machine..."
+# because the latter is apparently inaccurate
+# However, this only covers back to April 13th, so take fcast_dump... for rest
+agg_dump_file <- dir(in_dir,
+                     pattern = "dump\\_aggregated",
+                     full.names = TRUE) %>%
+  tail(1)
+agg_dump <- read_csv(agg_dump_file,
+                     col_types = cols(
+                       date = col_datetime(format = ""),
+                       slot_id = col_character(),
+                       method_name = col_character(),
+                       method_type = col_character(),
+                       ifp_id = col_integer(),
+                       discover_id = col_integer(),
+                       num_options = col_integer(),
+                       option_1 = col_double(),
+                       option_2 = col_double(),
+                       option_3 = col_double(),
+                       option_4 = col_double(),
+                       option_5 = col_double()
+                     ))
+agg_dump <- agg_dump %>%
+  filter(method_name=="Machine_only_models") %>%
+  mutate(date = as.Date(date)) %>%
+  left_join(., answers %>% select(ifp_id, has_historic_data), by = "ifp_id") %>%
+  # if no historic_data, take it out
+  filter(has_historic_data %in% TRUE) %>%
+  # some binary questions only have 1 option, fix this
+  mutate(
+    num_options = case_when(
+      is.na(option_2) ~ 2L,
+      TRUE ~ num_options
+    ),
+    option_2 = case_when(
+      is.na(option_2) ~ 1 - option_1,
+      TRUE ~ option_2
+    ),
+    has_historic_data = NULL) %>%
+  # if multiple forecasts for same IFP, take last one only
+  group_by(ifp_id, date) %>%
+  filter(1:n()==n())
+
+min(agg_dump$date)
+
+# Use other dump file to fill in forecasts before April 13th
+fcast_dump_file <- dir(file.path(in_dir),
[…]

### 2018-11-30 R/make-question-key.R

@@ -1,0 +1,53 @@
+
+
+dropbox_path  <- "~/Dropbox/Work/asia-polmeth-2019"
+in_dir  <- file.path(dropbox_path, "input-data")
+out_dir <- file.path(dropbox_path, "output")
+question_dump <- "dump_questions_11_29_2018_20_17_01.csv"
+
+
+q_dump <- read_csv(
+  file.path(in_dir, question_dump),
+  col_types = cols(
+    .default = col_character(),
+    ifp_id = col_integer(),
+    discover_id = col_integer(),
+    start_date = col_datetime(format = ""),
+    end_date = col_datetime(format = ""),
+    resolved_date = col_datetime(format = ""),
+    voided_date = col_datetime(format = ""),
+    num_options = col_integer()
+  )
+) %>%
+  mutate_at(vars(contains("date")), as.Date) %>%
+  mutate_at(vars(starts_with("has"), starts_with("is")), ~.x=="True")
+
+# only 2 practice questions
+no_id <- filter(q_dump, is.na(ifp_id))
+
+q_dump <- q_dump %>%
+  # filter practice questions
+  filter(!str_detect(title, "Practice Q"))
+
+# Questions that require data aggregation
+agg_q <- q_dump %>%
+  group_by(ifp_id, title) %>%
+  summarize() %>%
+  filter(str_detect(title, "ACLED|ICEWS|earthquakes|Boko|sea ice")) %>%
+  pull(ifp_id)
+
+# Get correct option and has_arima, has_chart for RCT-A (closed) questions
+answers <- q_dump %>%
+  # make sure only RCT-A IFPs
+  filter(end_date <= "2018-09-07") %>%
+  filter(!is.na(resolution)) %>%
+  # wide to long so we can code correct option_#
+  tidyr::gather(option, option_label, starts_with("option_")) %>%
+  dplyr::mutate(is_correct = option_label==resolution,
+                needs_agg = ifp_id %in% agg_q) %>%
+  # take out incorrect options now
+  filter(is_correct) %>%
+  rename(correct_option = option) %>%
+  select(ifp_id, has_historic_data, is_ordinal, has_arima, needs_agg, correct_option)
+
+write_rds(answers, file.path(out_dir, "answer_key.rds"))

### 2018-11-30 R/mbrier.R

@@ -1,0 +1,22 @@
+mbrier <- function(f, o, ordered = TRUE) {
+  prediction = f
+  outcome = o
+  
+  stopifnot(length(prediction)==length(outcome))
+  stopifnot(all(outcome %in% c(0L, 1L, NA)))
+  stopifnot(all.equal(sum(prediction), sum(outcome), 1L) | any(is.na(prediction)))
+  
+  if (ordered) {
+    pairs <- NULL
+    for (i in 1:(length(prediction)-1)) {
+      p_i <- sapply(split(prediction, 1:length(prediction) > i), sum)
+      o_i <- sapply(split(outcome, 1:length(outcome) > i), max)
+      pairs <- c(pairs, sum((p_i - o_i)^2))
+    }
+    brier <- mean(pairs)
+  } else {
+    brier <- sum((prediction - outcome)^2)
+  }
+  
+  return(brier)
+}

### 2018-11-30 paper/paper.Rmd
paper writing

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-
-library("tidyverse")
-```
-
-```{r load-data}
-dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
-in_dir  <- file.path(dropbox_path, "input-data")
-out_dir <- file.path(dropbox_path, "output")
-
-fcasts <- read_rds(file.path(out_dir, "user_forecasts_w_features.rds")) %>%
-  mutate(date = as.Date(date),
-         x = as.numeric(date) - min(as.numeric(date)),
-         sees_chart = sees_chart==1,
-         sees_model = sees_model==1) 
 ```
@@
 
-### Turker
+Comparison of groups A and B should have shown the effect of seeing a time series chart, and of groups B and C for the effect of seeing a model forecast.
+In practice, there were issues that complicated the effective design or forecaster groupings and treatments:
 
+- **Amazon Mechanical Turk forecasters**.
 Due to lower than expected activity levels, turker forecasters started to be provided several weeks intot he first trial period.
@@
 This mean that the group whose treatment was to only see charts (B) consisted only of volunteers, while the other two groups contained mixed populations of volunteers and turkers, thus adding a confounding factor for assessing the impact of charts and machine models.
+- **Gaps in data and machine coverage**.
+Only about 1/3 of IFPs had time series data available, meaning that within each condition group, most questions did not have a chart nor model.
+There was also a small number of instances where data was available, but a machine forecast was not.
+One example were questions related to FluNet influenza case counts, where a change in the data source broke the ability to update chart data, which would have required models to forecast over excessive time horizons.
+Since the availability of data was related to the type of question, it is not possible to rule out the possibility that questions with data were systematically different in their difficulty from questions without data.
+- **Changes in the data and machine forecasting platforms over time**.
+Both the data platform and the machine forecasting system suffered from various bugs and related issues, especially during the earlier portions of RCT-A.
+These problems in some cases resulted in incorrectly aggregated or otherwise inaccurate data, insufficient updating which led to data in the platform falling behind source availability and thus requiring forecasts over longer time periods than neccessary, and bugs in the machine forecaster that led to no or bad forecasts.
+The quality of the data and machine forecasts displayed to some users thus varied over time and IFPs, in ways that are difficult to reconstruct in retrospective.
 
-### Time series data was not available for all questions
+In respect to the first two problems, which we can quantify, the effective treatment groups were more complicated and are showin in Table XX.
 
-Additionally, chartable data was only available for about a third of the questions.
-The chart and model treatments thus are really part of a nested three layer set of factors: whether a question had chartable data, then whether a chart was displayed to a user, then whether a machine forecast was also displayed to a user.
-Whether a question had associated data depended very much on the type of question, ranging from questions like "What will be the gold price on [date]" where there is a clearly relevant time series that is easy to acquire and update, to "How many battle deaths will ACLED report in [country] in [date]?" where data are available but have to be transformed to a relevant series, to, on the other end, questions like "Who will win the World Cup?" where it is exceedingly difficult to acquire relevant time series data.
-Thus it is also possible that questions with chartable and without chartable data were systemtically different in their difficutly and average acccuracy levels.
-
-The effects of the two randomly assigned treatments, seeing a chart and seeing a model, are thus confounded by two other non-ranom factors: whether a forecaster was a volunteer or turker, and whether a question had chartable data or not.
-This effectively gives us an experiment design like in Table XXXX.
-
-| Condition | Is Turker? | Chartable data | Sees chart? | Sees models? |
-|-----------|------------|----------------|-------------|--------------|
-| A.1 |   |
-| A.2 | X |
-| A.3 |   | X |
-| A.4 | X | X |
-| B.1 |   |   |
-| B.2 |   | X | X |
-| C.1 |   |   |
-| C.2 | X |   |
-| C.3 |   | X | X | X |
-| C.4 | X | X | X | X |
+| Group|Condition          |Forecaster |IFP_Group       |Sees chart? |Sees model? |
+|-----:|:------------------|:----------|:---------------|:-----------|:-----------|
+|     1|A: no chart        |Turker     |No TS data      |            |            |
+|     2|A: no chart        |Turker     |Chart only      |            |            |
+|     3|A: no chart        |Turker     |Chart and model |            |            |
+|     4|A: no chart        |Volunteer  |No TS data      |            |            |
+|     5|A: no chart        |Volunteer  |Chart only      |            |            |
+|     6|A: no chart        |Volunteer  |Chart and model |            |            |
+|     7|B: chart only      |Volunteer  |No TS data      |            |            |
+|     8|B: chart only      |Volunteer  |Chart only      |X           |            |
+|     9|B: chart only      |Volunteer  |Chart and model |X           |            |
+|    10|C: chart and model |Turker     |No TS data      |            |            |
+|    11|C: chart and model |Turker     |Chart only      |X           |            |
+|    12|C: chart and model |Turker     |Chart and model |X           |X           |
+|    13|C: chart and model |Volunteer  |No TS data      |            |            |
+|    14|C: chart and model |Volunteer  |Chart only      |X           |            |
+|    15|C: chart and model |Volunteer  |Chart and model |X           |X           |
+|    16|Machine            |Machine    |Chart and model |            |            |
 
 Table: Effective treatments by group after addition of turkers to conditions A and C.
-
-### Varying quality of data and machine forecasts over time
-
-To complicate matters further, the quality of the data acquisition and machine forecasting modules changed over time during the RCT.
-Once a broader set of questions started flowing into the platform when the RCT started, a range of issues affecting both data quality and machine forecast quality arose.
-Examples include:
-
-- Incorrect aggregation of event data like ACLED and ICEWS that produced incorrect monthly counts of events or casualties.
-- Several questions were about monthly oil production rates for various countries.
-The resolution data are in PDF tables in monthly OPEC oil market reports.
-As this is hard to automatically scrape, an alternative source was used, but the values in this source do not always match the OPEC values.
-Nor, for that matter, do OPEC values always match in successive monthly reports.
-- Parsing failures that led the forecaster to produce non-sensical forecasts like negative values for count time series, or complete failures, e.g.
-when one question had a date like "April (04) 2018" instead of the expected "April 2018".
-- A limitation in the way Elastic Search aggregates non-standard fixed length time periods, e.g.
-periods of 40 days, required that for some questions data aggregation had to occur in the forecasting module, as opposed to data platform module, unlike planned.
-
-Most of these issues and bugs were fixed during the course of the RCT, and thus the quality of both the data and machine forecasts should have improved over time.
 
@@
 
-## Basic findings
+|Forecaster | avg_Brier|     n|
+|:----------|---------:|-----:|
+|Machine    |     0.386|  1975|
+|Turker     |     0.433| 39140|
+|Volunteer  |     0.322|  7816|
 
-### Brier over time
+Table: Average Brier score by forecaster group.
 
-Make point here that Brier generaly decreased over time? Maybe questions easier, maybe forecasters learn over time? Just establish that all analysis from here is for May 2nd and on only.
+Volunteers who had only charts did the best
 
-### Other
+|Forecaster | A: no chart| B: chart only| C: chart and model| Machine|
+|:----------|-----------:|-------------:|------------------:|-------:|
+|Machine    |          NA|            NA|                 NA|   0.386|
+|Turker     |       0.436|            NA|              0.432|      NA|
+|Volunteer  |       0.297|          0.25|              0.372|      NA|
 
-Machine forecasts were not more accurate than human forecasts.
+Table: Average Brier score by forecaster and condition groups.
 
-|Forecaster |Question         | A: no chart| B: chart only| C: chart and model|
-|:----------|:----------------|-----------:|-------------:|------------------:|
-|Volunteer  |IFP without data |        0.15|          0.09|               0.12|
-|Volunteer  |IFP had data     |        0.25|          0.22|               0.28|
-|Turker     |IFP without data |        0.15|            NA|               0.14|
-|Turker     |IFP had data     |        0.28|            NA|               0.27|
+Table XX summarizes the average Brier scores for all 16 treatment/forecaster/IFP groups in Table XX.
+As it is quite unwieldy, we include it for reference and will discuss specific insights in more details below.
 
-Table: MMDB by forecaster, data availability, and condition.
+| Group|Condition          |Forecaster |IFP_Group       | avg_Brier| sd_Brier|     n|
+|-----:|:------------------|:----------|:---------------|---------:|--------:|-----:|
+|     1|A: no chart        |Turker     |No TS data      |      0.45|     0.50|  6690|
+|     2|A: no chart        |Turker     |Chart only      |      0.55|     0.56|   304|
+|     3|A: no chart        |Turker     |Chart and model |      0.39|     0.40|  2907|
+|     4|A: no chart        |Volunteer  |No TS data      |      0.31|     0.54|   510|
+|     5|A: no chart        |Volunteer  |Chart only      |      0.36|     0.46|    18|
+|     6|A: no chart        |Volunteer  |Chart and model |      0.28|     0.34|   274|
+|     7|B: chart only      |Volunteer  |No TS data      |      0.26|     0.48|  1702|
+|     8|B: chart only      |Volunteer  |Chart only      |      0.42|     0.42|    57|
+|     9|B: chart only      |Volunteer  |Chart and model |      0.23|     0.31|   973|
+|    10|C: chart and model |Turker     |No TS data      |      0.45|     0.50| 19502|
+|    11|C: chart and model |Turker     |Chart only      |      0.56|     0.52|   990|
+|    12|C: chart and model |Turker     |Chart and model |      0.38|     0.40|  8747|
+|    13|C: chart and model |Volunteer  |No TS data      |      0.38|     0.60|  3090|
+|    14|C: chart and model |Volunteer  |Chart only      |      0.59|     0.57|    73|
+|    15|C: chart and model |Volunteer  |Chart and model |      0.32|     0.40|  1119|
+|    16|Machine            |Machine    |Chart and model |      0.39|     0.36|  1975|
 
-The low MMDB for condition B forecasters appears to be primarily driven by good performance on questions that *did not* have charts (keeping in mind that 2/3 of questions would not have had a chart).
+Table: Average Brier scores for all 15 distinct human IFP/treatment groups, as well as the machine forecasts.
 
-```{r tbl-accuracy-by-condition}
-read_csv("../output/tables/brier-by-conditions.csv") %>%
-  mutate(Condition = paste0(Condition, ".", c(1:4, 1:2, 1:4))) %>%
-  knitr::kable(digits = 2)
+## Turkers generally did worse
+
+Turkers overall had worse forecast accuracy than either the machine models or volunteers.
+To rule a spurious relationship, we can compare the accuracy in directly comparable pairs, e.g.
+rows 1 and 4 in Table XX consists of the average Brier scores for questions without TS data and in condition A treatment group for turkers and volunteers respectively, and show a difference of 0.14.
+
+Figure \ref{fig:turkers} shows the results of pairwise comparisons of the average Brier scores for all groups relative to Turker forecasters.
+Each row in the plot corresponds to a linear model for forecasters from the corresponding condition and IFP group, and the pointranges show coefficient estimates and 95\% confidence intervals for Turkers, Machine, and Volunteer forecasters.
+The Turker values correspond to the average Brier for them in this grouping, while the Machine and Volunteer estimates indicate the *relative* performance.
+I.e.
+for them, values below zero indicate an improvment.
+Note that the x-scale is reversed, so that better is further to the right.
+
+\begin{figure}
+\label{fig:turkers}
+\caption{Pairwise comparisons of turker versus machine and volunteer forecast accuracy.
+Each row shows estimates from a linear regression model with turkers as reference group, for different groupings of the experimental conditions and IFP data/model availability.
+The turker estimates (red dots) correspond to the average Brier for turker forecasters, while the other estimates correspond to the relative difference from the turker average for other forecasters.
+Points further to the right correspond to lower Brier scores, i.e.
+better peformance.}
+\includegraphics{../output/figures/pairwise-comparisons-turker.png}
+\end{figure}
+
+In most groups volunteers have an advantage sufficient to generate p-values below 0.05, although the effects are minute in comparison to general between-forecaster differences: all models have very small adjusted R^2 values, below 0.01.
+
+## Chart and model effects
+
+Teasing out the chart and model effects is a bit harder through pairwise comparison.
+
+Add in the LM and LMER model results here.
+
+```
+Coefficients:
+                          Estimate Std. Error t value Pr(>|t|)    
+(Intercept)               0.447126   0.002887 154.862  < 2e-16 ***
+ForecasterVolunteer      -0.104321   0.006208 -16.805  < 2e-16 ***
+IFP_GroupChart only       0.142377   0.017834   7.984 1.45e-15 ***
+IFP_GroupChart and model -0.063318   0.008648  -7.322 2.48e-13 ***
+sees_chartTRUE           -0.038067   0.015751  -2.417   0.0157 *  
+sees_modelTRUE            0.042337   0.015251   2.776   0.0055 ** 
+---
+Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+
+Residual standard error: 0.4776 on 46950 degrees of freedom
+Multiple R-squared:  0.01339,	Adjusted R-squared:  0.01328 
+F-statistic: 127.4 on 5 and 46950 DF,  p-value: < 2.2e-16
 ```
 
-Table: Mean Brier scores by condition
+```
+Linear mixed model fit by REML ['lmerMod']
+Formula: brier ~ Forecaster + IFP_Group + sees_chart + sees_model + (1 |      ifp_id)
+   Data: common_period[common_period$Forecaster != "Machine", ]
 
-## Turker performance compared to volunteer forecasters
+REML criterion at convergence: 50272.9
 
-## Machine forecast impact
+Scaled residuals: 
+    Min      1Q  Median      3Q     Max 
+-2.9464 -0.5432 -0.2071  0.3053  4.9013
 
-![Average mean daily Brier score (MDB) by forecaster type over time.
-Both volunteers and turkers improved over time.\label{fig:volunteers-only}](../output/figures/volunteers-only-accuracy-by-chartability-over-time.png)
+Random effects:
+ Groups   Name        Variance Std.Dev.
+ ifp_id   (Intercept) 0.06386  0.2527  
+ Residual             0.16808  0.4100  
+Number of obs: 46956, groups:  ifp_id, 156
 
-Figure \ref{fig:volunteers-only} looks only at volunteer forecasters in conditions B and C, showing again mean daily Brier scores over time.
-The right chart shows that forecasters who saw both a chart and model had slighly worse performance throughout the RCT compared to condition B forecasters who only saw the chart.
-This matches the hypothesis that seeing the machine forecasts hurt performance.
-However, if we look at the chart on the left, which shows performance for questions for which no data was available, i.e.
-where both condition B and C forecasters saw neither a chart nor a model, condition C performance was also worse up until sometime in late July.
+Fixed effects:
+                          Estimate Std. Error t value
+(Intercept)               0.438521   0.025392  17.270
+ForecasterVolunteer      -0.077777   0.005519 -14.091
+IFP_GroupChart only       0.160131   0.107529   1.489
+IFP_GroupChart and model -0.061051   0.044711  -1.365
+sees_chartTRUE           -0.021495   0.013689  -1.570
+sees_modelTRUE            0.026931   0.013280   2.028
 
-# Discussion/conclusion
+Correlation of Fixed Effects:
+            (Intr) FrcstV IFP_Go IFP_am ss_cTRUE
+FrcstrVlntr -0.042                              
+IFP_GrpChro -0.237  0.029                       
+IFP_GrpCham -0.567  0.005  0.140                
+ss_chrtTRUE  0.011 -0.266 -0.099 -0.066         
+ses_mdlTRUE -0.012  0.284  0.081 -0.016 -0.816  
+```
 
-Based on the quick check I did during the site visit, probably something like this:
+Relatively consistent resulst. First, turkers have worse performance. Second, seeing a chart somewhat helps, seeing a model somewhat hurts.
 
-- the machine forecasts were not accurate enough, although sans data and code issues they may have been 
-- Turkers don't forecast as well as volunteers
-- seeing the machine forecasts had a slight detrimental impact on accuracy, although maybe this changed as the forecasts became more accurate later in the RCT
-- seeing the charts improved performance, but I don't think it is clear why.
-Seeing the chart for instances where the data are hard to acquire (e.g.
-some data transformation are needed) did not boost accuracy more than in other series.
+## Models generally were disappointing, but did they do well in certain areas?
+
+This is now only looking at IFPs for which a machine forecast was available.
+
+Put the model based version with stars here instead of this:
+
+|Data_source      | N_IFPs| Machine| Turker| Volunteer|
+|:----------------|------:|-------:|------:|---------:|
+|ACLED 2op        |      7|   0.611|  0.533|     0.396|
+|ACLED 5op        |      4|   0.371|  0.554|     0.790|
+|earthquakes      |      2|   0.285|  0.207|     0.199|
+|eia              |      5|   0.184|  0.204|     0.145|
+|fao-fpi          |      3|   0.386|  0.220|     0.121|
+|fred             |      2|   0.580|  0.488|     0.290|
+|gold             |      3|   0.184|  0.274|     0.171|
+|imf              |      1|   0.222|  0.252|     0.325|
+|nigeria-security |      2|   0.355|  0.504|     0.648|
+|oecd             |      7|   0.430|  0.534|     0.356|
+|opec             |      5|   0.283|  0.222|     0.209|
+|other            |      2|   0.416|  0.574|     0.188|
+|privacyrights    |      2|   0.231|  0.321|     0.517|
+|stocks           |      4|   0.204|  0.204|     0.223|
+
+Models did relatively well with ACLED 5op, IMF, NST, privacyrights. Poorly on ACLED 2op, FAO, FRED, other?.
+
+## Why did condition B volunteers do so well?
+
+Condition B volunteers had the best overall performance out of the 16 groups shown in Table XX.
+One immediate and plausible story to explain why condition B forecasters did better than both A and C, even after controlling for the impact of Turker forecasters in the other groups, goes like this: we know the machine forecasts were disappointing and less accurate in general.
+Users who thus saw the chart and machine forecasts were adversely pulled to forecast in line with the machine prediction.
+On the other than, users who only saw the chart had a good baseline and were able to somehow accurately extrapolate from there.
+
+There are several inconsistencies in the data with this story.
+First, condition B forecasters did not only do better than other groups on questions that had data and a chart, but also on questions that did not have a chart.
+
+**evidence here**
+
+Maybe there is a "spillover" effect, where seeing the charts somehow made users just generally better forecasters.
+
+Do we see this in condition C for questions that didn't have a model?
+No.
+Condition C Turkers forecasting on a question without a chart did as poorly as condition A Turkers on the same set of questions (rows 1,2,10,11).
+Condition C volunteers on questions without a machine forecast did *worse* than both condition B and A volunteers on the same question sets.
+
+**evidence here**
+
+Another piece of evidence against the spillover effect is that accuracy is actually higher among users who only forecasted a small number of times, which suggests that learning did not take place.
+
+**evidence here**
+
+Not only were condition B volunteers uncannily accurate on questions even without a chart, they were uncannily accurate on questions where we know the displayed data were wrong.
+
+**evidence here**
+
+One possibility that we should not discard is that we are digging in noise.
+All of the effects---for turkers versus other forecasts, for the disappointing performance of the machine forecasts, for conditions or different groupings of IFPs---pale in comparison to differences in accuracy between individual forecasters and between forecasts for different IFPs.
+
+**maybe this is a good time to talk about the superactive forecasters and see whether excluding them changes results**
+
+# Conclusion
+
+Are we digging in noise?
+
+Many of the questions involve monthly time series but are only open for periods spanning a single or few months.
+Some of the advantage of the human forecasters may thus have been due to operating on a sub-monthly time scale.
+
+We also know that model performance suffered from issues related to automation.
+Were we able to run the RCT-A questions with the current system and with a data platform that optimally ingests data sources on the bleeding edge of their own updates, our forecasts would have been ...
+
+**check with updated results from test bed what the performance gain would have been, i think the site visit stats were not accurate anymore**
 

### 2018-11-30 R/analysis.R

@@ -7,6 +7,7 @@ library("betareg")
 library("rstan")
 library("rstanarm")
 library("forcats")
+library("broom")
 
 dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
 in_dir  <- file.path(dropbox_path, "input-data")
@@ -41,10 +42,17 @@ all_fcasts <- bind_rows(df1, df2) %>%
   left_join(data_sources %>% select(ifp_id, data_source), by = "ifp_id") %>%
   replace_na(list(data_source = "other")) %>%
   # add in summary info for chart/arima status
-  mutate(Data_status = case_when(
+  mutate(IFP_Group = case_when(
     has_historic_data & has_arima ~ "Chart and model",
     has_historic_data & !has_arima ~ "Chart only",
-    !has_historic_data ~ "None",
+    !has_historic_data ~ "No TS data",
+    TRUE ~ NA_character_
+  ), IFP_Group = factor(IFP_Group, levels = c("No TS data", "Chart only", "Chart and model"))) %>%
+  mutate(Condition = case_when(
+    condition=="a" ~ "A: no chart",
+    condition=="b" ~ "B: chart only",
+    condition=="c" ~ "C: chart and model",
+    condition=="n/a" ~ "Machine",
     TRUE ~ NA_character_
   ))
 
@@ -63,58 +71,172 @@ common_period %>%
 
 by_ifp <- common_period %>%
   group_by(ifp_id) %>%
-  summarize_at(vars(one_of("has_arima", "has_historic_data", "Data_status")), unique) %>%
+  summarize_at(vars(one_of("has_arima", "has_historic_data", "IFP_Group")), unique) %>%
   group_by(has_arima, has_historic_data) %>%
-  summarize(Data_status = unique(Data_status), n = n()) 
+  summarize(IFP_Group = unique(IFP_Group), n = n()) 
+write_csv(by_ifp, "output/tables/n-ifps-by-data-status.csv")
 sum(by_ifp$n)
 by_ifp
 
 # Overall condition table
-cond_tbl <- common_period %>%
+brier_by_forecaster <- common_period %>%
   group_by(Forecaster) %>%
-  summarize(avg_Brier = mean(brier))
-cond_tbl
+  summarize(avg_Brier = mean(brier), n = n()) %>%
+  write_csv(., "output/tables/brier-by-forecaster.csv")
+brier_by_forecaster %>%
+  knitr::kable(digits = 3)
+
+brier_by_forecaster_condition <- common_period %>%
[…]

### 2018-12-03 paper/paper.Rmd
add chart and model effect models to paper

@@
 header-includes:
-  - 
+  - \usepackage{mathpazo}
 ---
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+
+library("readr")
+library("dplyr")
+library("knitr")
 ```
@@
 
-|Forecaster | avg_Brier|     n|
-|:----------|---------:|-----:|
-|Machine    |     0.386|  1975|
-|Turker     |     0.433| 39140|
-|Volunteer  |     0.322|  7816|
-
-Table: Average Brier score by forecaster group.
+```{r}
+knitr::opts_current$set(label = "brier-by-forecaster")
+tbl <- read_csv("../output/tables/brier-by-forecaster.csv")
+knitr::kable(tbl, format = "latex", digits = 2, booktabs = TRUE,
+             caption = "Average Brier score by forecaster group")
+```
 
@@
 Teasing out the chart and model effects is a bit harder through pairwise comparison.
+Instead, we directly coded whether a forecaster saw a chart or model forecast.
+Forcasters in conditions B and C and for questions that had a chart saw a chart, and forecasters in condition C looking at a question which had a machine forecast will have seen a model forecast.
+We also leave out the machine forecasts from the data, leaving almost 47,000 forecasts, of which about 25% were made with a chart available, and 21% also had a model forecast available.
 
-Add in the LM and LMER model results here.
+We estimated two linear models to predict Brier scores for a forecast. The first had the specifiction:
 
-```
-Coefficients:
-                          Estimate Std. Error t value Pr(>|t|)    
-(Intercept)               0.447126   0.002887 154.862  < 2e-16 ***
-ForecasterVolunteer      -0.104321   0.006208 -16.805  < 2e-16 ***
-IFP_GroupChart only       0.142377   0.017834   7.984 1.45e-15 ***
-IFP_GroupChart and model -0.063318   0.008648  -7.322 2.48e-13 ***
-sees_chartTRUE           -0.038067   0.015751  -2.417   0.0157 *  
-sees_modelTRUE            0.042337   0.015251   2.776   0.0055 ** 
----
-Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+\[
+\textrm{Brier} \sim \alpha + \beta_1\textrm{Sees model} + \beta_2\textrm{Sees chart} + \beta_3\textrm{}
+\]
 
-Residual standard error: 0.4776 on 46950 degrees of freedom
-Multiple R-squared:  0.01339,	Adjusted R-squared:  0.01328 
-F-statistic: 127.4 on 5 and 46950 DF,  p-value: < 2.2e-16
-```
-
-```
-Linear mixed model fit by REML ['lmerMod']
-Formula: brier ~ Forecaster + IFP_Group + sees_chart + sees_model + (1 |      ifp_id)
-   Data: common_period[common_period$Forecaster != "Machine", ]
-
-REML criterion at convergence: 50272.9
-
-Scaled residuals: 
-    Min      1Q  Median      3Q     Max 
--2.9464 -0.5432 -0.2071  0.3053  4.9013
-
-Random effects:
- Groups   Name        Variance Std.Dev.
- ifp_id   (Intercept) 0.06386  0.2527  
- Residual             0.16808  0.4100  
-Number of obs: 46956, groups:  ifp_id, 156
-
-Fixed effects:
-                          Estimate Std. Error t value
-(Intercept)               0.438521   0.025392  17.270
-ForecasterVolunteer      -0.077777   0.005519 -14.091
-IFP_GroupChart only       0.160131   0.107529   1.489
-IFP_GroupChart and model -0.061051   0.044711  -1.365
-sees_chartTRUE           -0.021495   0.013689  -1.570
-sees_modelTRUE            0.026931   0.013280   2.028
-
-Correlation of Fixed Effects:
-            (Intr) FrcstV IFP_Go IFP_am ss_cTRUE
-FrcstrVlntr -0.042                              
-IFP_GrpChro -0.237  0.029                       
-IFP_GrpCham -0.567  0.005  0.140                
-ss_chrtTRUE  0.011 -0.266 -0.099 -0.066         
-ses_mdlTRUE -0.012  0.284  0.081 -0.016 -0.816  
-```
+The intercept $\alpha$ corresponds to the average Brier score for a Turker forecasting on a IFP that did not have chartable data, and who neither saw a chart nor machine forecast.
 
 Relatively consistent resulst. First, turkers have worse performance. Second, seeing a chart somewhat helps, seeing a model somewhat hurts.
+
+\begin{figure}
+\caption{\label{fig:model-chart-and-model-effects} Model estimates}
+\includegraphics{../output/figures/model-chart-and-model-effects.png}
+\end{figure}
+
+\begin{table}
+\caption{\label{tab:mdl-summary-stats}foo}
+\centering
+\begin{tabular}[t]{llr}
+\toprule
+Model & Statistic & Value\\
+\midrule
+Model 1: Linear model & N & 46956\\
+ & R$^2$ & 0.013\\
+ & Adj. R$^2$ & 0.013\\
+Model 2: Linear model with IFP random intercepts & N & 46956\\
+ & Marginal R$^2$ & 0.011\\
+ & Conditional R$^2$ & 0.283\\
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2018-12-03 R/analysis.R

@@ -137,7 +137,7 @@ turker_pairs <- common_period %>%
   filter(condition!="b") %>%
   mutate(Condition = fct_recode(Condition, `C: chart and model` = "Machine")) %>%
   mutate(Forecaster = factor(Forecaster, levels = c("Turker", "Volunteer", "Machine"))) %>%
-  unite(Group, Condition, IFP_Group)
+  unite(Group, Condition, IFP_Group, sep = "\n")
 group_all <- turker_pairs %>%
   mutate(Group = "All") 
 turker_pairs <- bind_rows(turker_pairs, group_all) %>%
@@ -159,17 +159,62 @@ turker_pairs %>%
   geom_hline(yintercept = 0, linetype = 3) +
   theme_minimal() +
   labs(y = "Average Brier / change in Average Brier") +
-  theme(legend.position = "top")
+  theme(legend.position = "top") +
+  scale_color_discrete("Coefficient:")
 ggsave("output/figures/pairwise-comparisons-turker.png", height = 4, width = 7)
 
 
 
 # Linear models for chart/model effects -----------------------------------
+#
+#   Maybe add an interaction model here?
+#   summary(lm(brier ~ Forecaster + IFP_Group + sees_chart:Forecaster + sees_model:Forecaster, data = common_period_no_machine))
+#
 
-summary(lm(brier ~ Forecaster + IFP_Group + sees_chart + sees_model, data = common_period[common_period$Forecaster!="Machine", ]))
 
-summary(lmer(brier ~ Forecaster + IFP_Group + sees_chart + sees_model + (1|ifp_id), common_period[common_period$Forecaster!="Machine", ]))
 
+common_period_no_machine <- common_period[common_period$Forecaster!="Machine", ]
+
+# % who saw a chart
+table(common_period_no_machine$sees_chart) / nrow(common_period_no_machine) * 100
+# % who saw a model
+table(common_period_no_machine$sees_model) / nrow(common_period_no_machine) * 100
+
+mdl1 <- lm(brier ~ Forecaster + IFP_Group + sees_chart + sees_model, 
+           data = common_period_no_machine)
+mdl2 <- lmer(brier ~ Forecaster + IFP_Group + sees_chart + sees_model + (1|ifp_id),
+             data = common_period_no_machine)
+
+bind_rows(
+  tidy(mdl1) %>% mutate(Model = "Model 1: linear model"),
[…]

### 2018-12-05 paper/paper.Rmd
first draft, more or less

@@
     latex_engine: xelatex
+    keep_tex: true
 abstract: "\\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project to develop hybrid geopolitical forecasting systems that combine human- and machine-generated forecasts.
@@
 header-includes:
-  - \usepackage{mathpazo}
+  - \usepackage[T1]{fontenc}
 ---
@@
 library("knitr")
+library("tidyr")
 ```
@@
 
-Table: Summary of original research design.
+Table: Summary of original research design. \label{tab:original-design}
 
@@
 
-In respect to the first two problems, which we can quantify, the effective treatment groups were more complicated and are showin in Table XX.
+In respect to the first two problems, which we can quantify, the effective treatment groups were more complicated and are showin in Table \ref{tab:all-conditions}.
 
@@
 
-Table: Effective treatments by group after addition of turkers to conditions A and C.
+Table: Effective treatments by group after addition of turkers to conditions A and C. \label{tab:all-conditions}
 
@@
 
-|Forecaster | A: no chart| B: chart only| C: chart and model| Machine|
-|:----------|-----------:|-------------:|------------------:|-------:|
-|Machine    |          NA|            NA|                 NA|   0.386|
-|Turker     |       0.436|            NA|              0.432|      NA|
-|Volunteer  |       0.297|          0.25|              0.372|      NA|
+```{r}
+options(knitr.kable.NA = '')
+knitr::opts_current$set(label = "brier-by-forecaster-condition")
+tbl <- read_csv("../output/tables/brier-by-forecaster-condition.csv")
+tbl %>%
+  spread(Condition, avg_Brier) %>%
+  knitr::kable(., format = "latex", digits = 2, booktabs = TRUE,
+             caption = "Average Brier score by forecaster group")
+```
 
-Table: Average Brier score by forecaster and condition groups.
-
-Table XX summarizes the average Brier scores for all 16 treatment/forecaster/IFP groups in Table XX.
+Table \ref{tab:brier-by-forecaster-ifp-group-condition} summarizes the average Brier scores for all 16 treatment/forecaster/IFP groups in Table XX.
 As it is quite unwieldy, we include it for reference and will discuss specific insights in more details below.
 
-| Group|Condition          |Forecaster |IFP_Group       | avg_Brier| sd_Brier|     n|
-|-----:|:------------------|:----------|:---------------|---------:|--------:|-----:|
-|     1|A: no chart        |Turker     |No TS data      |      0.45|     0.50|  6690|
-|     2|A: no chart        |Turker     |Chart only      |      0.55|     0.56|   304|
-|     3|A: no chart        |Turker     |Chart and model |      0.39|     0.40|  2907|
-|     4|A: no chart        |Volunteer  |No TS data      |      0.31|     0.54|   510|
-|     5|A: no chart        |Volunteer  |Chart only      |      0.36|     0.46|    18|
-|     6|A: no chart        |Volunteer  |Chart and model |      0.28|     0.34|   274|
-|     7|B: chart only      |Volunteer  |No TS data      |      0.26|     0.48|  1702|
-|     8|B: chart only      |Volunteer  |Chart only      |      0.42|     0.42|    57|
-|     9|B: chart only      |Volunteer  |Chart and model |      0.23|     0.31|   973|
-|    10|C: chart and model |Turker     |No TS data      |      0.45|     0.50| 19502|
-|    11|C: chart and model |Turker     |Chart only      |      0.56|     0.52|   990|
-|    12|C: chart and model |Turker     |Chart and model |      0.38|     0.40|  8747|
-|    13|C: chart and model |Volunteer  |No TS data      |      0.38|     0.60|  3090|
-|    14|C: chart and model |Volunteer  |Chart only      |      0.59|     0.57|    73|
-|    15|C: chart and model |Volunteer  |Chart and model |      0.32|     0.40|  1119|
-|    16|Machine            |Machine    |Chart and model |      0.39|     0.36|  1975|
-
-Table: Average Brier scores for all 15 distinct human IFP/treatment groups, as well as the machine forecasts.
+```{r brier-by-forecaster-ifp-group-condition}
+knitr::opts_current$set(label = "brier-by-forecaster-ifp-group-condition")
+tbl <- read_csv("../output/tables/brier-by-forecaster-ifp-group-condition.csv")
+knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
+             caption = "Average Brier scores for all 15 distinct human IFP/treatment groups, as well as the machine forecasts",
+             linesep = c(rep("", 5), '\\addlinespace', 
+                         rep("", 2), '\\addlinespace', 
+                         rep("", 5), '\\addlinespace'))
+```
 
@@
 To rule a spurious relationship, we can compare the accuracy in directly comparable pairs, e.g.
-rows 1 and 4 in Table XX consists of the average Brier scores for questions without TS data and in condition A treatment group for turkers and volunteers respectively, and show a difference of 0.14.
+rows 1 and 4 in Table \ref{tab:brier-by-forecaster-ifp-group-condition} consists of the average Brier scores for questions without TS data and in condition A treatment group for turkers and volunteers respectively, and show a difference of 0.14.
 
@@
 Points further to the right correspond to lower Brier scores, i.e.
-better peformance.}
+better peformance.
+Volunteers on average performed better than turker forecasters in all conditions and question groups, except on the small number of charts-only IFPs, where estimates are uncertain.}
 \includegraphics{../output/figures/pairwise-comparisons-turker.png}
@@
 
-We estimated two linear models to predict Brier scores for a forecast. The first had the specifiction:
+We estimated two linear models to predict Brier scores for a forecast. The first has the specifiction:
 
-\[
-\textrm{Brier} \sim \alpha + \beta_1\textrm{Sees model} + \beta_2\textrm{Sees chart} + \beta_3\textrm{}
-\]
+$$
+\begin{aligned}
+\textrm{Brier} =~& \alpha + \beta_1\textrm{SeesModel} + \beta_2\textrm{SeesChart} + \beta_3\textrm{IFPGroupChartOnly} + \beta_4\textrm{IFPGroupChartAndModel} \\
+ & + \beta_5\textrm{ForecasterVolunteer} 
+\end{aligned}
+$$
 
-The intercept $\alpha$ corresponds to the average Brier score for a Turker forecasting on a IFP that did not have chartable data, and who neither saw a chart nor machine forecast.
+All variables are dummy terms.
+The intercept $\alpha$ corresponds to the average Brier score for a Turker forecasting on an IFP that did not have chartable data, and who neither saw a chart nor machine forecast.
 
-Relatively consistent resulst. First, turkers have worse performance. Second, seeing a chart somewhat helps, seeing a model somewhat hurts.
+The second model includes random intercepts for IFP to account for the possibility that seeings charts or models may make users differentially willing to forecast on harder IFPs, i.e.
+those with higher average Brier scores.
+Otherwise the specification of covariates is the same.
+
+$$
+\begin{aligned}
+\textrm{Brier} =~& \alpha_{\textrm{Global}} + \alpha_{\textrm{IFP}} + \beta_1\textrm{SeesModel} + \beta_2\textrm{SeesChart} + \beta_3\textrm{IFPGroupChartOnly}  \\
+ & + \beta_4\textrm{IFPGroupChartAndModel} + \beta_5\textrm{ForecasterVolunteer} 
+\end{aligned}
+$$
+
+Figure \ref{fig:model-chart-and-model-effects} plots the coefficient estimates from both models.
+The results are consistent.
+Seeing a chart slightly helps, seeing a model slightly hurts forecast performance.
+However, the effects are small and not as important as whether a forecasters was a volunteer or turker.
+In addition, the models themselves capture only a small portion of the variation in Brier scores overall.
+Table \ref{tab:model-summary-stats} shows $R^2$ statistics for both models.
+The covariates account for less than 0.01 of the variation in Brier scores; the inclusion of random IFP intercepts in model 2 increases this to 0.28 (conditional $R^2$), but the fraction explained by the covariates together is still 0.01 (marginal $R^2$).
+Most of the variation remains either between forecasters or due to other random factors.
+
+Thus, while we find that exposure to charts and model forecasts has a small impact on forecast quality, the effect sizes are small in comparison to other systematic factors in the model, and small relative to unsystematic and random factors not in the model.
 
 \begin{figure}
-\caption{\label{fig:model-chart-and-model-effects} Model estimates}
+\caption{\label{fig:model-chart-and-model-effects} Model estimates for the direct effects of seeing a chart and machine forecast.}
 \includegraphics{../output/figures/model-chart-and-model-effects.png}
@@
 
-\begin{table}
-\caption{\label{tab:mdl-summary-stats}foo}
-\centering
-\begin{tabular}[t]{llr}
-\toprule
-Model & Statistic & Value\\
-\midrule
-Model 1: Linear model & N & 46956\\
- & R$^2$ & 0.013\\
- & Adj. R$^2$ & 0.013\\
-Model 2: Linear model with IFP random intercepts & N & 46956\\
- & Marginal R$^2$ & 0.011\\
- & Conditional R$^2$ & 0.283\\
-\bottomrule
-\end{tabular}
-\end{table}
+```{r model-summary-stats, result='asis'}
+knitr::opts_current$set(label = "model-summary-stats")
+tbl <- read_csv("../output/tables/model-summary-stats.csv")
+knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
+             caption = "Model summary statistics") %>%
+  gsub("R\\\\textasciicircum\\{\\}2", "$R^2$", .) %>%
+  # cut off the integer trailing digits
+  gsub("\\.00", "", .)
+```
 
@@
 
-This is now only looking at IFPs for which a machine forecast was available.
+Although the machine models overall underperformed, it is possible that they performed well relative to human forecasters in certain areas.
+To examine this we looked at all forecasts for questions that had a machine forecast.
+This leaves us with 15,995 forecasts, approximately 2,000 of which were machine and volunteer forecasts, respectively, and the remaining 12,000 from turkers.
 
-Put the model based version with stars here instead of this:
+To estimate the relative performance of different forecasters, we again estimated a simple linear model, similar to those used for the turker pariwise comparisons. The model has the general form:
+
+$$
+\begin{aligned}
+\textrm{Brier} =~& \beta_{i}\textrm{DataSource} + \beta'_{ij}(\textrm{DataSource}\times\textrm{Forecaster})
+\end{aligned}
+$$
+
+Data source is a discrete variable with $i=14$ levels for the different data sources reflected in the set of IFPs.
+Based on previous exploratory analysis, we separated one data source, ACLED questions, by whether a question was binary with two answer options ("yes"/"no"), or not (3 or 5 options).
+Forecaster is also a discrete variable with 3 levels, but we leave out "Machine" as reference category, giving us $j=2$ different levels.
+Thus, each of the $\beta$ coefficients corresponds to the average Brier score for machine forecasts for that group of IFPs, and each $\beta'_i$ coefficient is the relative difference of the average volunteer and turker Brier score compared to the corresponding baseline in $\beta_i$.
+
+\begin{figure}
+\caption{\label{fig:model-machine-by-data-source} Linear model estimates for the relative performance of machine to volunteer and turker forecasts by data source.
+The left panel shows the baseline machine performance for each data source, the right panel shows the relative difference of the volunteer and turker forecasts to the corresponding baseline.
+The outcome variable is a forecasts's Brier score.}
+\includegraphics{../output/figures/model-machine-by-data-source.png}
+\end{figure}
+
+Figure \ref{fig:model-machine-by-data-source} shows the coefficient estimates.
+The left panel shows the baseline machine forecast performance.
+The right panel shows the relative difference in average Brier scores for turker and volunteer forecasts; points to the left of the dotted 0 line indicate that volunteer or turker forecasters on average outperformed the machine forecasts.
+
+The machine forecasts did better than human forecasts on ACLED 5-option, Boko Haram killings ("nigeria-security"), and hacking ("privacyrights") questions.
+All three data sources consist of count data derived from original event datasets, i.e.
+they require data aggregation before one can see a time series relevant to the question at hand, which might be a factor in the relative performance on these questions.
+On the other hand, human forecasters consistently did better on ACLED 2-option, food price index ("FAO-FPI"), financial ("fred"), and oil production ("opec") questions.
+The results for ACLED 2-option questions might be influenced by the particular mechanism by which time series forecasts were converted to answer option probabilities, which may explain the inconsistency with ACLED 5-option questions.
+
+There are some hard to quantify factors that may explain some of the variation in human to machine performance:
+
+1.
+As mentioned, some data sources (ACLED, privacyrights, nigeria-security) required relatively complicated data aggregation to derive usable count series, and which therefore may not have been accessible to all human forecasters.
+2.
+Not all data sources are equally easy to automatically download and update, and at least in some of the data sources it was the case that models had to forecast over much longer time horizons due to outdated data, while a human forecaster would have been able to access more recent data.
+3.
+Due to obstacles in data acquisition, in at least one case--OPEC oil production rates--a secondary and less accurate source was substituted.
+OPEC production values are delivered in monthly PDF reports that are difficult to automatically extract data from, and the alternative data source in some instances significantly deviated from the OPEC values.
+Thus models for these questions would have been forecasting with incorrect input data.
+4.
+For questions in which the data source is at a monthly (or higher) resolution, humans have an advantage over univariate time series models in that they can incorporate sub-monthly information into their forecast.
+For many of these questions, if they were open for a period of one or two months, there would in fact have only been one or two distinct time series forecasts (barring practical live system issues and bugs).
+
+A caveat to these trends is that the number of IFPs and forecasts for each data source are small and the results may thus be due to random variation. Results from future RCTs should test this.
+
+## Why did condition B volunteers do so well?
+
+Condition B volunteers had the best overall performance out of the 16 groups shown in Table \ref{tab:brier-by-forecaster-ifp-group-condition}.
+Considering that the machine forecasts were relatively bad, one possible explanation is that the charts overall had a positive impact on human forecast accuracy, while seeing machine forecasts had a negative impact.
+
+However, there are several additional patterns that are either inconsistent with this simple explanation or at least suggest the need for a more complex alternative.
+The first is that condition B forecasters did not only do better than other groups on questions that had a chart, but also on questions that did not have a chart.
+Table \ref{tab:b-advantage} shows the relevant rows from the average Brier scores for all forecaster groups in Table \ref{tab:brier-by-forecaster-ifp-group-condition}.
+With the partial exception of the few chart-only questions, condition B forecasters' advantage on questions without a chart is as pronounced as their advantage in questions that did have a chart.
+Considering that there were two to three times as many forecasts on questions without chartable data, most of the condition B volunteer advantage is driven by doing better on questions that did not have a chart.
+
+```{r}
+knitr::opts_current$set(label = "b-advantage")
+tbl <- read_csv("../output/tables/volunteer-brier-by-condition-ifp-group.csv")
+knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
+             caption = "Condition B advantage",
+             linesep = c(rep("", 2), '\\addlinespace')) 
+```
+
+We could exlain this if there is, in addition to the direct beneficial effect of seeing a chart, but not a potentially misleading machine forecast, a spillover effect whereby exposure to charts also improves a volunteer forecaster's ability to forecast on other questions without charts.
+
+Curious inconsistencies remain:
+
+**There appears to be no chart spillover effect for condition C human and volunteer forecasters.** Tabe \ref{tab:brier-by-forecaster-ifp-group-condition} shows that condition C volunteers did worse on questions without time-series data than the condition A control group.
+Condition C turkers also did not do better than their condition A control group counterparts.
+Unlike condition B forecasters, they were exposed to machine forecasts as well, so maybe there is an equal but opposite negative machine spillover effect.
+
+**Condition B volunteers who forecast a lot do no better on non-chart questions than novices.** Figure \ref{fig:cond-b-volunteers-n-vs-brier} show average Brier scores as a function of the total number of forecasts a user has.
+The data consist of only forecasts from condition B volunteers.
+Each point is the average Brier score for a user's forecasts on the set of questions that did (blue) or did not have a chart (red).
+Novice users tend to be as accurate as intensive users on non-chart questions.
+This implies that the condition B volunteer advantage on non-chart questions is not related to how active a user was.
+If there is a spillover effect, even very inactive users seem to somehow benefit from it.
+
+\begin{figure}
+\caption{\label{fig:cond-b-volunteers-n-vs-brier} Average Brier for condition B volunteers by user and by whether a IFP had a chart.
+Intensive users are slightly more accurate than novice users on questions with a chart, but no more accurate on questions without a chart.}
+\includegraphics{../output/figures/cond-b-volunteers-n-vs-brier.png}
+\end{figure}
+
+**Condition C volunteers do well on questions where we know that data displayed was wrong.** There were several instances where we know that the data displayed in a chart were inaccurate.
+Figure \ref{fig:iraq} shows the first example, oil production figures for Iraq.
+The main chart shows several teim series depicting oil production in Iraq, according to different sources.
+The canonical data are contained in monthly OPEC oil market reports, each of which contains production figures for the past 3 months.
+Those are shown in the sequence of short time series on the mid-right part of the main chart--the figures are somewhat inconsistent between reports, but this is another peculiarity.
+Because this data is difficult to obtain programatically, an alternative data source is used instead.
+This is what would have been shown in the charts, and what informed the machine forecast.
+Obviously it is incorrect data.
+The dotted lines show the values separating the answer options for this question.
+The chart placed the production values in an entirely incorrect answer category.
+
+\begin{figure}
+\caption{\label{fig:iraq} Iraq oil production}
+\includegraphics{../output/figures/opec-iraq-divergence.png}
+\end{figure}
+
+Curiously, the condition B volunteers performed better on this question than any other forecaster.
+The number of forecasts here are low, and only the last three groups, starting from the machine forecasts, have pairwise statistically significant differences in average Brier score from the condition B volunteers.
+
+|Forecaster |condition | avg_Brier|  n|
+|:----------|:---------|---------:|--:|
+|Volunteer  |b         |      0.07|  5|
+|Volunteer  |a         |      0.11|  5|
+|Turker     |c         |      0.18| 87|
+|Machine    |n/a       |      0.25| 14|
+|Turker     |a         |      0.35| 23|
+|Volunteer  |c         |      0.45| 21|
+
+Another example with incorrect data is an early ICEWS question that fell outside the common period and thus did not have forecasts from turkers.
+Condition B volunteers also outperformed all other groups on this question, although the pairwise differences with the other groups are not statistically significant.
+
+|Forecaster |condition | avg_Brier|  n|
+|:----------|:---------|---------:|--:|
+|Volunteer  |b         |      0.92| 14|
+|Machine    |n/a       |      1.00| 18|
+|Volunteer  |a         |      1.11|  7|
+|Volunteer  |c         |      1.11| 25|
+
+\clearpage
+
+**Even when the machine predictions were good, condition B volunteers did better than their C counterparts.** If bad machine predictions influence human forecasts and are the reason why the beneficial chart effecgt is only apparent in condition B forecasters, and not condition C forecasters, then it should also be the case that condition C forecasters do well on those questions where the machine forecasts were good.
+This is no the case.
+The table below shows the subset of forecasts on questions that had a machine forecast (and thus chart) and where the machine forecast average Brier was lower than the average Brier score for the forecasts from all other forecaster $\times$ condition groups.
+Condition B volunteers beat their condition C counterparts even when the machine model should have improved their forecasts because they happened to be accurate.
+
+|Forecaster |Condition          | avg_Brier| n_fcasts|
+|:----------|:------------------|---------:|--------:|
+|Machine    |Machine            |      0.39|     1975|
+|Turker     |A: no chart        |      0.39|     2810|
+|Turker     |C: chart and model |      0.38|     8482|
+|Volunteer  |A: no chart        |      0.28|      267|
+|Volunteer  |B: chart only      |      0.23|      958|
+|Volunteer  |C: chart and model |      0.32|     1094|
+
+One possibility that we should not discard with these apparent inconsistencies is that we are digging in noise.
+All of the effects---for turkers versus other forecasts, for the disappointing performance of the machine forecasts, for conditions or different groupings of IFPs---pale in comparison to differences in accuracy between individual forecasters and between forecasts for different IFPs.
+But on the other hand there are several inconsistencies so it also is a stretch to write all of them off to random coincidence.
+Condition B volunteers clearly did better than all other forecasters/condition groups, but it is not clear why.
+
+# Conclusion
+
+Are we digging in noise?
+
+Many of the questions involve monthly time series but are only open for periods spanning a single or few months.
+Some of the advantage of the human forecasters may thus have been due to operating on a sub-monthly time scale.
+
+We also know that model performance suffered from issues related to automation.
+Were we able to run the RCT-A questions with the current system and with a data platform that optimally ingests data sources on the bleeding edge of their own updates, our forecasts would have been ...
+
+**[check with updated results from test bed what the performance gain would have been, i think the site visit stats were not accurate anymore]**
+
+# Additional information
 
@@
 
-Models did relatively well with ACLED 5op, IMF, NST, privacyrights. Poorly on ACLED 2op, FAO, FRED, other?.
-
-## Why did condition B volunteers do so well?
-
-Condition B volunteers had the best overall performance out of the 16 groups shown in Table XX.
-One immediate and plausible story to explain why condition B forecasters did better than both A and C, even after controlling for the impact of Turker forecasters in the other groups, goes like this: we know the machine forecasts were disappointing and less accurate in general.
-Users who thus saw the chart and machine forecasts were adversely pulled to forecast in line with the machine prediction.
-On the other than, users who only saw the chart had a good baseline and were able to somehow accurately extrapolate from there.
-
-There are several inconsistencies in the data with this story.
-First, condition B forecasters did not only do better than other groups on questions that had data and a chart, but also on questions that did not have a chart.
-
-**evidence here**
-
-Maybe there is a "spillover" effect, where seeing the charts somehow made users just generally better forecasters.
-
-Do we see this in condition C for questions that didn't have a model?
-No.
-Condition C Turkers forecasting on a question without a chart did as poorly as condition A Turkers on the same set of questions (rows 1,2,10,11).
-Condition C volunteers on questions without a machine forecast did *worse* than both condition B and A volunteers on the same question sets.
-
-**evidence here**
-
-Another piece of evidence against the spillover effect is that accuracy is actually higher among users who only forecasted a small number of times, which suggests that learning did not take place.
-
-**evidence here**
-
-Not only were condition B volunteers uncannily accurate on questions even without a chart, they were uncannily accurate on questions where we know the displayed data were wrong.
-
-**evidence here**
-
-One possibility that we should not discard is that we are digging in noise.
-All of the effects---for turkers versus other forecasts, for the disappointing performance of the machine forecasts, for conditions or different groupings of IFPs---pale in comparison to differences in accuracy between individual forecasters and between forecasts for different IFPs.
-
-**maybe this is a good time to talk about the superactive forecasters and see whether excluding them changes results**
-
-# Conclusion
-
-Are we digging in noise?
-
-Many of the questions involve monthly time series but are only open for periods spanning a single or few months.
-Some of the advantage of the human forecasters may thus have been due to operating on a sub-monthly time scale.
-
-We also know that model performance suffered from issues related to automation.
-Were we able to run the RCT-A questions with the current system and with a data platform that optimally ingests data sources on the bleeding edge of their own updates, our forecasts would have been ...
-
-**check with updated results from test bed what the performance gain would have been, i think the site visit stats were not accurate anymore**
+Table: Average Brier scores for machine and other forecasts by question data source.
 

### 2018-12-05 paper/paper.tex
first draft, more or less

--- 2018-12-05 paper/paper.Rmd
@@
----
-title: "Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
-author: "Andreas Beger and Michael D. Ward"
-date: "`r Sys.Date()`"
-output: 
-  pdf_document:
-    latex_engine: xelatex
-    keep_tex: true
-abstract: "\\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project to develop hybrid geopolitical forecasting systems that combine human- and machine-generated forecasts.
-The first project trial period took place in 2018, during the course of which volunteer participants, Amazon Mechanical Turkers, and an automated time series forecasting module we developed provided forecasts for more than 150 questions covering a diverse set of topics.
-We investigate two questions: (1) how does the performance of turker forecasters compare to volunteers, and (2) what impact did access to the machine forecasts have on forecaster accuracy?
-\\newline \\newline
-"
-bibliography: "../../whistle/master.bib"
-header-includes:
-  - \usepackage[T1]{fontenc}
----
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition},
+            pdfauthor={Andreas Beger and Michael D. Ward},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-library("readr")
-library("dplyr")
-library("knitr")
-library("tidyr")
-```
+%%% Change title format to be more compact
+\usepackage{titling}
 
-The Hybrid Forecasting Competition (HFC)^[**Funding acknowledgement**: This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
-The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S.
-Government.
-The U.S.
-Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.] is an IARPA program that seeks to develop methods for hybrid geopolitical forecasting system that combine human and machine forecasts to answer a broad range of questions about economic, political, health, and other events and trends.
-The first trial period, or RCT, took place in 2018, during the course of which hundreds of volunteer and Amazon Mechnial Turk forecasters, as well as automated machine models, answered more than 150 questions covering a broad range of issues.
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
 
-We worked on one of the competition teams, and specifically by contributing a time series forecasting module.
-Out of the large set of interesting questions one could examine with the results so far, given our specific focus on this project, we will try to examine in this paper two questions: (1) how does the accuracy of turker forecasters compare to volunteer forecasters, and (2) what was the impact of the machine forecasts on human forecaster accuracy?
+\setlength{\droptitle}{-2em}
 
-# The Hybrid Forecasting Competition
+  \title{Assessing Amazon turker and automated machine forecasts in the Hybrid
+Forecasting Competition}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{Andreas Beger and Michael D. Ward}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+      \predate{\centering\large\emph}
+  \postdate{\par}
+    \date{2018-12-05}
 
-The goal of the HFC is to find ways to optimally combine human and machine forecasts.
-For example, machine forecasts can be reliable and scalable, but are constrained by available data, idiosyncratic questions, and cold start problems when a corpus of historical data is not available.
-Human-generated forecasts on the other hand are more flexible, but also more costly to scale and subject to various cognitive biases.[e.g.
-@kahneman:2011].
+\usepackage[T1]{fontenc}
 
-The competition is organized around Good Judgement-style forecasting tournaments [@tetlock:2005, @tetlock:etal:2017], where forecasters answer and are scored on a diverse pool of questions.
-Examples from the first trial period, RCT-A, include:
+\begin{document}
+\maketitle
+\begin{abstract}
+\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project
+to develop hybrid geopolitical forecasting systems that combine human-
+and machine-generated forecasts. The first project trial period took
+place in 2018, during the course of which volunteer participants, Amazon
+Mechanical Turkers, and an automated time series forecasting module we
+developed provided forecasts for more than 150 questions covering a
+diverse set of topics. We investigate two questions: (1) how does the
+performance of turker forecasters compare to volunteers, and (2) what
+impact did access to the machine forecasts have on forecaster accuracy?
+\newline \newline
+\end{abstract}
 
-- What will be the long-term interest rate for South Africa (ZAF) in July 2018?
-- How many deaths perpetrated by Boko Haram will the Council on Foreign Relations report for June 2018?
-- What will be the daily closing spot price of Brent crude oil (USD per barrel) on 31 May 2018, according to the U.S. EIA?
+The Hybrid Forecasting Competition (HFC)\footnote{\textbf{Funding
+  acknowledgement}: This research is based upon work supported in part
+  by the Office of the Director of National Intelligence (ODNI),
+  Intelligence Advanced Research Projects Activity (IARPA), via
+  2017-17071900005. The views and conclusions contained herein are those
+  of the authors and should not be interpreted as necessarily
+  representing the official policies, either expressed or implied, of
+  ODNI, IARPA, or the U.S. Government. The U.S. Government is authorized
+  to reproduce and distribute reprints for governmental purposes
+  notwithstanding any copyright annotation therein.} is an IARPA program
+that seeks to develop methods for hybrid geopolitical forecasting system
+that combine human and machine forecasts to answer a broad range of
+questions about economic, political, health, and other events and
+trends. The first trial period, or RCT, took place in 2018, during the
+course of which hundreds of volunteer and Amazon Mechnial Turk
+forecasters, as well as automated machine models, answered more than 150
+questions covering a broad range of issues.
 
-Each question includes 2 to 5 answer options to which forecasters must assign weights summing to 1. Performance for a single forecast is then based on an ordered multinomial Brier score.
+We worked on one of the competition teams, and specifically by
+contributing a time series forecasting module. Out of the large set of
+interesting questions one could examine with the results so far, given
+our specific focus on this project, we will try to examine in this paper
+two questions: (1) how does the accuracy of turker forecasters compare
+to volunteer forecasters, and (2) what was the impact of the machine
+forecasts on human forecaster accuracy?
 
-In addition to this basic "human forecaster" tournament, competitors were expected to implement features that would in some fashion or agument these human forecasts with machine-generated tools and forecasts.
-An explicity requirement for the latter is that they are automated systems, e.g.
-an ad-hoc hand-tuned and expert-implemented machine model to forecast on a question would not be allowed, rather it would have to be a system that can generate such a model.
-On the other hand, if a forecaster has the skills and inclination to use data and model as a forecasting tool, that is perfectly allowed.
+\section{The Hybrid Forecasting
+Competition}\label{the-hybrid-forecasting-competition}
 
-## SAGE approach
+The goal of the HFC is to find ways to optimally combine human and
+machine forecasts. For example, machine forecasts can be reliable and
+scalable, but are constrained by available data, idiosyncratic
+questions, and cold start problems when a corpus of historical data is
+not available. Human-generated forecasts on the other hand are more
+flexible, but also more costly to scale and subject to various cognitive
+biases.(e.g. Kahneman and Egan 2011).
 
-One of the distinguishing features of our team's approach was a system that could automatically associate some of the RCT-A questions with a clearly corresponding time series and display these to a forcaster.
-This is based on a data platform which automatically collects and updates data from a variety of sources, and which can associate questions, based on their title, to an appropriate transformation of a data set if it is in the platform and matches a known, pre-specified pattern or template.
-If data is found for a question, it can then be shown to users as a simple time series chart accompanying the relevant question.
+The competition is organized around Good Judgement-style forecasting
+tournaments (P. Tetlock 2005, P. E. Tetlock, Mellers, and Scoblic
+(2017)), where forecasters answer and are scored on a diverse pool of
+questions. Examples from the first trial period, RCT-A, include:
 
-Additionally, we could use the time series to generate a machine forecast based on a univariate ARIMA model, and which then would either be shown to a user, and/or submitted seperately as a standalone forecast.
-The module generating these forecasts, "basil-ts", was based on the Auto ARIMA model in @hyndman:khandakar:2008, which consists of a ARIMA-family model with an automated algorithm for determining a reasonble specific model structure, e.g.
-whether and how many differencing orders to apply, AR and MA orders, and some other parameters.
-This is wrapped in additional functionality needed to meet the automation requirements, e.g.
-recognizing how far a forecast needs to extend, data pre-processing, converting time series to answer option forecasts, updating forecasts with additional information if available, etc.
+\begin{itemize}
+\tightlist
+\item
+  What will be the long-term interest rate for South Africa (ZAF) in
+  July 2018?
+\item
+  How many deaths perpetrated by Boko Haram will the Council on Foreign
+  Relations report for June 2018?
+\item
+  What will be the daily closing spot price of Brent crude oil (USD per
+  barrel) on 31 May 2018, according to the U.S. EIA?
+\end{itemize}
 
-## Research design for RCT-A
+Each question includes 2 to 5 answer options to which forecasters must
+assign weights summing to 1. Performance for a single forecast is then
+based on an ordered multinomial Brier score.
 
-Since the scientific goal of HFC is to evaluate various hybrid forecasting techniques, the pripmary aspect of our research design for RCT-A was designed to assess what impact exposure to the time series charts and model forecasts would have on human forecaster accuracy.
-Incoming forecasters were assigned to one of three experimental conditions.
-The first group, A, served as control group and only had access to a basic version of the online platform showin the question information and tools to enter weights for the answer options.
-The second group, B, could also see the time series charts, and the third group, C, could see the chart and machine forecast.
-All forecasters, regardless of group, were forecasting on the same set of questions.
+In addition to this basic ``human forecaster'' tournament, competitors
+were expected to implement features that would in some fashion or
+agument these human forecasts with machine-generated tools and
+forecasts. An explicity requirement for the latter is that they are
+automated systems, e.g.~an ad-hoc hand-tuned and expert-implemented
+machine model to forecast on a question would not be allowed, rather it
+would have to be a system that can generate such a model. On the other
+hand, if a forecaster has the skills and inclination to use data and
+model as a forecasting tool, that is perfectly allowed.
 
-There were elements in the research design to assess other design choices but they are not relevant to the set of questions we seek to examine here, thus we will not discuss them.
+\subsection{SAGE approach}\label{sage-approach}
 
-| Condition | Treatment                 | 
-|-----------|---------------------------|
-|A          |None; control              |
-|B          |Chart                      |
-|C          |Chart and machine forecast |
+One of the distinguishing features of our team's approach was a system
+that could automatically associate some of the RCT-A questions with a
+clearly corresponding time series and display these to a forcaster. This
+is based on a data platform which automatically collects and updates
+data from a variety of sources, and which can associate questions, based
+on their title, to an appropriate transformation of a data set if it is
+in the platform and matches a known, pre-specified pattern or template.
+If data is found for a question, it can then be shown to users as a
+simple time series chart accompanying the relevant question.
 
-Table: Summary of original research design. \label{tab:original-design}
+Additionally, we could use the time series to generate a machine
+forecast based on a univariate ARIMA model, and which then would either
+be shown to a user, and/or submitted seperately as a standalone
+forecast. The module generating these forecasts, ``basil-ts'', was based
+on the Auto ARIMA model in Hyndman and Khandakar (2008), which consists
+of a ARIMA-family model with an automated algorithm for determining a
+reasonble specific model structure, e.g.~whether and how many
+differencing orders to apply, AR and MA orders, and some other
+parameters. This is wrapped in additional functionality needed to meet
+the automation requirements, e.g.~recognizing how far a forecast needs
+to extend, data pre-processing, converting time series to answer option
+forecasts, updating forecasts with additional information if available,
+etc.
 
-Comparison of groups A and B should have shown the effect of seeing a time series chart, and of groups B and C for the effect of seeing a model forecast.
-In practice, there were issues that complicated the effective design or forecaster groupings and treatments:
+\subsection{Research design for RCT-A}\label{research-design-for-rct-a}
 
-- **Amazon Mechanical Turk forecasters**.
-Due to lower than expected activity levels, turker forecasters started to be provided several weeks intot he first trial period.
-Given activity levels at that time, a decision was made to assign turkers to either condition A or C, but not B.
-This mean that the group whose treatment was to only see charts (B) consisted only of volunteers, while the other two groups contained mixed populations of volunteers and turkers, thus adding a confounding factor for assessing the impact of charts and machine models.
-- **Gaps in data and machine coverage**.
-Only about 1/3 of IFPs had time series data available, meaning that within each condition group, most questions did not have a chart nor model.
-There was also a small number of instances where data was available, but a machine forecast was not.
-One example were questions related to FluNet influenza case counts, where a change in the data source broke the ability to update chart data, which would have required models to forecast over excessive time horizons.
-Since the availability of data was related to the type of question, it is not possible to rule out the possibility that questions with data were systematically different in their difficulty from questions without data.
-- **Changes in the data and machine forecasting platforms over time**.
-Both the data platform and the machine forecasting system suffered from various bugs and related issues, especially during the earlier portions of RCT-A.
-These problems in some cases resulted in incorrectly aggregated or otherwise inaccurate data, insufficient updating which led to data in the platform falling behind source availability and thus requiring forecasts over longer time periods than neccessary, and bugs in the machine forecaster that led to no or bad forecasts.
-The quality of the data and machine forecasts displayed to some users thus varied over time and IFPs, in ways that are difficult to reconstruct in retrospective.
+Since the scientific goal of HFC is to evaluate various hybrid
+forecasting techniques, the pripmary aspect of our research design for
+RCT-A was designed to assess what impact exposure to the time series
+charts and model forecasts would have on human forecaster accuracy.
+Incoming forecasters were assigned to one of three experimental
+conditions. The first group, A, served as control group and only had
+access to a basic version of the online platform showin the question
+information and tools to enter weights for the answer options. The
+second group, B, could also see the time series charts, and the third
+group, C, could see the chart and machine forecast. All forecasters,
+regardless of group, were forecasting on the same set of questions.
 
-In respect to the first two problems, which we can quantify, the effective treatment groups were more complicated and are showin in Table \ref{tab:all-conditions}.
+There were elements in the research design to assess other design
+choices but they are not relevant to the set of questions we seek to
+examine here, thus we will not discuss them.
 
-| Group|Condition          |Forecaster |IFP_Group       |Sees chart? |Sees model? |
-|-----:|:------------------|:----------|:---------------|:-----------|:-----------|
-|     1|A: no chart        |Turker     |No TS data      |            |            |
-|     2|A: no chart        |Turker     |Chart only      |            |            |
-|     3|A: no chart        |Turker     |Chart and model |            |            |
-|     4|A: no chart        |Volunteer  |No TS data      |            |            |
-|     5|A: no chart        |Volunteer  |Chart only      |            |            |
-|     6|A: no chart        |Volunteer  |Chart and model |            |            |
-|     7|B: chart only      |Volunteer  |No TS data      |            |            |
-|     8|B: chart only      |Volunteer  |Chart only      |X           |            |
-|     9|B: chart only      |Volunteer  |Chart and model |X           |            |
-|    10|C: chart and model |Turker     |No TS data      |            |            |
-|    11|C: chart and model |Turker     |Chart only      |X           |            |
-|    12|C: chart and model |Turker     |Chart and model |X           |X           |
-|    13|C: chart and model |Volunteer  |No TS data      |            |            |
-|    14|C: chart and model |Volunteer  |Chart only      |X           |            |
-|    15|C: chart and model |Volunteer  |Chart and model |X           |X           |
-|    16|Machine            |Machine    |Chart and model |            |            |
+\begin{longtable}[]{@{}ll@{}}
+\caption{Summary of original research design.
+\label{tab:original-design}}\tabularnewline
+\toprule
+Condition & Treatment\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Condition & Treatment\tabularnewline
+\midrule
+\endhead
+A & None; control\tabularnewline
+B & Chart\tabularnewline
+C & Chart and machine forecast\tabularnewline
+\bottomrule
+\end{longtable}
 
-Table: Effective treatments by group after addition of turkers to conditions A and C. \label{tab:all-conditions}
+Comparison of groups A and B should have shown the effect of seeing a
+time series chart, and of groups B and C for the effect of seeing a
+model forecast. In practice, there were issues that complicated the
+effective design or forecaster groupings and treatments:
 
-# Design, data, and method
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Amazon Mechanical Turk forecasters}. Due to lower than
+  expected activity levels, turker forecasters started to be provided
+  several weeks intot he first trial period. Given activity levels at
+  that time, a decision was made to assign turkers to either condition A
+  or C, but not B. This mean that the group whose treatment was to only
+  see charts (B) consisted only of volunteers, while the other two
+  groups contained mixed populations of volunteers and turkers, thus
+  adding a confounding factor for assessing the impact of charts and
+  machine models.
+\item
+  \textbf{Gaps in data and machine coverage}. Only about 1/3 of IFPs had
+  time series data available, meaning that within each condition group,
+  most questions did not have a chart nor model. There was also a small
+  number of instances where data was available, but a machine forecast
+  was not. One example were questions related to FluNet influenza case
+  counts, where a change in the data source broke the ability to update
+  chart data, which would have required models to forecast over
+  excessive time horizons. Since the availability of data was related to
+  the type of question, it is not possible to rule out the possibility
+  that questions with data were systematically different in their
+  difficulty from questions without data.
+\item
+  \textbf{Changes in the data and machine forecasting platforms over
+  time}. Both the data platform and the machine forecasting system
+  suffered from various bugs and related issues, especially during the
+  earlier portions of RCT-A. These problems in some cases resulted in
+  incorrectly aggregated or otherwise inaccurate data, insufficient
+  updating which led to data in the platform falling behind source
+  availability and thus requiring forecasts over longer time periods
+  than neccessary, and bugs in the machine forecaster that led to no or
+  bad forecasts. The quality of the data and machine forecasts displayed
+  to some users thus varied over time and IFPs, in ways that are
+  difficult to reconstruct in retrospective.
+\end{itemize}
 
-## DGP
+In respect to the first two problems, which we can quantify, the
+effective treatment groups were more complicated and are showin in Table
+\ref{tab:all-conditions}.
 
-Question comes from T&E to our platform.
+\begin{longtable}[]{@{}rlllll@{}}
+\caption{Effective treatments by group after addition of turkers to
+conditions A and C. \label{tab:all-conditions}}\tabularnewline
+\toprule
+Group & Condition & Forecaster & IFP\_Group & Sees chart? & Sees
+model?\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Group & Condition & Forecaster & IFP\_Group & Sees chart? & Sees
+model?\tabularnewline
+\midrule
+\endhead
+1 & A: no chart & Turker & No TS data & &\tabularnewline
+2 & A: no chart & Turker & Chart only & &\tabularnewline
+3 & A: no chart & Turker & Chart and model & &\tabularnewline
+4 & A: no chart & Volunteer & No TS data & &\tabularnewline
+5 & A: no chart & Volunteer & Chart only & &\tabularnewline
+6 & A: no chart & Volunteer & Chart and model & &\tabularnewline
+7 & B: chart only & Volunteer & No TS data & &\tabularnewline
+8 & B: chart only & Volunteer & Chart only & X &\tabularnewline
+9 & B: chart only & Volunteer & Chart and model & X &\tabularnewline
+10 & C: chart and model & Turker & No TS data & &\tabularnewline
+11 & C: chart and model & Turker & Chart only & X &\tabularnewline
+12 & C: chart and model & Turker & Chart and model & X &
+X\tabularnewline
+13 & C: chart and model & Volunteer & No TS data & &\tabularnewline
+14 & C: chart and model & Volunteer & Chart only & X &\tabularnewline
+15 & C: chart and model & Volunteer & Chart and model & X &
+X\tabularnewline
+16 & Machine & Machine & Chart and model & &\tabularnewline
+\bottomrule
+\end{longtable}
+
+\section{Design, data, and method}\label{design-data-and-method}
+
+\subsection{DGP}\label{dgp}
+
+Question comes from T\&E to our platform.
 
@@
 
-1. Human
-  - sees question, and depending on condition the chart and machine forecast
-  - forecasts, possibly multiple times per day
-  - Mean daily Brier score
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Human
+\end{enumerate}
 
-2. Platform
-  - parses question, finds associated data
-  - sends question and data to forecaster
-  - forecaster parses question and data, TS forecast, convert to MN forecast
-  - mean daily Brier score and/or shown to user
+\begin{itemize}
+\tightlist
+\item
+  sees question, and depending on condition the chart and machine
+  forecast
+\item
+  forecasts, possibly multiple times per day
+\item
+  Mean daily Brier score
+\end{itemize}
 
-## Data
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Platform
+\end{enumerate}
 
-RCT-A lasted from March 7th to September 7th 2018.
-However, Turkers did not enter until May 2nd, and for an unrelated reason we also discard data after August 2nd.
-This leaves a total of 48,931 forecasts---7,816 from volunteer forecasters, 39,140 from Turkers, and 1,975 from machine---for 156 IFPs, of which 101 did not have TS data, 6 had TS data but not a machine forecast, and 49 had both TS data and a machine forecast.
+\begin{itemize}
+\tightlist
+\item
+  parses question, finds associated data
+\item
+  sends question and data to forecaster
+\item
+  forecaster parses question and data, TS forecast, convert to MN
+  forecast
+\item
+  mean daily Brier score and/or shown to user
+\end{itemize}
 
-# Results
+\subsection{Data}\label{data}
 
-```{r}
-knitr::opts_current$set(label = "brier-by-forecaster")
-tbl <- read_csv("../output/tables/brier-by-forecaster.csv")
-knitr::kable(tbl, format = "latex", digits = 2, booktabs = TRUE,
-             caption = "Average Brier score by forecaster group")
-```
+RCT-A lasted from March 7th to September 7th 2018. However, Turkers did
+not enter until May 2nd, and for an unrelated reason we also discard
+data after August 2nd. This leaves a total of 48,931 forecasts---7,816
+from volunteer forecasters, 39,140 from Turkers, and 1,975 from
+machine---for 156 IFPs, of which 101 did not have TS data, 6 had TS data
+but not a machine forecast, and 49 had both TS data and a machine
+forecast.
+
+\section{Results}\label{results}
+
+\begin{table}
+
+\caption{\label{tab:brier-by-forecaster}Average Brier score by forecaster group}
+\centering
+\begin{tabular}[t]{lrr}
+\toprule
+Forecaster & avg\_Brier & n\\
+\midrule
+Machine & 0.39 & 1975\\
+Turker & 0.43 & 39140\\
+Volunteer & 0.32 & 7816\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-```{r}
-options(knitr.kable.NA = '')
-knitr::opts_current$set(label = "brier-by-forecaster-condition")
-tbl <- read_csv("../output/tables/brier-by-forecaster-condition.csv")
-tbl %>%
-  spread(Condition, avg_Brier) %>%
-  knitr::kable(., format = "latex", digits = 2, booktabs = TRUE,
-             caption = "Average Brier score by forecaster group")
-```
+\begin{table}
 
-Table \ref{tab:brier-by-forecaster-ifp-group-condition} summarizes the average Brier scores for all 16 treatment/forecaster/IFP groups in Table XX.
-As it is quite unwieldy, we include it for reference and will discuss specific insights in more details below.
+\caption{\label{tab:brier-by-forecaster-condition}Average Brier score by forecaster group}
+\centering
+\begin{tabular}[t]{lrrrr}
+\toprule
+Forecaster & A: no chart & B: chart only & C: chart and model & Machine\\
+\midrule
+Machine &  &  &  & 0.39\\
+Turker & 0.44 &  & 0.43 & \\
+Volunteer & 0.30 & 0.25 & 0.37 & \\
+\bottomrule
+\end{tabular}
+\end{table}
 
-```{r brier-by-forecaster-ifp-group-condition}
-knitr::opts_current$set(label = "brier-by-forecaster-ifp-group-condition")
-tbl <- read_csv("../output/tables/brier-by-forecaster-ifp-group-condition.csv")
-knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
-             caption = "Average Brier scores for all 15 distinct human IFP/treatment groups, as well as the machine forecasts",
-             linesep = c(rep("", 5), '\\addlinespace', 
-                         rep("", 2), '\\addlinespace', 
-                         rep("", 5), '\\addlinespace'))
-```
+Table \ref{tab:brier-by-forecaster-ifp-group-condition} summarizes the
+average Brier scores for all 16 treatment/forecaster/IFP groups in Table
+XX. As it is quite unwieldy, we include it for reference and will
+discuss specific insights in more details below.
 
-## Turkers generally did worse
+\begin{table}
 
-Turkers overall had worse forecast accuracy than either the machine models or volunteers.
-To rule a spurious relationship, we can compare the accuracy in directly comparable pairs, e.g.
-rows 1 and 4 in Table \ref{tab:brier-by-forecaster-ifp-group-condition} consists of the average Brier scores for questions without TS data and in condition A treatment group for turkers and volunteers respectively, and show a difference of 0.14.
+\caption{\label{tab:brier-by-forecaster-ifp-group-condition}Average Brier scores for all 15 distinct human IFP/treatment groups, as well as the machine forecasts}
+\centering
+\begin{tabular}[t]{rlllrrr}
+\toprule
+Group & Condition & Forecaster & IFP\_Group & avg\_Brier & sd\_Brier & n\\
+\midrule
+1 & A: no chart & Turker & No TS data & 0.45 & 0.50 & 6690\\
+2 & A: no chart & Turker & Chart only & 0.55 & 0.56 & 304\\
+3 & A: no chart & Turker & Chart and model & 0.39 & 0.40 & 2907\\
+4 & A: no chart & Volunteer & No TS data & 0.31 & 0.54 & 510\\
+5 & A: no chart & Volunteer & Chart only & 0.36 & 0.46 & 18\\
+6 & A: no chart & Volunteer & Chart and model & 0.28 & 0.34 & 274\\
+\addlinespace
+7 & B: chart only & Volunteer & No TS data & 0.26 & 0.48 & 1702\\
+8 & B: chart only & Volunteer & Chart only & 0.42 & 0.42 & 57\\
+9 & B: chart only & Volunteer & Chart and model & 0.23 & 0.31 & 973\\
+\addlinespace
+10 & C: chart and model & Turker & No TS data & 0.45 & 0.50 & 19502\\
+11 & C: chart and model & Turker & Chart only & 0.56 & 0.52 & 990\\
+12 & C: chart and model & Turker & Chart and model & 0.38 & 0.40 & 8747\\
+13 & C: chart and model & Volunteer & No TS data & 0.38 & 0.60 & 3090\\
+14 & C: chart and model & Volunteer & Chart only & 0.59 & 0.57 & 73\\
+15 & C: chart and model & Volunteer & Chart and model & 0.32 & 0.40 & 1119\\
+16 & Machine & Machine & Chart and model & 0.39 & 0.36 & 1975\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-Figure \ref{fig:turkers} shows the results of pairwise comparisons of the average Brier scores for all groups relative to Turker forecasters.
-Each row in the plot corresponds to a linear model for forecasters from the corresponding condition and IFP group, and the pointranges show coefficient estimates and 95\% confidence intervals for Turkers, Machine, and Volunteer forecasters.
-The Turker values correspond to the average Brier for them in this grouping, while the Machine and Volunteer estimates indicate the *relative* performance.
-I.e.
-for them, values below zero indicate an improvment.
-Note that the x-scale is reversed, so that better is further to the right.
+\subsection{Turkers generally did
+worse}\label{turkers-generally-did-worse}
+
+Turkers overall had worse forecast accuracy than either the machine
+models or volunteers. To rule a spurious relationship, we can compare
+the accuracy in directly comparable pairs, e.g.~rows 1 and 4 in Table
+\ref{tab:brier-by-forecaster-ifp-group-condition} consists of the
+average Brier scores for questions without TS data and in condition A
+treatment group for turkers and volunteers respectively, and show a
+difference of 0.14.
+
+Figure \ref{fig:turkers} shows the results of pairwise comparisons of
+the average Brier scores for all groups relative to Turker forecasters.
+Each row in the plot corresponds to a linear model for forecasters from
+the corresponding condition and IFP group, and the pointranges show
+coefficient estimates and 95\% confidence intervals for Turkers,
+Machine, and Volunteer forecasters. The Turker values correspond to the
+average Brier for them in this grouping, while the Machine and Volunteer
+estimates indicate the \emph{relative} performance. I.e. for them,
+values below zero indicate an improvment. Note that the x-scale is
+reversed, so that better is further to the right.
 
@@
 
-In most groups volunteers have an advantage sufficient to generate p-values below 0.05, although the effects are minute in comparison to general between-forecaster differences: all models have very small adjusted R^2 values, below 0.01.
+In most groups volunteers have an advantage sufficient to generate
+p-values below 0.05, although the effects are minute in comparison to
+general between-forecaster differences: all models have very small
+adjusted R\^{}2 values, below 0.01.
 
-## Chart and model effects
+\subsection{Chart and model effects}\label{chart-and-model-effects}
 
-Teasing out the chart and model effects is a bit harder through pairwise comparison.
-Instead, we directly coded whether a forecaster saw a chart or model forecast.
-Forcasters in conditions B and C and for questions that had a chart saw a chart, and forecasters in condition C looking at a question which had a machine forecast will have seen a model forecast.
-We also leave out the machine forecasts from the data, leaving almost 47,000 forecasts, of which about 25% were made with a chart available, and 21% also had a model forecast available.
+Teasing out the chart and model effects is a bit harder through pairwise
+comparison. Instead, we directly coded whether a forecaster saw a chart
+or model forecast. Forcasters in conditions B and C and for questions
+that had a chart saw a chart, and forecasters in condition C looking at
+a question which had a machine forecast will have seen a model forecast.
+We also leave out the machine forecasts from the data, leaving almost
+47,000 forecasts, of which about 25\% were made with a chart available,
+and 21\% also had a model forecast available.
 
-We estimated two linear models to predict Brier scores for a forecast. The first has the specifiction:
+We estimated two linear models to predict Brier scores for a forecast.
+The first has the specifiction:
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-All variables are dummy terms.
-The intercept $\alpha$ corresponds to the average Brier score for a Turker forecasting on an IFP that did not have chartable data, and who neither saw a chart nor machine forecast.
+All variables are dummy terms. The intercept \(\alpha\) corresponds to
+the average Brier score for a Turker forecasting on an IFP that did not
+have chartable data, and who neither saw a chart nor machine forecast.
 
-The second model includes random intercepts for IFP to account for the possibility that seeings charts or models may make users differentially willing to forecast on harder IFPs, i.e.
-those with higher average Brier scores.
-Otherwise the specification of covariates is the same.
+The second model includes random intercepts for IFP to account for the
+possibility that seeings charts or models may make users differentially
+willing to forecast on harder IFPs, i.e.~those with higher average Brier
+scores. Otherwise the specification of covariates is the same.
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Figure \ref{fig:model-chart-and-model-effects} plots the coefficient estimates from both models.
-The results are consistent.
-Seeing a chart slightly helps, seeing a model slightly hurts forecast performance.
-However, the effects are small and not as important as whether a forecasters was a volunteer or turker.
-In addition, the models themselves capture only a small portion of the variation in Brier scores overall.
-Table \ref{tab:model-summary-stats} shows $R^2$ statistics for both models.
-The covariates account for less than 0.01 of the variation in Brier scores; the inclusion of random IFP intercepts in model 2 increases this to 0.28 (conditional $R^2$), but the fraction explained by the covariates together is still 0.01 (marginal $R^2$).
-Most of the variation remains either between forecasters or due to other random factors.
+Figure \ref{fig:model-chart-and-model-effects} plots the coefficient
+estimates from both models. The results are consistent. Seeing a chart
+slightly helps, seeing a model slightly hurts forecast performance.
+However, the effects are small and not as important as whether a
+forecasters was a volunteer or turker. In addition, the models
+themselves capture only a small portion of the variation in Brier scores
+overall. Table \ref{tab:model-summary-stats} shows \(R^2\) statistics
+for both models. The covariates account for less than 0.01 of the
+variation in Brier scores; the inclusion of random IFP intercepts in
+model 2 increases this to 0.28 (conditional \(R^2\)), but the fraction
+explained by the covariates together is still 0.01 (marginal \(R^2\)).
+Most of the variation remains either between forecasters or due to other
+random factors.
 
-Thus, while we find that exposure to charts and model forecasts has a small impact on forecast quality, the effect sizes are small in comparison to other systematic factors in the model, and small relative to unsystematic and random factors not in the model.
+Thus, while we find that exposure to charts and model forecasts has a
+small impact on forecast quality, the effect sizes are small in
+comparison to other systematic factors in the model, and small relative
+to unsystematic and random factors not in the model.
 
@@
 
-```{r model-summary-stats, result='asis'}
-knitr::opts_current$set(label = "model-summary-stats")
-tbl <- read_csv("../output/tables/model-summary-stats.csv")
-knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
-             caption = "Model summary statistics") %>%
-  gsub("R\\\\textasciicircum\\{\\}2", "$R^2$", .) %>%
-  # cut off the integer trailing digits
-  gsub("\\.00", "", .)
-```
+\begin{table}
 
-## Models generally were disappointing, but did they do well in certain areas?
+\caption{\label{tab:model-summary-stats}Model summary statistics}
+\centering
+\begin{tabular}[t]{llr}
+\toprule
+Model & Statistic & Value\\
+\midrule
+Model 1: Linear model & N & 46956\\
+ & $R^2$ & 0.01\\
+ & Adj. $R^2$ & 0.01\\
+Model 2: Linear model with IFP random intercepts & N & 46956\\
+ & Marginal $R^2$ & 0.01\\
+ & Conditional $R^2$ & 0.28\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-Although the machine models overall underperformed, it is possible that they performed well relative to human forecasters in certain areas.
-To examine this we looked at all forecasts for questions that had a machine forecast.
-This leaves us with 15,995 forecasts, approximately 2,000 of which were machine and volunteer forecasts, respectively, and the remaining 12,000 from turkers.
+\subsection{Models generally were disappointing, but did they do well in
+certain
+areas?}\label{models-generally-were-disappointing-but-did-they-do-well-in-certain-areas}
 
-To estimate the relative performance of different forecasters, we again estimated a simple linear model, similar to those used for the turker pariwise comparisons. The model has the general form:
+Although the machine models overall underperformed, it is possible that
+they performed well relative to human forecasters in certain areas. To
+examine this we looked at all forecasts for questions that had a machine
+forecast. This leaves us with 15,995 forecasts, approximately 2,000 of
+which were machine and volunteer forecasts, respectively, and the
+remaining 12,000 from turkers.
 
-$$
+To estimate the relative performance of different forecasters, we again
+estimated a simple linear model, similar to those used for the turker
+pariwise comparisons. The model has the general form:
+
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Data source is a discrete variable with $i=14$ levels for the different data sources reflected in the set of IFPs.
-Based on previous exploratory analysis, we separated one data source, ACLED questions, by whether a question was binary with two answer options ("yes"/"no"), or not (3 or 5 options).
-Forecaster is also a discrete variable with 3 levels, but we leave out "Machine" as reference category, giving us $j=2$ different levels.
-Thus, each of the $\beta$ coefficients corresponds to the average Brier score for machine forecasts for that group of IFPs, and each $\beta'_i$ coefficient is the relative difference of the average volunteer and turker Brier score compared to the corresponding baseline in $\beta_i$.
+Data source is a discrete variable with \(i=14\) levels for the
+different data sources reflected in the set of IFPs. Based on previous
+exploratory analysis, we separated one data source, ACLED questions, by
+whether a question was binary with two answer options (``yes''/``no''),
+or not (3 or 5 options). Forecaster is also a discrete variable with 3
+levels, but we leave out ``Machine'' as reference category, giving us
+\(j=2\) different levels. Thus, each of the \(\beta\) coefficients
+corresponds to the average Brier score for machine forecasts for that
+group of IFPs, and each \(\beta'_i\) coefficient is the relative
+difference of the average volunteer and turker Brier score compared to
+the corresponding baseline in \(\beta_i\).
 
@@
 
-Figure \ref{fig:model-machine-by-data-source} shows the coefficient estimates.
-The left panel shows the baseline machine forecast performance.
-The right panel shows the relative difference in average Brier scores for turker and volunteer forecasts; points to the left of the dotted 0 line indicate that volunteer or turker forecasters on average outperformed the machine forecasts.
+Figure \ref{fig:model-machine-by-data-source} shows the coefficient
+estimates. The left panel shows the baseline machine forecast
+performance. The right panel shows the relative difference in average
+Brier scores for turker and volunteer forecasts; points to the left of
+the dotted 0 line indicate that volunteer or turker forecasters on
+average outperformed the machine forecasts.
 
-The machine forecasts did better than human forecasts on ACLED 5-option, Boko Haram killings ("nigeria-security"), and hacking ("privacyrights") questions.
-All three data sources consist of count data derived from original event datasets, i.e.
-they require data aggregation before one can see a time series relevant to the question at hand, which might be a factor in the relative performance on these questions.
-On the other hand, human forecasters consistently did better on ACLED 2-option, food price index ("FAO-FPI"), financial ("fred"), and oil production ("opec") questions.
-The results for ACLED 2-option questions might be influenced by the particular mechanism by which time series forecasts were converted to answer option probabilities, which may explain the inconsistency with ACLED 5-option questions.
+The machine forecasts did better than human forecasts on ACLED 5-option,
+Boko Haram killings (``nigeria-security''), and hacking
+(``privacyrights'') questions. All three data sources consist of count
+data derived from original event datasets, i.e.~they require data
+aggregation before one can see a time series relevant to the question at
+hand, which might be a factor in the relative performance on these
+questions. On the other hand, human forecasters consistently did better
+on ACLED 2-option, food price index (``FAO-FPI''), financial (``fred''),
+and oil production (``opec'') questions. The results for ACLED 2-option
+questions might be influenced by the particular mechanism by which time
+series forecasts were converted to answer option probabilities, which
+may explain the inconsistency with ACLED 5-option questions.
 
-There are some hard to quantify factors that may explain some of the variation in human to machine performance:
+There are some hard to quantify factors that may explain some of the
+variation in human to machine performance:
 
-1.
-As mentioned, some data sources (ACLED, privacyrights, nigeria-security) required relatively complicated data aggregation to derive usable count series, and which therefore may not have been accessible to all human forecasters.
-2.
-Not all data sources are equally easy to automatically download and update, and at least in some of the data sources it was the case that models had to forecast over much longer time horizons due to outdated data, while a human forecaster would have been able to access more recent data.
-3.
-Due to obstacles in data acquisition, in at least one case--OPEC oil production rates--a secondary and less accurate source was substituted.
-OPEC production values are delivered in monthly PDF reports that are difficult to automatically extract data from, and the alternative data source in some instances significantly deviated from the OPEC values.
-Thus models for these questions would have been forecasting with incorrect input data.
-4.
-For questions in which the data source is at a monthly (or higher) resolution, humans have an advantage over univariate time series models in that they can incorporate sub-monthly information into their forecast.
-For many of these questions, if they were open for a period of one or two months, there would in fact have only been one or two distinct time series forecasts (barring practical live system issues and bugs).
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  As mentioned, some data sources (ACLED, privacyrights,
+  nigeria-security) required relatively complicated data aggregation to
+  derive usable count series, and which therefore may not have been
+  accessible to all human forecasters.
+\item
+  Not all data sources are equally easy to automatically download and
+  update, and at least in some of the data sources it was the case that
+  models had to forecast over much longer time horizons due to outdated
+  data, while a human forecaster would have been able to access more
+  recent data.
+\item
+  Due to obstacles in data acquisition, in at least one case--OPEC oil
+  production rates--a secondary and less accurate source was
+  substituted. OPEC production values are delivered in monthly PDF
+  reports that are difficult to automatically extract data from, and the
+  alternative data source in some instances significantly deviated from
+  the OPEC values. Thus models for these questions would have been
+  forecasting with incorrect input data.
+\item
+  For questions in which the data source is at a monthly (or higher)
+  resolution, humans have an advantage over univariate time series
+  models in that they can incorporate sub-monthly information into their
+  forecast. For many of these questions, if they were open for a period
+  of one or two months, there would in fact have only been one or two
+  distinct time series forecasts (barring practical live system issues
+  and bugs).
+\end{enumerate}
 
-A caveat to these trends is that the number of IFPs and forecasts for each data source are small and the results may thus be due to random variation. Results from future RCTs should test this.
+A caveat to these trends is that the number of IFPs and forecasts for
+each data source are small and the results may thus be due to random
+variation. Results from future RCTs should test this.
 
-## Why did condition B volunteers do so well?
+\subsection{Why did condition B volunteers do so
+well?}\label{why-did-condition-b-volunteers-do-so-well}
 
-Condition B volunteers had the best overall performance out of the 16 groups shown in Table \ref{tab:brier-by-forecaster-ifp-group-condition}.
-Considering that the machine forecasts were relatively bad, one possible explanation is that the charts overall had a positive impact on human forecast accuracy, while seeing machine forecasts had a negative impact.
+Condition B volunteers had the best overall performance out of the 16
+groups shown in Table \ref{tab:brier-by-forecaster-ifp-group-condition}.
+Considering that the machine forecasts were relatively bad, one possible
+explanation is that the charts overall had a positive impact on human
+forecast accuracy, while seeing machine forecasts had a negative impact.
 
-However, there are several additional patterns that are either inconsistent with this simple explanation or at least suggest the need for a more complex alternative.
-The first is that condition B forecasters did not only do better than other groups on questions that had a chart, but also on questions that did not have a chart.
-Table \ref{tab:b-advantage} shows the relevant rows from the average Brier scores for all forecaster groups in Table \ref{tab:brier-by-forecaster-ifp-group-condition}.
-With the partial exception of the few chart-only questions, condition B forecasters' advantage on questions without a chart is as pronounced as their advantage in questions that did have a chart.
-Considering that there were two to three times as many forecasts on questions without chartable data, most of the condition B volunteer advantage is driven by doing better on questions that did not have a chart.
+However, there are several additional patterns that are either
+inconsistent with this simple explanation or at least suggest the need
+for a more complex alternative. The first is that condition B
+forecasters did not only do better than other groups on questions that
+had a chart, but also on questions that did not have a chart. Table
+\ref{tab:b-advantage} shows the relevant rows from the average Brier
+scores for all forecaster groups in Table
+\ref{tab:brier-by-forecaster-ifp-group-condition}. With the partial
+exception of the few chart-only questions, condition B forecasters'
+advantage on questions without a chart is as pronounced as their
+advantage in questions that did have a chart. Considering that there
+were two to three times as many forecasts on questions without chartable
+data, most of the condition B volunteer advantage is driven by doing
+better on questions that did not have a chart.
 
-```{r}
-knitr::opts_current$set(label = "b-advantage")
-tbl <- read_csv("../output/tables/volunteer-brier-by-condition-ifp-group.csv")
-knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
-             caption = "Condition B advantage",
-             linesep = c(rep("", 2), '\\addlinespace')) 
-```
+\begin{table}
 
-We could exlain this if there is, in addition to the direct beneficial effect of seeing a chart, but not a potentially misleading machine forecast, a spillover effect whereby exposure to charts also improves a volunteer forecaster's ability to forecast on other questions without charts.
+\caption{\label{tab:b-advantage}Condition B advantage}
+\centering
+\begin{tabular}[t]{rllrrr}
+\toprule
+Group & Condition & IFP\_Group & avg\_Brier & sd\_Brier & n\\
+\midrule
+7 & B: chart only & No TS data & 0.26 & 0.48 & 1702\\
+4 & A: no chart & No TS data & 0.31 & 0.54 & 510\\
+13 & C: chart and model & No TS data & 0.38 & 0.60 & 3090\\
+\addlinespace
+8 & B: chart only & Chart only & 0.42 & 0.42 & 57\\
+5 & A: no chart & Chart only & 0.36 & 0.46 & 18\\
+14 & C: chart and model & Chart only & 0.59 & 0.57 & 73\\
+\addlinespace
+9 & B: chart only & Chart and model & 0.23 & 0.31 & 973\\
+6 & A: no chart & Chart and model & 0.28 & 0.34 & 274\\
+15 & C: chart and model & Chart and model & 0.32 & 0.40 & 1119\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+We could exlain this if there is, in addition to the direct beneficial
+effect of seeing a chart, but not a potentially misleading machine
+forecast, a spillover effect whereby exposure to charts also improves a
+volunteer forecaster's ability to forecast on other questions without
+charts.
 
@@
 
-**There appears to be no chart spillover effect for condition C human and volunteer forecasters.** Tabe \ref{tab:brier-by-forecaster-ifp-group-condition} shows that condition C volunteers did worse on questions without time-series data than the condition A control group.
-Condition C turkers also did not do better than their condition A control group counterparts.
-Unlike condition B forecasters, they were exposed to machine forecasts as well, so maybe there is an equal but opposite negative machine spillover effect.
+\textbf{There appears to be no chart spillover effect for condition C
+human and volunteer forecasters.} Tabe
+\ref{tab:brier-by-forecaster-ifp-group-condition} shows that condition C
+volunteers did worse on questions without time-series data than the
+condition A control group. Condition C turkers also did not do better
+than their condition A control group counterparts. Unlike condition B
+forecasters, they were exposed to machine forecasts as well, so maybe
+there is an equal but opposite negative machine spillover effect.
 
-**Condition B volunteers who forecast a lot do no better on non-chart questions than novices.** Figure \ref{fig:cond-b-volunteers-n-vs-brier} show average Brier scores as a function of the total number of forecasts a user has.
-The data consist of only forecasts from condition B volunteers.
-Each point is the average Brier score for a user's forecasts on the set of questions that did (blue) or did not have a chart (red).
-Novice users tend to be as accurate as intensive users on non-chart questions.
-This implies that the condition B volunteer advantage on non-chart questions is not related to how active a user was.
-If there is a spillover effect, even very inactive users seem to somehow benefit from it.
+\textbf{Condition B volunteers who forecast a lot do no better on
+non-chart questions than novices.} Figure
+\ref{fig:cond-b-volunteers-n-vs-brier} show average Brier scores as a
+function of the total number of forecasts a user has. The data consist
+of only forecasts from condition B volunteers. Each point is the average
+Brier score for a user's forecasts on the set of questions that did
+(blue) or did not have a chart (red). Novice users tend to be as
+accurate as intensive users on non-chart questions. This implies that
+the condition B volunteer advantage on non-chart questions is not
+related to how active a user was. If there is a spillover effect, even
+very inactive users seem to somehow benefit from it.
 
@@
 
-**Condition C volunteers do well on questions where we know that data displayed was wrong.** There were several instances where we know that the data displayed in a chart were inaccurate.
-Figure \ref{fig:iraq} shows the first example, oil production figures for Iraq.
-The main chart shows several teim series depicting oil production in Iraq, according to different sources.
-The canonical data are contained in monthly OPEC oil market reports, each of which contains production figures for the past 3 months.
-Those are shown in the sequence of short time series on the mid-right part of the main chart--the figures are somewhat inconsistent between reports, but this is another peculiarity.
-Because this data is difficult to obtain programatically, an alternative data source is used instead.
-This is what would have been shown in the charts, and what informed the machine forecast.
-Obviously it is incorrect data.
-The dotted lines show the values separating the answer options for this question.
-The chart placed the production values in an entirely incorrect answer category.
+\textbf{Condition C volunteers do well on questions where we know that
+data displayed was wrong.} There were several instances where we know
+that the data displayed in a chart were inaccurate. Figure
+\ref{fig:iraq} shows the first example, oil production figures for Iraq.
+The main chart shows several teim series depicting oil production in
+Iraq, according to different sources. The canonical data are contained
+in monthly OPEC oil market reports, each of which contains production
+figures for the past 3 months. Those are shown in the sequence of short
+time series on the mid-right part of the main chart--the figures are
+somewhat inconsistent between reports, but this is another peculiarity.
+Because this data is difficult to obtain programatically, an alternative
+data source is used instead. This is what would have been shown in the
+charts, and what informed the machine forecast. Obviously it is
+incorrect data. The dotted lines show the values separating the answer
+options for this question. The chart placed the production values in an
+entirely incorrect answer category.
 
@@
 
-Curiously, the condition B volunteers performed better on this question than any other forecaster.
-The number of forecasts here are low, and only the last three groups, starting from the machine forecasts, have pairwise statistically significant differences in average Brier score from the condition B volunteers.
+Curiously, the condition B volunteers performed better on this question
+than any other forecaster. The number of forecasts here are low, and
+only the last three groups, starting from the machine forecasts, have
+pairwise statistically significant differences in average Brier score
+from the condition B volunteers.
 
-|Forecaster |condition | avg_Brier|  n|
-|:----------|:---------|---------:|--:|
-|Volunteer  |b         |      0.07|  5|
-|Volunteer  |a         |      0.11|  5|
-|Turker     |c         |      0.18| 87|
-|Machine    |n/a       |      0.25| 14|
-|Turker     |a         |      0.35| 23|
-|Volunteer  |c         |      0.45| 21|
+\begin{longtable}[]{@{}llrr@{}}
+\toprule
+Forecaster & condition & avg\_Brier & n\tabularnewline
+\midrule
+\endhead
+Volunteer & b & 0.07 & 5\tabularnewline
+Volunteer & a & 0.11 & 5\tabularnewline
+Turker & c & 0.18 & 87\tabularnewline
+Machine & n/a & 0.25 & 14\tabularnewline
+Turker & a & 0.35 & 23\tabularnewline
+Volunteer & c & 0.45 & 21\tabularnewline
+\bottomrule
+\end{longtable}
 
-Another example with incorrect data is an early ICEWS question that fell outside the common period and thus did not have forecasts from turkers.
-Condition B volunteers also outperformed all other groups on this question, although the pairwise differences with the other groups are not statistically significant.
+Another example with incorrect data is an early ICEWS question that fell
+outside the common period and thus did not have forecasts from turkers.
+Condition B volunteers also outperformed all other groups on this
+question, although the pairwise differences with the other groups are
+not statistically significant.
 
-|Forecaster |condition | avg_Brier|  n|
-|:----------|:---------|---------:|--:|
-|Volunteer  |b         |      0.92| 14|
-|Machine    |n/a       |      1.00| 18|
-|Volunteer  |a         |      1.11|  7|
-|Volunteer  |c         |      1.11| 25|
+\begin{longtable}[]{@{}llrr@{}}
+\toprule
+Forecaster & condition & avg\_Brier & n\tabularnewline
+\midrule
+\endhead
+Volunteer & b & 0.92 & 14\tabularnewline
+Machine & n/a & 1.00 & 18\tabularnewline
+Volunteer & a & 1.11 & 7\tabularnewline
+Volunteer & c & 1.11 & 25\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
-**Even when the machine predictions were good, condition B volunteers did better than their C counterparts.** If bad machine predictions influence human forecasts and are the reason why the beneficial chart effecgt is only apparent in condition B forecasters, and not condition C forecasters, then it should also be the case that condition C forecasters do well on those questions where the machine forecasts were good.
-This is no the case.
-The table below shows the subset of forecasts on questions that had a machine forecast (and thus chart) and where the machine forecast average Brier was lower than the average Brier score for the forecasts from all other forecaster $\times$ condition groups.
-Condition B volunteers beat their condition C counterparts even when the machine model should have improved their forecasts because they happened to be accurate.
+\textbf{Even when the machine predictions were good, condition B
+volunteers did better than their C counterparts.} If bad machine
+predictions influence human forecasts and are the reason why the
+beneficial chart effecgt is only apparent in condition B forecasters,
+and not condition C forecasters, then it should also be the case that
+condition C forecasters do well on those questions where the machine
+forecasts were good. This is no the case. The table below shows the
+subset of forecasts on questions that had a machine forecast (and thus
+chart) and where the machine forecast average Brier was lower than the
+average Brier score for the forecasts from all other forecaster
+\(\times\) condition groups. Condition B volunteers beat their condition
+C counterparts even when the machine model should have improved their
+forecasts because they happened to be accurate.
 
-|Forecaster |Condition          | avg_Brier| n_fcasts|
-|:----------|:------------------|---------:|--------:|
-|Machine    |Machine            |      0.39|     1975|
-|Turker     |A: no chart        |      0.39|     2810|
-|Turker     |C: chart and model |      0.38|     8482|
-|Volunteer  |A: no chart        |      0.28|      267|
-|Volunteer  |B: chart only      |      0.23|      958|
-|Volunteer  |C: chart and model |      0.32|     1094|
+\begin{longtable}[]{@{}llrr@{}}
+\toprule
+Forecaster & Condition & avg\_Brier & n\_fcasts\tabularnewline
+\midrule
+\endhead
+Machine & Machine & 0.39 & 1975\tabularnewline
+Turker & A: no chart & 0.39 & 2810\tabularnewline
+Turker & C: chart and model & 0.38 & 8482\tabularnewline
+Volunteer & A: no chart & 0.28 & 267\tabularnewline
+Volunteer & B: chart only & 0.23 & 958\tabularnewline
+Volunteer & C: chart and model & 0.32 & 1094\tabularnewline
+\bottomrule
+\end{longtable}
 
-One possibility that we should not discard with these apparent inconsistencies is that we are digging in noise.
-All of the effects---for turkers versus other forecasts, for the disappointing performance of the machine forecasts, for conditions or different groupings of IFPs---pale in comparison to differences in accuracy between individual forecasters and between forecasts for different IFPs.
-But on the other hand there are several inconsistencies so it also is a stretch to write all of them off to random coincidence.
-Condition B volunteers clearly did better than all other forecasters/condition groups, but it is not clear why.
+One possibility that we should not discard with these apparent
+inconsistencies is that we are digging in noise. All of the
+effects---for turkers versus other forecasts, for the disappointing
+performance of the machine forecasts, for conditions or different
+groupings of IFPs---pale in comparison to differences in accuracy
+between individual forecasters and between forecasts for different IFPs.
+But on the other hand there are several inconsistencies so it also is a
+stretch to write all of them off to random coincidence. Condition B
+volunteers clearly did better than all other forecasters/condition
+groups, but it is not clear why.
 
-# Conclusion
+\section{Conclusion}\label{conclusion}
 
@@
 
-Many of the questions involve monthly time series but are only open for periods spanning a single or few months.
-Some of the advantage of the human forecasters may thus have been due to operating on a sub-monthly time scale.
+Many of the questions involve monthly time series but are only open for
+periods spanning a single or few months. Some of the advantage of the
+human forecasters may thus have been due to operating on a sub-monthly
+time scale.
 
-We also know that model performance suffered from issues related to automation.
-Were we able to run the RCT-A questions with the current system and with a data platform that optimally ingests data sources on the bleeding edge of their own updates, our forecasts would have been ...
+We also know that model performance suffered from issues related to
+automation. Were we able to run the RCT-A questions with the current
+system and with a data platform that optimally ingests data sources on
+the bleeding edge of their own updates, our forecasts would have been
+\ldots{}
 
-**[check with updated results from test bed what the performance gain would have been, i think the site visit stats were not accurate anymore]**
+\textbf{{[}check with updated results from test bed what the performance
+gain would have been, i think the site visit stats were not accurate
+anymore{]}}
 
-# Additional information
+\section{Additional information}\label{additional-information}
 
-|Data_source      | N_IFPs| Machine| Turker| Volunteer|
-|:----------------|------:|-------:|------:|---------:|
-|ACLED 2op        |      7|   0.611|  0.533|     0.396|
-|ACLED 5op        |      4|   0.371|  0.554|     0.790|
-|earthquakes      |      2|   0.285|  0.207|     0.199|
-|eia              |      5|   0.184|  0.204|     0.145|
-|fao-fpi          |      3|   0.386|  0.220|     0.121|
-|fred             |      2|   0.580|  0.488|     0.290|
-|gold             |      3|   0.184|  0.274|     0.171|
-|imf              |      1|   0.222|  0.252|     0.325|
-|nigeria-security |      2|   0.355|  0.504|     0.648|
-|oecd             |      7|   0.430|  0.534|     0.356|
-|opec             |      5|   0.283|  0.222|     0.209|
-|other            |      2|   0.416|  0.574|     0.188|
-|privacyrights    |      2|   0.231|  0.321|     0.517|
-|stocks           |      4|   0.204|  0.204|     0.223|
+\begin{longtable}[]{@{}lrrrr@{}}
+\caption{Average Brier scores for machine and other forecasts by
+question data source.}\tabularnewline
+\toprule
+Data\_source & N\_IFPs & Machine & Turker & Volunteer\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Data\_source & N\_IFPs & Machine & Turker & Volunteer\tabularnewline
+\midrule
+\endhead
+ACLED 2op & 7 & 0.611 & 0.533 & 0.396\tabularnewline
+ACLED 5op & 4 & 0.371 & 0.554 & 0.790\tabularnewline
+earthquakes & 2 & 0.285 & 0.207 & 0.199\tabularnewline
+eia & 5 & 0.184 & 0.204 & 0.145\tabularnewline
+fao-fpi & 3 & 0.386 & 0.220 & 0.121\tabularnewline
+fred & 2 & 0.580 & 0.488 & 0.290\tabularnewline
+gold & 3 & 0.184 & 0.274 & 0.171\tabularnewline
+imf & 1 & 0.222 & 0.252 & 0.325\tabularnewline
+nigeria-security & 2 & 0.355 & 0.504 & 0.648\tabularnewline
+oecd & 7 & 0.430 & 0.534 & 0.356\tabularnewline
+opec & 5 & 0.283 & 0.222 & 0.209\tabularnewline
+other & 2 & 0.416 & 0.574 & 0.188\tabularnewline
+privacyrights & 2 & 0.231 & 0.321 & 0.517\tabularnewline
+stocks & 4 & 0.204 & 0.204 & 0.223\tabularnewline
+\bottomrule
+\end{longtable}
 
-Table: Average Brier scores for machine and other forecasts by question data source.
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
-# References
+\hypertarget{refs}{}
+\hypertarget{ref-hyndman:khandakar:2008}{}
+Hyndman, Rob J., and Yeasmin Khandakar. 2008. ``Automatic Time Series
+for Forecasting: The Forecast Package for R.'' \emph{Journal of
+Statistical Software} 27 (3). Monash University, Department of
+Econometrics; Business Statistics.
+
+\hypertarget{ref-kahneman:2011}{}
+Kahneman, Daniel, and Patrick Egan. 2011. \emph{Thinking, Fast and
+Slow}. Vol. 1. Farrar, Straus; Giroux New York.
+
+\hypertarget{ref-tetlock:2005}{}
+Tetlock, Philip. 2005. \emph{Expert Political Judgment: How Good Is It?
+How Can We Know?} Princeton, NJ: Princeton University Press.
+
+\hypertarget{ref-tetlock:etal:2017}{}
+Tetlock, Philip E, Barbara A Mellers, and J Peter Scoblic. 2017.
+``Bringing Probability Judgments into Policy Debates via Forecasting
+Tournaments.'' \emph{Science} 355 (6324). American Association for the
+Advancement of Science: 481--83.
+
+\end{document}

### 2018-12-05 R/analysis.R

@@ -35,7 +35,7 @@ df1 <- machine_fcasts %>%
          sees_chart = NA, sees_model = NA)
 df2 <- user_fcasts %>%
   mutate(Forecaster = ifelse(turker, "Turker", "Volunteer")) %>%
-  select(ifp_id, date, has_historic_data, has_arima, needs_data_agg,
+  select(ifp_id, user_id, date, has_historic_data, has_arima, needs_data_agg,
          num_options, p_correct, brier, Forecaster, condition, sub_condition_1,
          sees_chart, sees_model)
 all_fcasts <- bind_rows(df1, df2) %>%
@@ -104,9 +104,14 @@ brier_by_forecaster_ifp_group_condition <- common_period %>%
   arrange(Condition, Forecaster, IFP_Group) %>%
   mutate(Group = 1:n()) %>%
   select(Group, Condition, Forecaster, IFP_Group, everything()) %>%
-  write_csv("output/tables/brier-by-forecast-ifp-group-condition.csv")
+  write_csv("output/tables/brier-by-forecaster-ifp-group-condition.csv")
 brier_by_forecaster_ifp_group_condition %>%
   knitr::kable(digits = 2)
+
+caption = paste0(
+  create_label('tab:', opts_current$get('label'), latex = (format == 'latex')),
+  caption
+)
 
 # Design table
 brier_by_forecaster_ifp_group_condition %>%
@@ -135,11 +140,14 @@ brier_by_ifp_group
 
 turker_pairs <- common_period %>% 
   filter(condition!="b") %>%
-  mutate(Condition = fct_recode(Condition, `C: chart and model` = "Machine")) %>%
+  mutate(Condition = fct_recode(Condition, `C: chart and model` = "Machine"),
+         Condition = fct_relabel(Condition, ~ paste0("Condition ", .x))) %>%
   mutate(Forecaster = factor(Forecaster, levels = c("Turker", "Volunteer", "Machine"))) %>%
+  mutate(IFP_Group = factor(IFP_Group),
+         IFP_Group = fct_relabel(IFP_Group, ~ paste0("IFP Group: ", .x))) %>%
   unite(Group, Condition, IFP_Group, sep = "\n")
 group_all <- turker_pairs %>%
-  mutate(Group = "All") 
+  mutate(Group = "All conditions and IFPs") 
 turker_pairs <- bind_rows(turker_pairs, group_all) %>%
   group_by(Group) %>%
   nest(Forecaster, brier) %>%
[…]

### 2018-12-05 R/opec-iraq-example.R

@@ -1,0 +1,60 @@
+#
+#   Iraq oil production data divergence example
+#
+
+library("tidyverse")
+library("cowplot")
+
+df <- read_csv("input-data/iraq-oil-production.csv")
+
+p1 <- df %>%
+  filter(date >= as.Date("2018-06-01") %m-% months(120)) %>%
+  ggplot(aes(x = date, y = value, group = source)) +
+  geom_line(size = 1) +
+  geom_hline(yintercept = cutpoints, lty = 3, size = .3) +
+  theme_void() +
+  theme(plot.background = element_rect(color = "black", fill = "white")) +
+  labs(x = "", y = "") +
+  annotate("rect", xmin = as.Date("2017-01-01"), xmax = as.Date("2018-06-30"), 
+           ymin = 4100, ymax = 4700, fill = NA, color = "gray50", size = 1)
+
+p2 <- df %>% 
+  filter(date > "2017-01-01") %>%
+  ggplot(aes(x = date, y = value, colour = source)) +
+  geom_point() +
+  geom_line() +
+  geom_hline(yintercept = cutpoints, lty = 3) +
+  theme_minimal() +
+  labs(x = "Month", y = "Oil tb/d") +
+  scale_colour_discrete("Source") +
+  theme(legend.position = "top")
+
+ggdraw() +
+  draw_plot(p2, 0, 0, 1, 1) +
+  draw_plot(p1, 0.1, 0.13, .32, .2)
+ggsave(file = "output/figures/opec-iraq-divergence.png", height = 5, width = 8)
+
+
+plot.mpg <- ggplot(mpg, aes(x = cty, y = hwy, colour = factor(cyl))) + 
+  geom_point(size=2.5) + theme_cowplot()
+plot.mpg
+plot.diamonds <- ggplot(diamonds, aes(clarity, fill = cut)) + geom_bar() +
+  theme_cowplot() + theme(axis.text.x = element_text(angle=70, vjust=0.5, hjust = 0.9))
+plot.diamonds
+
+ggdraw() +
+  draw_plot(
+    plot.diamonds + theme(legend.justification = "bottom"),
+    0, 0, 1, 1
+  ) +
+  draw_plot(
+    plot.mpg + scale_color_viridis_d() + 
+      theme(legend.justification = "top", plot.background = element_rect(fill = "white")),
+    0.5, 0.54, 0.5, 0.4
+  ) +
+  draw_plot_label(
[…]

### 2018-12-05 R/opec-reports.R

@@ -1,0 +1,27 @@
+#
+# Rscript opec-reports.R
+
+library("tidyverse")
+library("readxl")
+library("hrbrthemes")
+library("here")
+library("cowplot")
+
+setwd(here::here("data-modules/opec"))
+
+opec <- read_xlsx("data/opec_reports.xlsx") %>%
+  mutate(Report = as.Date(Report),
+         Date = as.Date(Date))
+
+opec %>% group_by(Country, Date) %>%
+  summarize(Crude_oil_production_secondary_sources_tbd = mean(Crude_oil_production_secondary_sources_tbd)) %>%
+  ungroup() %>%
+  print(n = 100)
+
+ggplot(opec, aes(x = Date, y = Crude_oil_production_secondary_sources_tbd)) +
+  facet_wrap(~ Country, scales = "free_y") +
+  geom_line(aes(color = factor(Report))) +
+  scale_color_discrete(guide = FALSE) +
+  theme_ipsum_rc()
+
+ggsave("opec-divergence.png", height = 8, width = 15)

### 2018-12-07 paper/paper.Rmd
finish second draft; sent off for IARPA review

@@
 ---
-title: "Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
-author: "Andreas Beger and Michael D. Ward"
+title: "Assessing Amazon Turker and automated machine forecasts in the Hybrid Forecasting Competition^[**Funding acknowledgement**: This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
+The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S.
+Government.
+The U.S.
+Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.\\newline \\indent Prepared for the 7th Annual Asian Political Methodology Conference, 5-6 January 2019, Kyoto, Japan]"
+author: "Andreas Beger^[Predictive Heuristics, adbeger@gmail.com] and Michael D. Ward^[Predictive Heuristics, michael.don.ward@gmail.com]"
 date: "`r Sys.Date()`"
@@
     latex_engine: xelatex
-    keep_tex: true
 abstract: "\\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project to develop hybrid geopolitical forecasting systems that combine human- and machine-generated forecasts.
 The first project trial period took place in 2018, during the course of which volunteer participants, Amazon Mechanical Turkers, and an automated time series forecasting module we developed provided forecasts for more than 150 questions covering a diverse set of topics.
-We investigate two questions: (1) how does the performance of turker forecasters compare to volunteers, and (2) what impact did access to the machine forecasts have on forecaster accuracy?
+We assess several aspects of the accuracy of human and machine forecasts.
 \\newline \\newline
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+options(knitr.kable.NA = '')
 
@@
 
-The Hybrid Forecasting Competition (HFC)^[**Funding acknowledgement**: This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
-The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S.
-Government.
-The U.S.
-Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.] is an IARPA program that seeks to develop methods for hybrid geopolitical forecasting system that combine human and machine forecasts to answer a broad range of questions about economic, political, health, and other events and trends.
-The first trial period, or RCT, took place in 2018, during the course of which hundreds of volunteer and Amazon Mechnial Turk forecasters, as well as automated machine models, answered more than 150 questions covering a broad range of issues.
+The Hybrid Forecasting Competition (HFC) is an IARPA program that seeks to develop methods for hybrid geopolitical forecasting system that combine human and machine forecasts to answer a broad range of questions about economic, political, health, and other events and trends.
+The first trial period, or RCT, took place in 2018, during the course of which hundreds of volunteer and Amazon Mechancial Turk forecasters, as well as automated machine models, answered more than 150 questions (IFPs) covering a broad range of issues.
 
 We worked on one of the competition teams, and specifically by contributing a time series forecasting module.
-Out of the large set of interesting questions one could examine with the results so far, given our specific focus on this project, we will try to examine in this paper two questions: (1) how does the accuracy of turker forecasters compare to volunteer forecasters, and (2) what was the impact of the machine forecasts on human forecaster accuracy?
+Among the questions we investigate in this paper are: (1) how does the accuracy of turker forecasters compare to volunteer forecasters, (2) what was the impact of time series charts and machine forecasts on human accuracy, (3) are there areas in which machine forecasts performed better than human forecasts, and vice versa, and (4) what explains the good performance of one particular group of volunteer forecasters who saw time series charts but were not exposed to model forecasts.
 
@@
 For example, machine forecasts can be reliable and scalable, but are constrained by available data, idiosyncratic questions, and cold start problems when a corpus of historical data is not available.
-Human-generated forecasts on the other hand are more flexible, but also more costly to scale and subject to various cognitive biases.[e.g.
+Human-generated forecasts on the other hand are more flexible, but also more costly to scale and subject to various cognitive biases [e.g.
 @kahneman:2011].
@@
 
-Each question includes 2 to 5 answer options to which forecasters must assign weights summing to 1. Performance for a single forecast is then based on an ordered multinomial Brier score.
+Each question includes 2 to 5 answer options to which forecasters must assign weights summing to 1.
+Once a question has closed and a resolution is available, the forecasts are scored.
+The primary performance measures were various aggregations derived from multinomial Brier scores.
 
-In addition to this basic "human forecaster" tournament, competitors were expected to implement features that would in some fashion or agument these human forecasts with machine-generated tools and forecasts.
-An explicity requirement for the latter is that they are automated systems, e.g.
+In addition to this basic "human forecaster" tournament, competitors were expected to implement features that would in some fashion or argument these human forecasts with machine-generated tools and forecasts.
+An explicitly requirement for the latter is that they are automated systems, e.g.
 an ad-hoc hand-tuned and expert-implemented machine model to forecast on a question would not be allowed, rather it would have to be a system that can generate such a model.
@@
 
-## SAGE approach
+## Automatic systems for acquiring data and producing time series charts and forecasts
 
-One of the distinguishing features of our team's approach was a system that could automatically associate some of the RCT-A questions with a clearly corresponding time series and display these to a forcaster.
+One features of our team's approach was a system that could automatically associate some of the RCT-A questions with a clearly corresponding time series and display these to a forecaster.
 This is based on a data platform which automatically collects and updates data from a variety of sources, and which can associate questions, based on their title, to an appropriate transformation of a data set if it is in the platform and matches a known, pre-specified pattern or template.
@@
 
-Additionally, we could use the time series to generate a machine forecast based on a univariate ARIMA model, and which then would either be shown to a user, and/or submitted seperately as a standalone forecast.
-The module generating these forecasts, "basil-ts", was based on the Auto ARIMA model in @hyndman:khandakar:2008, which consists of a ARIMA-family model with an automated algorithm for determining a reasonble specific model structure, e.g.
+Additionally, we could use the time series to generate a machine forecast based on a univariate ARIMA model, and which then would either be shown to a user, and/or submitted separately as a standalone forecast.
+The module generating these forecasts, "basil-ts", was based on the Auto ARIMA model in @hyndman:khandakar:2008, which consists of a ARIMA-family model with an automated algorithm for determining a reasonable specific model structure, e.g.
 whether and how many differencing orders to apply, AR and MA orders, and some other parameters.
@@
 
-Since the scientific goal of HFC is to evaluate various hybrid forecasting techniques, the pripmary aspect of our research design for RCT-A was designed to assess what impact exposure to the time series charts and model forecasts would have on human forecaster accuracy.
+Since the scientific goal of HFC is to evaluate various hybrid forecasting techniques, the primary aspect of the research design for RCT-A was to assess what impact exposure to the time series charts and model forecasts would have on human forecaster accuracy.
 Incoming forecasters were assigned to one of three experimental conditions.
-The first group, A, served as control group and only had access to a basic version of the online platform showin the question information and tools to enter weights for the answer options.
+The first group, A, served as control group and only had access to a basic version of the online platform showing the question information and tools to enter weights for the answer options.
 The second group, B, could also see the time series charts, and the third group, C, could see the chart and machine forecast.
@@
 - **Amazon Mechanical Turk forecasters**.
-Due to lower than expected activity levels, turker forecasters started to be provided several weeks intot he first trial period.
+Due to lower than expected activity levels, turker forecasters started to be provided several weeks into the first trial period.
 Given activity levels at that time, a decision was made to assign turkers to either condition A or C, but not B.
-This mean that the group whose treatment was to only see charts (B) consisted only of volunteers, while the other two groups contained mixed populations of volunteers and turkers, thus adding a confounding factor for assessing the impact of charts and machine models.
+This meant that the group whose treatment was to only see charts (B) consisted only of volunteers, while the other two groups contained mixed populations of volunteers and turkers, thus adding a confounding factor for assessing the impact of charts and machine models.
 - **Gaps in data and machine coverage**.
@@
 - **Changes in the data and machine forecasting platforms over time**.
-Both the data platform and the machine forecasting system suffered from various bugs and related issues, especially during the earlier portions of RCT-A.
-These problems in some cases resulted in incorrectly aggregated or otherwise inaccurate data, insufficient updating which led to data in the platform falling behind source availability and thus requiring forecasts over longer time periods than neccessary, and bugs in the machine forecaster that led to no or bad forecasts.
-The quality of the data and machine forecasts displayed to some users thus varied over time and IFPs, in ways that are difficult to reconstruct in retrospective.
-
-In respect to the first two problems, which we can quantify, the effective treatment groups were more complicated and are showin in Table \ref{tab:all-conditions}.
+Both the data platform and the machine forecasting system experienced various bugs and related issues, especially during the earlier portions of RCT-A.
+These problems in some cases resulted in incorrectly aggregated or otherwise inaccurate data, insufficient updating which led to data in the platform falling behind source availability and thus requiring forecasts over longer time periods than necessary, and bugs in the machine forecaster that led to no or bad forecasts.
+As these were addressed over the course of time, the quality of the data and machine forecasts displayed to some users varied at any given point in time and IFP, in ways that are difficult to reconstruct in retrospective.
 
@@
 
-# Design, data, and method
+In respect to the first two problems, which we can quantify, the effective treatment groups were more complicated and are shown in Table \ref{tab:all-conditions}.
+Within each of the original experimental conditions (A, B, C), there were potentially both volunteer and turker forecasters.
+They made forecasts on group of IFPs for which data was not available at all, on which data and a chart and a machine forecast were available, and also a small number of questions for which data and a chart, but not a machine forecast, were available.
+Since it is not possible to show a chart for an IFP for which data were not available, even in conditions B and C actual exposure to charts and model forecasts was also conditional on the IFP group in respect to data availability.
+The last two columns thus mark the groups of forecasters who actually would have seen a chart or model forecast on a question.
+Finally, we also mark the subset of questions for which there was a machine forecast.
 
-## DGP
+# Data
 
-Question comes from T&E to our platform.
-
-Now on two tracks:
-
-1. Human
-  - sees question, and depending on condition the chart and machine forecast
-  - forecasts, possibly multiple times per day
-  - Mean daily Brier score
-
-2. Platform
-  - parses question, finds associated data
-  - sends question and data to forecaster
-  - forecaster parses question and data, TS forecast, convert to MN forecast
-  - mean daily Brier score and/or shown to user
-
-## Data
+The data that we will use for the empirical exploration consist of single forecasts that a user (volunteer, turker, machine) made at a particular time (date) for a question (IFP), i.e.
+a set of probabilities for the number of options that that question had.
 
 RCT-A lasted from March 7th to September 7th 2018.
-However, Turkers did not enter until May 2nd, and for an unrelated reason we also discard data after August 2nd.
-This leaves a total of 48,931 forecasts---7,816 from volunteer forecasters, 39,140 from Turkers, and 1,975 from machine---for 156 IFPs, of which 101 did not have TS data, 6 had TS data but not a machine forecast, and 49 had both TS data and a machine forecast.
+However, turkers did not enter until May 2nd, and for an unrelated reason we also discard data after August 2nd.
+This leaves a total of 48,931 forecasts---7,816 from volunteer forecasters, 39,140 from turkers, and 1,975 from machine---for 156 IFPs, of which 101 did not have TS data, 6 had TS data but not a machine forecast, and 49 had both TS data and a machine forecast.
 
-# Results
+The basic measure of forecast quality are multinomial Brier scores that range from 0 to 2, with lower values indicating better performance.
+This is used for questions with only two or several unordered responses, and calculated as:
+
+$$
+mBS = \sum_{i=1}^R(f_i - o_i)^2 
+$$
+
+In which $R$ is the number of answer options, $f$ is a vector of probabilities for each option, and $o$ is a vector indicating the correct answer option.
+The score ranges from 0 to 2, lower values indicating better accuracy.
+
+For questions with ordered answers, an ordinal variant of the multinomial Brier score is used so that "near-misses" are correctly penalized less than far misses, unlike in the multinomial Brier score.
+For a question with 4 answer options, it would be calculated with the following steps:
+
+1. Split the ordinal categories (A-B-C-D) into cumulative binary pairs, aggregating the forecast probabilities for each grouping of categories (A-BCD; AB-CD; ABC-D). 
+2. Calculate the multinomial Brier score for each of the binary categories.
+3. Average across the binary category scores to obtain the final Brier score.
+
+The values of this score also range from 0 to 2.
+
+The primary covariates or independent variables of interest correspond to the various factors defining the groups of forecasters in Table \ref{tab:all-conditions}.
+We did not aim to incorporate other external factors, e.g.
+characteristics of the users beyond whether they were volunteers or turkers, for example.
+
+# Empirical analysis
 
@@
 
-Volunteers who had only charts did the best
-
-```{r}
-options(knitr.kable.NA = '')
-knitr::opts_current$set(label = "brier-by-forecaster-condition")
-tbl <- read_csv("../output/tables/brier-by-forecaster-condition.csv")
-tbl %>%
-  spread(Condition, avg_Brier) %>%
-  knitr::kable(., format = "latex", digits = 2, booktabs = TRUE,
-             caption = "Average Brier score by forecaster group")
-```
-
-Table \ref{tab:brier-by-forecaster-ifp-group-condition} summarizes the average Brier scores for all 16 treatment/forecaster/IFP groups in Table XX.
+Table \ref{tab:brier-by-forecaster-ifp-group-condition} summarizes the average Brier scores for all 16 treatment/forecaster/IFP groups in Table \ref{tab:all-conditions}.
 As it is quite unwieldy, we include it for reference and will discuss specific insights in more details below.
@@
 
+## Aggregate results
+
+Table \ref{tab:brier-by-forecaster-condition} shows a high level summary of average Brier scores by forecaster type and condition.
+Overall, volunteer forecasters in condition B had the best accuracy, turkers overall appear to have had poorer accuracy, and the machine forecasts were somewhat middling and generally lower than expected, but without a doubt not competitive with volunteer forecasts, even if they were forecasting without the benefit of a time series chart or model forecast.
+
+```{r}
+knitr::opts_current$set(label = "brier-by-forecaster-condition")
+tbl <- read_csv("../output/tables/brier-by-forecaster-condition.csv")
+tbl %>%
+  spread(Condition, avg_Brier) %>%
+  knitr::kable(., format = "latex", digits = 2, booktabs = TRUE,
+             caption = "Average Brier score by forecaster group")
+```
+
+These aggregate patterns don't reflect confounding factors, e.g.
+the possibility that turker accuracy is somehow influenced by the fact that they were only present in conditions A and C, etc.
+The next sections will examine in more depths these areas:
+
+- Turker performance relative to volunteer forecasters
+- What were the direct chart and model effects on the quality of a human forecasts?
+- Are there specific areas in which machine forecasts seemed to outperform human forecasters?
+- Why did condition B volunteers do so well compared to other forecaster-condition groups?
+
 ## Turkers generally did worse
@@
 I.e.
-for them, values below zero indicate an improvment.
+for them, values below zero indicate an improvement.
 Note that the x-scale is reversed, so that better is further to the right.
@@
 Points further to the right correspond to lower Brier scores, i.e.
-better peformance.
+better performance.
 Volunteers on average performed better than turker forecasters in all conditions and question groups, except on the small number of charts-only IFPs, where estimates are uncertain.}
@@
 In most groups volunteers have an advantage sufficient to generate p-values below 0.05, although the effects are minute in comparison to general between-forecaster differences: all models have very small adjusted R^2 values, below 0.01.
+Thus, while turker forecasters on average did not perform as well as volunteer forecasters, the difference in forecast quality for specific users and questions differs far more than the means between the two groups do.
 
@@
 
-Teasing out the chart and model effects is a bit harder through pairwise comparison.
+Teasing out the chart and model effects is a bit harder through pairwise comparison since a forecaster seeing a machine forecast would have also always seen a chart.
 Instead, we directly coded whether a forecaster saw a chart or model forecast.
-Forcasters in conditions B and C and for questions that had a chart saw a chart, and forecasters in condition C looking at a question which had a machine forecast will have seen a model forecast.
-We also leave out the machine forecasts from the data, leaving almost 47,000 forecasts, of which about 25% were made with a chart available, and 21% also had a model forecast available.
+Forecasters in conditions B and C and for questions that had a chart saw a chart, and forecasters in condition C looking at a question which had a machine forecast will have seen a model forecast.
+We also leave out the machine forecasts from the data for this portion, leaving almost 47,000 forecasts, of which about 25% were made with a chart available, and 21% also had a model forecast available.
 
-We estimated two linear models to predict Brier scores for a forecast. The first has the specifiction:
+We estimated two linear models to predict Brier scores for a forecast. The first has the specification:
 
@@
 
-The second model includes random intercepts for IFP to account for the possibility that seeings charts or models may make users differentially willing to forecast on harder IFPs, i.e.
+The second model includes random intercepts for IFP to account for the possibility that seeing charts or models may make users differentially willing to forecast on harder IFPs, i.e.
 those with higher average Brier scores.
@@
 Seeing a chart slightly helps, seeing a model slightly hurts forecast performance.
-However, the effects are small and not as important as whether a forecasters was a volunteer or turker.
+However, the effects are small and not as important as whether a forecaster was a volunteer or turker.
 In addition, the models themselves capture only a small portion of the variation in Brier scores overall.
@@
 Thus, while we find that exposure to charts and model forecasts has a small impact on forecast quality, the effect sizes are small in comparison to other systematic factors in the model, and small relative to unsystematic and random factors not in the model.
+There is far more variation between individual forecasts (users, questions) than between the forecasts of users who saw a chart and/or model.
 
 \begin{figure}
-\caption{\label{fig:model-chart-and-model-effects} Model estimates for the direct effects of seeing a chart and machine forecast.}
+\caption{\label{fig:model-chart-and-model-effects} Model estimates for the direct effects of seeing a chart and machine forecast.
+Model 1 is a linear model with the charted terms, model 2 in addition has random intercepts for each question (IFP) to account for question difficulty.}
 \includegraphics{../output/figures/model-chart-and-model-effects.png}
@@
 
-To estimate the relative performance of different forecasters, we again estimated a simple linear model, similar to those used for the turker pariwise comparisons. The model has the general form:
+To estimate the relative performance of different forecasters, we again estimated a simple linear model, similar to those used for the turker pairwise comparisons. The model has the general form:
 
@@
 Condition B volunteers had the best overall performance out of the 16 groups shown in Table \ref{tab:brier-by-forecaster-ifp-group-condition}.
-Considering that the machine forecasts were relatively bad, one possible explanation is that the charts overall had a positive impact on human forecast accuracy, while seeing machine forecasts had a negative impact.
+Considering that the machine forecasts were of middling quality, one possible explanation is that the charts overall had a positive impact on human forecast accuracy, while seeing machine forecasts had a negative impact.
 
-However, there are several additional patterns that are either inconsistent with this simple explanation or at least suggest the need for a more complex alternative.
+However, there are several additional patterns in the forecasts that are either inconsistent with this simple explanation or at least suggest the need for a more complex alternative.
 The first is that condition B forecasters did not only do better than other groups on questions that had a chart, but also on questions that did not have a chart.
@@
 
-We could exlain this if there is, in addition to the direct beneficial effect of seeing a chart, but not a potentially misleading machine forecast, a spillover effect whereby exposure to charts also improves a volunteer forecaster's ability to forecast on other questions without charts.
-
-Curious inconsistencies remain:
+Maybe there is, in addition to the direct beneficial effect of seeing a chart for a question, an indirect spillover effect whereby exposure to charts also improves a volunteer forecaster's ability to forecast on other questions without charts.
+For example, it could be that seeing charts is an effective reminder of taking relevant baselines into account when forecasting.
+There are still some curiosities:
 
@@
 
-**Condition C volunteers do well on questions where we know that data displayed was wrong.** There were several instances where we know that the data displayed in a chart were inaccurate.
+**Condition C volunteers do well on questions where we know that the data displayed was wrong.** There were several instances where we know that the data displayed in a chart were inaccurate.
 Figure \ref{fig:iraq} shows the first example, oil production figures for Iraq.
-The main chart shows several teim series depicting oil production in Iraq, according to different sources.
+The main chart shows several time series depicting oil production in Iraq, according to different sources.
 The canonical data are contained in monthly OPEC oil market reports, each of which contains production figures for the past 3 months.
 Those are shown in the sequence of short time series on the mid-right part of the main chart--the figures are somewhat inconsistent between reports, but this is another peculiarity.
-Because this data is difficult to obtain programatically, an alternative data source is used instead.
+Because this data is difficult to obtain programmatically, an alternative data source is used instead.
 This is what would have been shown in the charts, and what informed the machine forecast.
-Obviously it is incorrect data.
+It is incorrect data.
 The dotted lines show the values separating the answer options for this question.
@@
 \begin{figure}
-\caption{\label{fig:iraq} Iraq oil production}
+\caption{\label{fig:iraq} Iraq oil production by data source.
+Official figures are to be drawn from monthly OPEC reports, each of which includes values for the past 3 months.
+Due to difficulty in automatically processing the PDF reports, an alternative data source was used in the chart and to create the machine forecast.}
 \includegraphics{../output/figures/opec-iraq-divergence.png}
@@
 
-Curiously, the condition B volunteers performed better on this question than any other forecaster.
-The number of forecasts here are low, and only the last three groups, starting from the machine forecasts, have pairwise statistically significant differences in average Brier score from the condition B volunteers.
+Curiously, the condition B volunteers performed better on this question than any other forecaster, with an average Brier score of 0.07, compared to 0.11 and 0.45 for volunteer forecasts in conditions A and C (Table \ref{tab:perf-1892}).
+The number of forecasts here are low however, and only the last three groups (machine, turkers in condition A, volunteers in condition C) have pairwise statistically significant differences in average Brier score from the condition B volunteers.
 
-|Forecaster |condition | avg_Brier|  n|
+|Forecaster |Condition | avg_Brier|  N|
 |:----------|:---------|---------:|--:|
@@
 
+Table: Forecaster performance for "How much crude oil will Iraq produce in May 2018?" \label{tab:perf-1892}
+
 Another example with incorrect data is an early ICEWS question that fell outside the common period and thus did not have forecasts from turkers.
+Performance values for this question are shown in Table \ref{tab:perf-911}.
 Condition B volunteers also outperformed all other groups on this question, although the pairwise differences with the other groups are not statistically significant.
@@
 
-\clearpage
+Table: Forecaster performance for "How many material conflict events involving Occupied Palestinian Territory will ICEWS record in March 2018?" \label{tab:perf-911}
 
-**Even when the machine predictions were good, condition B volunteers did better than their C counterparts.** If bad machine predictions influence human forecasts and are the reason why the beneficial chart effecgt is only apparent in condition B forecasters, and not condition C forecasters, then it should also be the case that condition C forecasters do well on those questions where the machine forecasts were good.
+**Even when the machine predictions were good, condition B volunteers did better than their C counterparts.** If bad machine predictions influence human forecasts and are the reason why the beneficial chart effect is only apparent in condition B forecasters, and not condition C forecasters, then it should also be the case that condition C forecasters do well on those questions where the machine forecasts were good.
 This is no the case.
-The table below shows the subset of forecasts on questions that had a machine forecast (and thus chart) and where the machine forecast average Brier was lower than the average Brier score for the forecasts from all other forecaster $\times$ condition groups.
-Condition B volunteers beat their condition C counterparts even when the machine model should have improved their forecasts because they happened to be accurate.
+Table \ref{tab:perf-good-machine} summarizes performance for forecasts on the 10 questions where the average machine forecast performance was better than the average performance for all other groups of forecasts, and where all groups of forecasters were represented, i.e.
+had forecasted.
+Since forecasters forecasted a differential amount of times in a way that is related to question difficulty, we cannot summarize the performance through a simple average.
+The relative performance measure shown in the table is instead calculated by first averaging Brier scores for each group on each question, calculating each group's relative performance as percentage deviation from the overall average of group averages for that question, and then finally averaging these per-question relative performances over questions for a final measure for each forecaster group.
 
-|Forecaster |Condition          | avg_Brier| n_fcasts|
-|:----------|:------------------|---------:|--------:|
-|Machine    |Machine            |      0.39|     1975|
-|Turker     |A: no chart        |      0.39|     2810|
-|Turker     |C: chart and model |      0.38|     8482|
-|Volunteer  |A: no chart        |      0.28|      267|
-|Volunteer  |B: chart only      |      0.23|      958|
-|Volunteer  |C: chart and model |      0.32|     1094|
+\newpage
 
-One possibility that we should not discard with these apparent inconsistencies is that we are digging in noise.
+|Forecaster |Condition          | avg_rel_Brier (%) | N_users| N_fcasts|
+|:----------|:------------------|-------------:|-------:|--------:|
+|Machine    |Machine            |         -0.55|       1|      423|
+|Turker     |A: no chart        |          0.23|     152|      610|
+|Turker     |C: chart and model |          0.25|     469|     1890|
+|Volunteer  |A: no chart        |         -0.13|       8|       50|
+|Volunteer  |B: chart only      |         -0.04|      20|      216|
+|Volunteer  |C: chart and model |          0.24|      40|      224|
+
+Table: Performance for questions where the machine forecasts had the lowest average Brier score accross groups.
+The 'avg_rel_Brier' column shows the expected deviation of a group's forecasts from the overall question average between groups, in %.
+Negative values indicate that a group's forecasts tended to do better than the overall average performance on a question.
+Had condition C forecasters sufficiently adjusted their forecasts in the direction of the machine forecasts, they should have outperformed other comparable forecaster groups.
+\label{tab:perf-good-machine}
+
+The forecasts summarized in the table are for the small set of questions where the machine forecasts did very well, with Brier scores that were 55% less than the between group average for a question.
+If human forecasters who see machine forecasts align their forecasts towards the machine forecasts, they should have done really well here.
+But they don't.
+The condition B volunteers still outperform their "i saw a really good machine forecast" condition C counterparts, and to a lesser extend the same trend is apparent with turkers who saw neither chart nor model and those that did.
+
+Each single example of an apparent consistency itself is based on small findings in noisy data.
 All of the effects---for turkers versus other forecasts, for the disappointing performance of the machine forecasts, for conditions or different groupings of IFPs---pale in comparison to differences in accuracy between individual forecasters and between forecasts for different IFPs.
@@
 
-Are we digging in noise?
+We reported an exploratory analysis that examined several aspects of the performance of volunteer, Amazon Mechanical Turker, and machine-generated forecasts for a portion of the first trial period in the Hybrid Forecasting Competition (HFC) project.
+These initial findings suggest that:
 
-Many of the questions involve monthly time series but are only open for periods spanning a single or few months.
-Some of the advantage of the human forecasters may thus have been due to operating on a sub-monthly time scale.
-
-We also know that model performance suffered from issues related to automation.
-Were we able to run the RCT-A questions with the current system and with a data platform that optimally ingests data sources on the bleeding edge of their own updates, our forecasts would have been ...
-
-**[check with updated results from test bed what the performance gain would have been, i think the site visit stats were not accurate anymore]**
+- Amazon Mechanical Turker forecasters underperform compared to volunteer forecasters.
+It may still be that there is a subset of turker forecasters that manage to consistently outperform the average volunteer forecaster, a question we did not examine but which is plausible due to the overall small difference in volunteer and turker average performance.
+- Volunteer forecasters who were able to see charts for the subset of questions with clearly associated time series overall had the best accuracy, even on questions that did not have chartable data.
+However, there are several potential empirical inconsistencies with the implied causal mechanism (charts have a direct beneficial effect on the question at hand, and some form of indirect beneficial effect on other, non-chart, question as well).
+- Machine forecasts on average did not perform as well as volunteer forecasts, but slightly better than turker forecasts.
+They appear to have had an edge on questions with count series derived by aggregating event data on protests or violence (ACLED; Boko Haram violence in Nigeria) as well as hacking (Privacyrights hacking incidents).
+One commonality in these sources is that they require data aggregation in a way that is not easily replicable by a user without some moderate level of technical skill.
 

### 2018-12-07 R/analysis.R

@@ -205,7 +205,7 @@ sink()
 
 bind_rows(
   tidy(mdl1) %>% mutate(Model = "Model 1: linear model"),
-  tidy(mdl2) %>% mutate(Model = "Model 2: linear model with IFP random intercepts") %>%
+  tidy(mdl2) %>% mutate(Model = "Model 2: linear model with IFP random\nintercepts") %>%
     filter(group=="fixed")
 ) %>%
   mutate(term = factor(term) %>% fct_recode(`Baseline\nForecasterTurker, IFP_GroupNo_TS_data` = "(Intercept)")) %>%
@@ -346,7 +346,7 @@ ggplot(cond_b_volunteers_n_vs_brier, aes(x = n_fcasts, y = avg_Brier, color = IF
   geom_smooth(method = "lm", se = FALSE) +
   theme_ipsum_rc() +
   labs(x = "Total number of forecasts by user")
-ggsave("output/figures/cond-b-volunteers-n-vs-brier.png", height = 5, width = 8)
+ggsave("output/figures/cond-b-volunteers-n-vs-brier.png", height = 4, width = 8)
 
 # neither slope seems to be significant
 summary(lm(avg_Brier ~ n_fcasts*IFP_had_chart, data = cond_b_volunteers_n_vs_brier))
@@ -427,23 +427,44 @@ tbl911
 # When the machine models did well, did condition C also do better, relative 
 # to B?
 good_machine <- common_period %>%
-  filter(Forecaster %in% c("Machine", "Volunteer")) %>%
+  # Get average for each group on each IFP
   group_by(ifp_id, Forecaster, Condition) %>%
   summarize(avg_Brier = mean(brier), n_fcasts = n()) %>%
   group_by(ifp_id) %>%
-  mutate(avg_Brier_for_ifp = weighted.mean(avg_Brier, n_fcasts),
-         machine_improvement = (avg_Brier - avg_Brier_for_ifp)/avg_Brier_for_ifp,
-         min_avg_Brier = min(avg_Brier)) %>%
+  # Don't do a weighted avg for Brier. We want an estimate of the performance
+  # of a group on an IFP, i.e. group average and the average of that
+  mutate(avg_Brier_for_ifp = mean(avg_Brier),
+         rel_performance = (avg_Brier - avg_Brier_for_ifp)/avg_Brier_for_ifp,
+         min_avg_Brier = min(avg_Brier),
+         n_groups = n()) %>%
+  # Make sure we had forecasts from all groups
[…]

### 2018-12-13 slides/slides.Rmd
clean up figure and start slides

@@
 ---
-title: "draft slides"
+title: Assessing Amazon Turker and automated machine forecasts in the Hybrid Forecasting Competition
 author: "Andreas Beger, Predictive Heuristics"
 date: "`r format(Sys.Date(), '%e %b %Y')`"
-output: xaringan::moon_reader
+output: 
+  xaringan::moon_reader:
+    chakra: libs/remark-latest.min.js
+    nature:
+      beforeInit: "macros.js"
 ---
@@
 
-# First slide
+class: middle
+
+# Disclaimer
+
+This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
+The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S.
+Government.
+The U.S.
+Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.
 
 ---
+class: center, middle, inverse
 
-# Second slides
+# Hybrid Forecasting Competition (HFC)
+
+---
+class: center, middle
+
+![:scale 70%](img/model-machine-by-data-source.png)

### 2018-12-13 R/analysis.R

@@ -292,6 +292,8 @@ mdl3_coefs <- tidy(mdl3) %>%
   mutate(Forecaster = gsub("Forecaster", "", Forecaster),
          `Data source` = gsub("Data_source", "", `Data source`)) %>%
   mutate(`Data source` = factor(`Data source`))
+# Reorder by median estimate; makes plot easier to read
+mdl3_coefs$`Data source` = fct_reorder(mdl3_coefs$`Data source`, mdl3_coefs$estimate)
 
 mdl3_coefs %>%
   mutate(facet = ifelse(Forecaster=="Machine (reference)", "Baseline", "Relative difference")) %>%
@@ -476,56 +478,9 @@ ggplot(fcasts_good_machine, aes(x = interaction(Forecaster, Condition), y = brie
 
 
 
-## older stuff, organize
 
-# activity by question
-fcasts_by_ifp <- all_fcasts %>%
-  group_by(ifp_id) %>%
-  summarize(avg_Brier = mean(brier), n = n()) %>%
-  arrange(desc(n))
-fcasts_by_ifp
-tail(fcasts_by_ifp)
+# Selection effects -------------------------------------------------------
 
-# 2126 had a lot of forecasts
-f2126 <- filter(all_fcasts, ifp_id==2126)
-table(f2126$Forecaster, f2126$condition)
-f2126 %>% 
-  group_by(Forecaster, condition) %>%
-  summarize(avg_Brier = mean(brier), n = n())
-
-# 2297
-f2297 <- filter(all_fcasts, ifp_id==2297)
-table(f2297$Forecaster, f2297$condition)
-f2297 %>% 
-  group_by(Forecaster, condition) %>%
-  summarize(avg_Brier = mean(brier), n = n())
-
-# low activity 
-# 1208
-f1208 <- filter(all_fcasts, ifp_id==1208)
-table(f1208$Forecaster, f1208$condition)
-f1208 %>% 
-  group_by(Forecaster, condition) %>%
-  summarize(avg_Brier = mean(brier), n = n())
-
-# 1235
-f1235 <- filter(all_fcasts, ifp_id==1235)
-table(f1235$Forecaster, f1235$condition)
-f1235 %>% 
-  group_by(Forecaster, condition) %>%
-  summarize(avg_Brier = mean(brier), n = n())
-
-# 1172
-f1172 <- filter(all_fcasts, ifp_id==1172)
-f1172 %>% 
[…]

### 2018-12-13 R/make-machine-forecast-data.R
check fcast data discrepancy between dump files

@@ -115,6 +115,12 @@ machine_fcasts %>%
   summarize(avg_brier = mean(brier))
 
 
+# Discrepancy between source files ----------------------------------------
+#
+#   The aggregation dump and machine forecast dump files don't match,
+#   it seems that both are incomplete.
+#
+
 fcast_dump <- read_csv(
   file.path(in_dir, "dump_machine_forecasts_11_29_2018_20_17_42.csv")
 ) %>%
@@ -151,3 +157,63 @@ check2 <- check %>%
 filter(check2, !ifp_id %in% fcast_dump$ifp_id)
 # both agg and m dump sometimes seem to be missing stuff, weird
 
+
+md_by_day <- fcast_dump %>%
+  group_by(date) %>%
+  summarize(n_forecasts = n()) %>%
+  complete(date = seq(min(date), max(date), by = "day"), 
+           fill = list(n_forecasts = 0)) %>%
+  mutate(source = "machine dump")
+
+agg_by_day <- machine_fcasts %>%
+  group_by(date) %>%
+  summarize(n_forecasts = n()) %>%
+  complete(date = seq(min(date), max(date), by = "day"), 
+           fill = list(n_forecasts = 0)) %>%
+  mutate(source = "agg dump") 
+
+by_day <- bind_rows(md_by_day, agg_by_day) 
+
+ggplot(by_day, aes(x = date, y = n_forecasts, color = source)) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(c("2018-04-13", "2018-08-01")), linetype = 3) +
+  annotate("text", x = as.Date(c("2018-04-13", "2018-08-01")) + 2, y = 26,
+           label = c("2018-04-13\nFirst forecast in aggregation dump", 
+                     "2018-08-01\nLast forecast in aggregation dump"), hjust = 0) +
+  theme_ipsum_ps()
+
+# How many IFPs with ARIMA were there after August 1st?
+
+ifp_have <- fcast_dump %>%
+  filter(date > "2018-08-01") 
+question_dump <- "dump_questions_11_29_2018_20_17_01.csv"
+q_dump <- read_csv(
+  file.path(in_dir, question_dump),
+  col_types = cols(
+    .default = col_character(),
+    ifp_id = col_integer(),
+    discover_id = col_integer(),
+    start_date = col_datetime(format = ""),
+    end_date = col_datetime(format = ""),
+    resolved_date = col_datetime(format = ""),
[…]

### 2018-12-13 R/scratchpad.R

@@ -1,164 +1,0 @@
-library("tidyverse")
-
-mbrier <- function(f, o, ordered = TRUE) {
-  prediction = f
-  outcome = o
-  
-  stopifnot(length(prediction)==length(outcome))
-  stopifnot(all(outcome %in% c(0L, 1L, NA)))
-  stopifnot(all.equal(sum(prediction), sum(outcome), 1L) | any(is.na(prediction)))
-  
-  if (ordered) {
-    pairs <- NULL
-    for (i in 1:(length(prediction)-1)) {
-      p_i <- sapply(split(prediction, 1:length(prediction) > i), sum)
-      o_i <- sapply(split(outcome, 1:length(outcome) > i), max)
-      pairs <- c(pairs, sum((p_i - o_i)^2))
-    }
-    brier <- mean(pairs)
-  } else {
-    brier <- sum((prediction - outcome)^2)
-  }
-  
-  return(brier)
-}
-
-
-user_raw <- read_csv("input-data/dump_user_forecasts_09_07_2018_09_15_12.csv") %>%
-  mutate_at(vars(contains("date")), as.Date) %>%
-  arrange(ifp_id, date, user_id) %>%
-  group_by(ifp_id, date, user_id) %>%
-  mutate(fcast_seq = 1:n()) %>%
-  mutate(
-    num_options = case_when(
-      num_options==1 ~ 2L,
-      TRUE ~ num_options
-    ),
-    option_2 = case_when(
-      is.na(option_2) ~ 1 - option_1,
-      TRUE ~ option_2
-    )) %>%
-  tidyr::gather(option, prob, starts_with("option")) %>%
-  mutate(prob = prob / 100) %>%
-  mutate(option = str_replace(option, "option\\_", "")) %>%
-  filter(option <= num_options) 
-
-
-suppressWarnings(
-  q_dump <- read_csv("input-data/dump_questions_09_07_2018_09_30_55.csv",
-                     col_types = cols(
-                       .default = col_character(),
-                       ifp_id = col_integer(),
-                       discover_id = col_integer(),
-                       start_date = col_datetime(format = ""),
-                       end_date = col_datetime(format = ""),
-                       resolved_date = col_datetime(format = ""),
-                       voided_date = col_datetime(format = ""),
-                       num_options = col_integer(),
-                       is_ordinal = col_logical(),
[…]

### 2018-12-13 R/analysis.R
Move 1003 up in analyais

@@ -426,6 +426,20 @@ tbl911 <- f911 %>%
   arrange(avg_Brier)
 tbl911
 
+# 1003 Loya Jirga
+# System persistently kept showing wrong chart, with ACLED riots/protests
+# B volunteers have lower Brier on this, why?
+# C turkers as expected had worse score than A turkers
+f1003 <- filter(common_period, ifp_id==1003) %>%
+  mutate(Group = paste0(Forecaster, ", ", condition),
+         Group = factor(Group) %>% fct_relevel("Volunteer, b"))
+summary(lm(brier ~ Group, data = f1003))
+tbl1003 <- f1003 %>%
+  group_by(Forecaster, condition) %>%
+  summarize(avg_Brier = mean(brier), n = n()) %>%
+  arrange(avg_Brier)
+tbl1003
+
 # When the machine models did well, did condition C also do better, relative 
 # to B?
 good_machine <- common_period %>%
@@ -549,19 +563,7 @@ foo %>%
 
 # Specific questions ------------------------------------------------------
 
-# 1003 Loya Jirga
-# System persistently kept showing wrong chart, with ACLED riots/protests
-# B volunteers have lower Brier on this, why?
-# C turkers as expected had worse score than A turkers
-f1003 <- filter(common_period, ifp_id==1003) %>%
-  mutate(Group = paste0(Forecaster, ", ", condition),
-         Group = factor(Group) %>% fct_relevel("Volunteer, b"))
-summary(lm(brier ~ Group, data = f1003))
-tbl1003 <- f1003 %>%
-  group_by(Forecaster, condition) %>%
-  summarize(avg_Brier = mean(brier), n = n()) %>%
-  arrange(avg_Brier)
-tbl1003
+
 
 # activity by question
 fcasts_by_ifp <- all_fcasts %>%

### 2018-12-13 docs/index.Rmd
rename slides and copy to docs for GH pages

---
title: Assessing Amazon Turker and automated machine forecasts in the Hybrid Forecasting Competition
author: "Andreas Beger, Predictive Heuristics"
date: "`r format(Sys.Date(), '%e %b %Y')`"
output: 
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom_theme.css"]
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(icon)
```

<!-- https://apreshill.github.io/data-vis-labs-2018/slides/06-slides_xaringan.html#13 -->

class: middle

# Disclaimer

This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005. The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S. Government. The U.S. Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.

---
class: center, middle, inverse

# Hybrid Forecasting Competition (HFC)

---

# Project goal

>The HFC program is developing and testing hybrid geopolitical forecasting systems. These systems integrate human and machine forecasting components to create maximally accurate, flexible, and scalable forecasting capabilities.
>
>Human-generated forecasts may be subject to cognitive biases and/or scalability limits. Machine-generated (i.e., statistical, computational) forecasting approaches may be more scalable and data-driven, but are often ill-suited to render forecasts for idiosyncratic or newly emerging geopolitical issues. Hybrid approaches hold promise for combining the strengths of these two approaches while mitigating their individual weaknesses.
>
>Performers are developing systems that integrate human and machine forecasting contributions in novel ways. These systems are competing in a multi-year competition to identify approaches that may enable the Intelligence Community (IC) to radically improve the accuracy and timeliness of geopolitical forecasts.

.small[From https://www.hybridforecasting.com]

---
background-image: url('img/screenshot-splash.png')
background-size: contain

---
background-image: url('img/screenshot-question.png')
background-size: contain

---

# Scoring

Multinomial Brier score:

$$mBS = \sum_{i=1}^R(f_i - o_i)^2$$

- $R$ is the number of answer options, $f$ the vector of weights summing to 1, and $o$ a 0/1 vector marking the correct option
- ranges from 0 (good) to 2 (bad)

--

Ordered Brier score:

1. Split the ordinal categories (A-B-C-D) into cumulative binary pairs, aggregating the forecast probabilities for each grouping of categories (A-BCD; AB-CD; ABC-D). 
2. Calculate the multinomial Brier score for each of the binary categories.
3. Average across the binary category scores to obtain the final Brier score.

- also ranges from 0 to 2
- "near misses" are penalized less than far misses

---
class: center, middle, inverse

# The Hybrid part:

# Some users could see time series charts and machine forecasts for some IFPs

---

# How this works

- dig

---

# Data

- RCT-A, the first trial period, lasted from 7 March to 7 September 2018
- Use subset of forecasts from 2 May to 1 August 2018

N fcasts, n IFPs

---

# Summary of some findings in the paper

Volunteers are better forecasts than Amazon Turkers

Machines are somewhere in the middle

Human forecasters who saw the machine forecasts did poorly; volunteer forecasts who saw only charts had overall the best performance

It's not clear why though, e.g.:
- they outperformed on questions that did not have data (charts and model)
- they outperformed volunteer forecasters who saw the machine forecasts even when the machine forecasts had good accuracy

Relative performance: where did the machine forecasts do better than human forecasters?

---

Two models

---
class: center, middle

![:scale 80%](img/model-machine-by-data-source.png)

---

# ACLED 5-option

---

# OECD interest rates

---

# ACLED binary question

---

Machine forecasts did well on count questions that require data aggregation

Some of the overall hardest questions, and where to some extent human forecasters did better, are economic/financial monthly series

- OECD interest rates
- FAO food price indices

---
class: center, middle, inverse

# Thank you!

Register to forecast at: 
Internal platform:

.left[
`r icon::fa("envelope")`: [adbeger@gmail.com](mailto:adbeger@gmail.com)

`r icon::fa("github-square")`: https://github.com/andybega/asia-polmeth-2019

`r icon::fa("file")`: [link to paper (on github under docs/)](https://github.com/andybega/asia-polmeth-2019/blob/master/docs/BegerWard_HFC_AsiaPolmeth2019.pdf)

`r icon::fa("slideshare")`: [slide source](https://github.com/andybega/asia-polmeth-2019/tree/master/slides)
]

### 2018-12-13 slides/index.Rmd
try to fix pdf link

--- 2018-12-13 docs/index.Rmd
@@
 
-`r icon::fa("file")`: [link to paper (on github under docs/)](https://github.com/andybega/asia-polmeth-2019/blob/master/docs/BegerWard_HFC_AsiaPolmeth2019.pdf)
-
-`r icon::fa("slideshare")`: [slide source](https://github.com/andybega/asia-polmeth-2019/tree/master/slides)
+`r icon::fa("file")`: [link to paper (on github under docs/pdf/)](pdf/BegerWard_HFC_AsiaPolmeth2019.pdf)
 ]

### 2018-12-13 slides/index.Rmd
enough slide work for the day

@@
 title: Assessing Amazon Turker and automated machine forecasts in the Hybrid Forecasting Competition
-author: "Andreas Beger, Predictive Heuristics"
+author: "Predictive Heuristics </br>Andreas Beger and Michael D. Ward"
 date: "`r format(Sys.Date(), '%e %b %Y')`"
@@
 
-class: middle
+class: center
+
+# Preview
+
+.left.large[I'm going to show results of an exploratory analysis of the relative accuracy of volunteer, Amazon Mechanical Turker, and automated machine forecasts for a broad set of questions (IFPs)<sup>1</sup> in the HFC<sup>2</sup> during the first trial period (RCT-A<sup>3</sup>) that took place this year.]
+
+.left.footnote[<sup>1</sup> Individual Forecasting Problem...I may not be able to keep myself from using HFC jargon. </br>
+<sup>2 </sup> Hybrid Forecasting Competition </br>
+<sup>3 </sup> Randomized controlled trial A]
+
+---
+class: middle, center
 
@@
 
-This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
+.left[This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005.
 The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S.
@@
 The U.S.
-Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.
+Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.]
 
@@
 
-# Hybrid Forecasting Competition (HFC)
+# Hybrid Forecasting Competition
+
+.pull-left[![](img/hfc-logo.png)]
+
+.pull-right[![](img/sage-logo.png)]
 
@@
 >The HFC program is developing and testing hybrid geopolitical forecasting systems.
-These systems integrate human and machine forecasting components to create maximally accurate, flexible, and scalable forecasting capabilities.
->
->Human-generated forecasts may be subject to cognitive biases and/or scalability limits.
-Machine-generated (i.e., statistical, computational) forecasting approaches may be more scalable and data-driven, but are often ill-suited to render forecasts for idiosyncratic or newly emerging geopolitical issues.
-Hybrid approaches hold promise for combining the strengths of these two approaches while mitigating their individual weaknesses.
->
->Performers are developing systems that integrate human and machine forecasting contributions in novel ways.
-These systems are competing in a multi-year competition to identify approaches that may enable the Intelligence Community (IC) to radically improve the accuracy and timeliness of geopolitical forecasts.
+These systems **integrate human and machine forecasting components** to create maximally accurate, flexible, and scalable forecasting capabilities.
 
-.small[From https://www.hybridforecasting.com]
+.right.small[From https://www.hybridforecasting.com]
+
+The idea is to overcome some of the respective weaknesses of human and machine-generated forecasts by combining them in some to be determined fashion.
+
+Human-generated forecasts are flexible, but subject to cognitive biases and scalability issues.
+
+Machine-generated forecasts are reliable and scalable, but ill-suited for idiosyncratic or emerging issues.
 
@@
 
-- dig
+- An automated system is collection and updating data from several data sources
+- An automated system identifies if data for a particular question is available
+
+If so:
+
+- Chart the data
+- An automated system generates a machine forecast
+
+---
+
+# basil-ts
+
+Automated time series forecaster microservice
+
+Implemented in R + Python Flask + RESTful API
+
+Sketch of the internals:
+
+1. Parse incoming question and data
+2. Produce a time series forecast using an automated ARIMA fitter<sup>1</sup>
+3. Convert the time series forecast to answer option probabilities
+
+Most of the complexity is related to automating the question/task parsing and handling edge cases.
+
+.footnote[<sup>1</sup> [Hyndman & Khandakar 2008](http://www.jstatsoft.org/article/view/v027i03); R [forecast](https://cran.r-project.org/package=forecast) package]
+
+---
+class: middle, center, inverse
+
+# Let's start looking at data and results
 
@@
 
-Two models
+# Model overview
+
+Group questions by data source, N=14 groups
+
+Compare average Brier scores of volunteer, turker, and machine forecasts (reference category)
+
+$$\textrm{Brier}_{ij} = \beta_{i}\textrm{DataSource}_i + \beta'_{ij}(\textrm{DataSource}_i\times\textrm{Forecaster}_j)$$
+
+$\beta_i$ = average Brier score for machine forecasts for question group $i$
+
+$\beta'_{ij}$ = average Brier score *relative to machine forecasts* for forecaster group $j$ in question group $i$
+
+- Negative values $\rightarrow$ human forecasters did better
+- Positive values $\rightarrow$ machine forecasts did better
 
@@
 
+# Conclusions
+
 Machine forecasts did well on count questions that require data aggregation
@@
 Register to forecast at:
+
 Internal platform:

### 2019-01-04 slides/index.Rmd
update slides

@@
 author: "Predictive Heuristics </br>Andreas Beger and Michael D. Ward"
-date: "`r format(Sys.Date(), '%e %b %Y')`"
+date: "`r format(Sys.Date(), '%e %B %Y')` </br>Asian POLMETH 2019, Kyoto, Japan"
 output: 
@@
 
-class: center
-
-# Preview
+class: center, middle
 
@@
 
-.left.footnote[<sup>1</sup> Individual Forecasting Problem...I may not be able to keep myself from using HFC jargon. </br>
+.left.footnote[<sup>1</sup> Individual Forecasting Problem</br>
 <sup>2 </sup> Hybrid Forecasting Competition </br>
@@
 ---
+class: middle
 
@@
 
+.middle[
 >The HFC program is developing and testing hybrid geopolitical forecasting systems.
 These systems **integrate human and machine forecasting components** to create maximally accurate, flexible, and scalable forecasting capabilities.
+]
 
@@
 The idea is to overcome some of the respective weaknesses of human and machine-generated forecasts by combining them in some to be determined fashion.
+
+???
 
@@
 
-# The Hybrid part:
+# The hybrid part:
 
-# Some users could see time series charts and machine forecasts for some IFPs
+# Some users could see time series charts and/or machine forecasts for some IFPs
 
@@
 
-- An automated system is collection and updating data from several data sources
-- An automated system identifies if data for a particular question is available
+- An automated system is collecting and updating data from several data sources, and matching them up, if possible, to questions
 
-If so:
+--
+
+If data for a particular question is available:
 
@@
 - An automated system generates a machine forecast
+
+---
+class: middle
+
+### How much crude oil will Iraq produce in May 2018?
+
+.center[
+![:scale 80%](img/opec-iraq-with-chart.png)
+]
+
+- Less than 4,280
+- Between 4,280 and 4,384, inclusive
+- More than 4,384 but less than 4,473
+- Between 4,473 and 4,576, inclusive
+- More than 4,576
+
+---
+class: middle
+
+### How much crude oil will Iraq produce in May 2018?
+
+.center[
+![:scale 80%](img/opec-iraq-with-forecast.png)
+]
+
+- Less than 4,280 *(38%)*
+- Between 4,280 and 4,384, inclusive *(15%)*
+- More than 4,384 but less than 4,473 *(13%)*
+- Between 4,473 and 4,576, inclusive *(13%)*
+- More than 4,576 *(21%)*
+
+???
+
+The model is ARIMA(0, 1, 1)
 
@@
 
+???
+
+Examples of parsing that needs to be done:
+
+- how far ahead to forecast, which requires parsing the data and question time periods
+- whether the data are count and should be constrained
+- partial data handling
+- data aggregation for fixed period questions
+
+---
+
+# Questions that were of primary concern to us
+
+--
+
+How well are the machine forecasts doing?
+
+--
+
+Are there particular question groups where performance is good or lacking? Basically, where do we need to focus improvements?
+
 ---
@@
 
-- RCT-A, the first trial period, lasted from 7 March to 7 September 2018
-- Use subset of forecasts from 2 May to 1 August 2018
+RCT-A, the first trial period, lasted from 7 March to 7 September 2018
 
-N fcasts, n IFPs
+- Use subset of forecasts from 2 May (turkers enter) to 1 August 2018 (change in tracking)
+
+~49,000 forecasts
+
+156 IFPs; 46 with machine forecasts
+
+971 unique users
+
+Forecasters:
+
+- Volunteers who joined the platform 
+- Amazon Mechanical Turkers
+- Machine (basil-ts)
 
@@
 
+--
+
 Volunteers are better forecasts than Amazon Turkers
@@
 
+--
+
 Human forecasters who saw the machine forecasts did poorly; volunteer forecasts who saw only charts had overall the best performance
 
+--
+
 It's not clear why though, e.g.:
+
+--
+
 - they outperformed on questions that did not have data (charts and model)
+
+--
+
 - they outperformed volunteer forecasters who saw the machine forecasts even when the machine forecasts had good accuracy
+
+--
 
@@
 
-# Model overview
+# Relative forecaster performance
+
+--
 
@@
 
-Compare average Brier scores of volunteer, turker, and machine forecasts (reference category)
+--
+
+Linear model to compare average Brier scores of volunteer, turker, and machine forecasts (reference category)
 
@@
 
+--
+
 $\beta_i$ = average Brier score for machine forecasts for question group $i$
+
+--
 
@@
 
+--
+
 - Negative values $\rightarrow$ human forecasters did better
+
+--
+
 - Positive values $\rightarrow$ machine forecasts did better
@@
 
-![:scale 80%](img/model-machine-by-data-source.png)
+![](img/model-machine-by-data-source-buildup.png)
+
+???
+
+- explain the two panels
+- x-axis left is average machine forecast performance, right is difference from left; both are Brier scores
+- y-axis are question groups, number is number of questions
+- sorted so that at top are better human, at bottom are better machine
+
+---
+class: center, middle
+
+![](img/model-machine-by-data-source.png)
+
+???
+
+- overall easiest, which can't see directly were oil prices, stocks
+- overall hardest were ACLED, OECD interest rates, and Boko Haram
+- humans generally tended to do better on questions with monthly period (food prices, oil production)
+- machines more clearly did better on questions that require data aggregation (ACLED, Nigeria, hacking); earthquakes aggregates are readily available
+- ACLED 2op versus 5op: maybe partly 0's, partly truncation issue
 
@@
 
+Overall hard questions; machine outperformed human forecasts
+
+> How many battle deaths will ACLED record in Central African Republic in August 2018?
+
+![](img/chart-2117.png)
+
 ---
@@
 
+Overall hard questions; machine has middling performance
+
+> What will be the long-term interest rate for Hungary (HUN) in July 2018?
+
+![](img/chart-1793.png)
+
 ---
@@
 # ACLED binary question
+
+Relatively easy questions, but machines underperformed a lot
+
+> Will ACLED record any riot/protest events in Gambia in July 2018?
+
+![](img/chart-1136.png)
 
@@
 
+- ACLED
+- Privacyrights hacking
+- Nigeria security tracker (Boko Haram)
+
+--
+
 Some of the overall hardest questions, and where to some extent human forecasters did better, are economic/financial monthly series
@@
 - FAO food price indices
+- exchange rates
+- oil production
+
+--
+
+---
+
+# Issues we did not address
+
+--
+
+What about selection issues; are volunteer forecasters able to self-select into questions they will do well on?
+
+--
+
+Do the chart volunteers do better on questions requiring data aggregation; generally there are some inconsistencies on why/how the chart volunteers did so well.
 
@@
 
-Register to forecast at:
-
-Internal platform:
+Register to forecast at https://sage-platform.isi.edu/
 

### 2019-01-04 R/analysis.R

@@ -8,6 +8,7 @@ library("rstan")
 library("rstanarm")
 library("forcats")
 library("broom")
+library("Hmisc")
 
 dropbox_path <- "~/Dropbox/Work/asia-polmeth-2019"
 in_dir  <- file.path(dropbox_path, "input-data")
@@ -65,9 +66,20 @@ common_period <- all_fcasts %>%
 
 nrow(common_period)
 
+# Unique IFPs
+length(unique(common_period$ifp_id))
+
+# IFPs with machine forecasts
+length(unique(common_period[common_period$Forecaster=="Machine", ]$ifp_id))
+
+# Unique forecasters
+length(unique(common_period$user_id))
+
+# N. forecasts by user group
 common_period %>%
-  group_by(Forecaster) %>%
-  summarize(n = n())
+  dplyr::group_by(Forecaster) %>%
+  dplyr::summarize(n_forecasts = n(),
+                   n_users = length(unique(user_id)))
 
 by_ifp <- common_period %>%
   group_by(ifp_id) %>%
@@ -251,8 +263,30 @@ arima_qs <- common_period %>%
   mutate(Data_source = case_when(
     data_source=="acled" & num_options==2 ~ "ACLED 2op",
     data_source=="acled" & num_options==5 ~ "ACLED 5op",
-    TRUE ~ data_source
+    data_source=="fao-fpi" ~ "FAO global food prices",
+    data_source=="oecd" ~ "OECD interest rates",
+    data_source=="eia" ~ "EIA oil prices",
+    data_source=="privacyrights" ~ "Privacyrights hacking",
+    data_source=="fred" ~ "FRED exchange rates",
+    data_source=="opec" ~ "OPEC oil production",
+    data_source=="imf" ~ "IMF inflation",
+    data_source=="nigeria-security" ~ "Nigeria Boko Haram violence",
+    TRUE ~ Hmisc::capitalize(data_source)
   ))
+
+# overall data source difficulty
+arima_qs %>%
+  dplyr::group_by(Data_source) %>%
+  dplyr::summarize(avg_Brier = mean(brier, na.rm = TRUE)) %>%
+  arrange(avg_Brier)
+
+
+# Add # of questions to each Data source label
+arima_qs <- arima_qs %>%
+  group_by(Data_source) %>%
+  mutate(by_Data_source_n_ifps = length(unique(ifp_id))) %>%
+  ungroup() %>%
[…]

### 2019-01-04 R/opec-iraq-example.R

@@ -4,8 +4,12 @@
 
 library("tidyverse")
 library("cowplot")
+library("lubridate")
+library("forecast")
 
 df <- read_csv("input-data/iraq-oil-production.csv")
+
+cutpoints <- c(4280, 4384, 4473, 4576)
 
 p1 <- df %>%
   filter(date >= as.Date("2018-06-01") %m-% months(120)) %>%
@@ -35,26 +39,40 @@ ggdraw() +
 ggsave(file = "output/figures/opec-iraq-divergence.png", height = 5, width = 8)
 
 
-plot.mpg <- ggplot(mpg, aes(x = cty, y = hwy, colour = factor(cyl))) + 
-  geom_point(size=2.5) + theme_cowplot()
-plot.mpg
-plot.diamonds <- ggplot(diamonds, aes(clarity, fill = cut)) + geom_bar() +
-  theme_cowplot() + theme(axis.text.x = element_text(angle=70, vjust=0.5, hjust = 0.9))
-plot.diamonds
+# 
+#   Example of chart and chart + forecast for slides
+#   ____________________________
 
-ggdraw() +
-  draw_plot(
-    plot.diamonds + theme(legend.justification = "bottom"),
-    0, 0, 1, 1
-  ) +
-  draw_plot(
-    plot.mpg + scale_color_viridis_d() + 
-      theme(legend.justification = "top", plot.background = element_rect(fill = "white")),
-    0.5, 0.54, 0.5, 0.4
-  ) +
-  draw_plot_label(
-    c("A", "B"),
-    c(0, 0.5),
-    c(1, 0.94),
-    size = 15
-  )
+train_df <- df %>%
+  filter(date <= as.Date("2018-02-01")) %>%
+  filter(date >= as.Date("2018-02-01") %m-% months(60)) %>%
+  filter(source=="Platform / tradingeconomics.com")
+
+p1 <- train_df %>%
+  tail(12*4) %>%
+  ggplot() +
+  geom_line(aes(x = date, y = value)) +
+  theme_minimal() +
+  labs(x = "", y = "") +
+  scale_x_date(limits = c(as.Date("2018-02-01") %m-% months(12*4), as.Date("2018-05-01"))) +
+  scale_y_continuous(limits = c(2700, 5000))
+ggsave(p1, file = "output/figures/opec-iraq-with-chart.png", height = 4, width = 8)
+
[…]

### 2019-01-04 R/analysis.R
update slides

@@ -92,8 +92,8 @@ by_ifp
 
 # Overall condition table
 brier_by_forecaster <- common_period %>%
-  group_by(Forecaster) %>%
-  summarize(avg_Brier = mean(brier), n = n()) %>%
+  dplyr::group_by(Forecaster) %>%
+  dplyr::summarize(avg_Brier = mean(brier), n = n()) %>%
   write_csv(., "output/tables/brier-by-forecaster.csv")
 brier_by_forecaster %>%
   knitr::kable(digits = 3)

### 2019-01-04 slides/index.Rmd
fix redundant transition
update slides

@@
 author: "Predictive Heuristics </br>Andreas Beger and Michael D. Ward"
-date: "`r format(Sys.Date(), '%e %B %Y')` </br>Asian POLMETH 2019, Kyoto, Japan"
+date: "`r format(Sys.Date(), '%e %B %Y')` </br>Asia POLMETH 2019, Kyoto, Japan"
 output: 
@@
 
---
-
 If data for a particular question is available:
@@
 
---
-
 How well are the machine forecasts doing?
-
---
 
@@
 
+---
+
+# Performance by forecaster group
+
+|Forecaster | avg Brier|     N|
+|:----------|---------:|-----:|
+|Machine    |     0.402|  1975|
+|Turker     |     0.433| 39140|
+|Volunteer  |     0.322|  7816|
+
+---
+
 # Relative forecaster performance
 
---
-
 Group questions by data source, N=14 groups
-
---
 
@@
 
---
-
 $\beta_i$ = average Brier score for machine forecasts for question group $i$
-
---
 
@@
 
---
-
 - Negative values $\rightarrow$ human forecasters did better
-
---
-
 - Positive values $\rightarrow$ machine forecasts did better
@@
 
---
-
 ---
